### 2019-05-19 main.tex
Initial Overleaf Import

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

% Paper specific includes and commands
\usepackage{times}
\usepackage{soul}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\urlstyle{same}
\usepackage{xargs}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{import}
\usepackage{svg}
\usepackage{pdfpages}
\usepackage[norelsize, linesnumbered, ruled, lined, boxed, commentsnumbered]{algorithm2e}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage[lofdepth,lotdepth]{subfig}

% Custom math commands
\input{mathdef}

% Todo notes
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\usetikzlibrary{patterns}
\newcommand{\quotes}[1]{``#1''}
\newcommandx{\nc}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{Nicolas:} #2}}
\newcommandx{\op}[2][1=]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Olivier:} #2}}
\newcommandx{\el}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{\textbf{Edouard:} #2}}
\newcommandx{\td}[2][1=]{\todo[inline,size=\large,#1]{#2}}

% Resources path
\makeatletter
\providecommand*{\input@path}{}
\g@addto@macro\input@path{{./source/}{./source/ressources/}{./ressources/}}
\makeatother

%---------------------------------------------
% Cross Referencing of Supplementary Material %---------------------------------------------
% https://tex.stackexchange.com/questions/14364/cross-referencing-between-different-files
% https://tex.stackexchange.com/questions/244239/correct-ordering-of-xr-hyper-hyperref-and-cleveref
% \usepackage{xr}
% \usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage[capitalise,noabbrev]{cleveref}
% \makeatletter
% \newcommand*{\addFileDependency}[1]{% argument=file name and extension
%   \typeout{(#1)}
%   \@addtofilelist{#1}
%   \IfFileExists{#1}{}{\typeout{No file #1.}}
% }
% \makeatother
% \newcommand*{\myexternaldocument}[1]{%
%     \externaldocument{#1}%
%     \addFileDependency{#1.tex}%
%     \addFileDependency{#1.aux}%
% }
% \myexternaldocument{supplementary}

\title{Budgeted Reinforcement Learning in Continuous State Space}

\author{
Nicolas Carrara$^{1,}$\thanks{equal contribution}\and
Edouard Leurent$^{1,3,}$\footnotemark[1]\and
Romain Laroche$^{4}$\and
Tanguy Urvoy$^{2}$\and
Odalric Maillard$^{1}$\and
Olivier Pietquin$^{1,2,}$\thanks{now with Google Brain, Paris.}\\
\affiliations $^1$SequeL team, INRIA Lille -- Nord Europe, France\\
$^2$Orange Labs, Lannion, France\\
$^3$Renault Group, France\\
$^4$Microsoft Research , Montreal, Canada
\emails
\{nicolas.carrara, edouard.leurent, odalric.maillard, olivier.pietquin\}@inria.fr,
romain.laroche@microsoft.com, tanguy.urvoy@orange.com
}

\begin{document}
\maketitle
\begin{abstract}
    A Budgeted Markov Decision Process (BMDP) is an extension of a Markov Decision Process to critical applications requiring safety constraints. It relies on a notion of risk implemented in the shape of a cost signal constrained  to lie below a threshold that -- importantly -- can be modified in real-time. So far, BMDPs could only be solved in the case of finite state spaces with known dynamics. This work extends the state-of-the-art to continuous spaces environments and unknown dynamics. We show that the solution to a BMDP is a fixed point of a novel Budgeted Bellman Optimality operator. This observation allows us to introduce natural extensions of Deep Reinforcement Learning algorithms to address large-scale BMDPs. We validate our approach on two simulated applications: spoken dialogue and autonomous driving.
\end{abstract}

% @Nicolas/Edouard: faites ici la liste des sections qui sont pretes a etre relues.
% Section 1: il manque les contributions
% Section 2: OK. On pense éventuellement rajouter une hypothèse de contractance de Tau (pas sûr que ce soit le cas)
% Section 3: OK
% Section 4.1: OK dans un premier temps, mais on va peut-être modifier. En fait, on remplace pi_greedy(Q) par pi_hull(Q), et on ne peut faire cela que lorsque Q est concave. Deux possibilité: soit on dit juste que Tau_greedy est remplacé par Tau_hull, qui admet également Q* comme point fixe (Tau_greedy(Q*) = Tau_hull(Q*) puisque Q* concave d'après le lemme). Soit on est intéressé par ce qui se passe le long de la trajectoire Q_k (par exemple si on suppose Tau_greedy contractant), et pour conserver cette propriété il faudrait alors montrer que Tau conserve la concavité et que donc on a bien Tau_hull = Tau_greedy tout le long de la trajectoire. Il faut aussi reprendre les notations des graphiques.
% Section 4.2 et 4.3: KO. Pas encore retravaillé, mais ça ne devrait pas trop bougé
% Section 5: OK pour environnements et results, KO pour discussion
% Section 6: KO
% Section 7: KO

\section{Introduction}

Reinforcement Learning (RL) is a general framework for decision-making under uncertainty. It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R, \gamma)$ with continuous state space $\cS$, discrete actions $\cA$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ and rewards $R\in\Real^{\cS \times \cA}$, where $\mathcal{M}(\mathcal{X})$ denotes the set of probability measures over a measurable space $\mathcal{X}$. Formally, we seek a policy $\pi\in\cM(A)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R(s_t, a_t)$.

However, this modelling assumption comes at a price: no control is given over the spread of the performance distribution. In many critical real-world applications where failures may turn out very costly, this is an issue as most decision-makers would rather give away some amount of expected optimality to increase the performances in the lower-tail of the distribution. This has led to the development of several risk-averse variants where the optimisation criteria include other statistics of the performance, such as the worst-case realisation \citep{Iyengar2005,Nilim2005,Wiesemann2013}, the variance-penalised expectation \citep{Garcia2015,Tamar2012}, or the Conditional Value-At-Risk (CVaR) \citep{Chow2014,DBLP:journals/corr/ChowGJP15}.

Reinforcement Learning also assumes that the performance can be described solely by a reward function $R$. Conversely, real problems typically involve many aspects, some of which can be contradictory \citep{ChunmingLiu2014}. For instance, a self-driving car needs to balance between progressing quickly on the road and avoiding collisions. When aggregating several objectives in a single scalar signal, as often in Multi-Objectives RL~\citep{Roijers2013ASO}, no control is given over their relative ratios, as high rewards can compensate high penalties. For instance, if a weighted sum is used to balance velocity $v$ and accidents $a$, then for any given choice of weights $\omega$ the optimality equation $\omega_v\expectedvalue[\sum\gamma^t v_t] + \omega_a\expectedvalue[\sum\gamma^t a_t] = G_r^* = \max_\pi G^\pi_r$ is the equation of a line in $(\expectedvalue[\sum\gamma^t v_t], \expectedvalue[\sum\gamma^t a_t])$, and the automotive company cannot control where its optimal policy $\pi^*$ lies on that line.

Both of these concerns can be addressed in the \emph{Constrained Markov Decision Process} (CMDP) setting \citep{BEUTLER1985236,Altman95constrainedmarkov}. In this multi-objective formulation, task completion and safety are considered separately. We equip the MDP with a cost signal $C \in \Real^{\cS\times \cA}$ and a cost budget $\beta\in\Real$. Similarly to $G_r$, we define the return of costs $G_c^\pi = \sum_{t=0}^\infty \gamma^t C(s_t, a_t)$ and the new cost-constrained objective:
\begin{equation}
\label{eq:cmdp}
\begin{array}{lcr}
 \max_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s, \pi] & \text{ s.t. } & \expectedvalue[G_c^\pi | s_0=s, \pi] \leq \beta
\end{array}
\end{equation}

This constrained framework allows for better control of the performance-safety tradeoff. However, it suffers from a major limitation: the budget has to be chosen before training, and cannot be changed afterwards.

To address this concern, the \emph{Budgeted Markov Decision Process} (BMDP) was introduced in \citep{Boutilier_Lu:uai16} as an extension of CMDPs to enable the online control over the budget $\beta$ within an interval $\cB \subset \Real$ of admissible budgets. Instead of fixing the budget prior to training, the objective is now to find a generic optimal policy $\pi^*$ takes $\beta$ as input so as to solve the corresponding CMDP \eqref{eq:cmdp} for all $\beta\in\cB$. This gives the system designer the ability to move the optimal policy $\pi^*$ in real-time along the Pareto-optimal curve of the different reward-cost trade-offs.

Our first contribution is to re-frame the original BMDP formulation in the context of continuous states and infinite discounted horizon. We then propose a novel Budgeted Bellman Optimality Operator and prove the optimal value function to be a fixed point of this operator. Second, we use this operator in \BFTQ, a Reinforcement Learning algorithm for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure. Third, we scale this algorithms to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex computing, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing. Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.

\section{Budgeted Dynamic Programming}

We will work in the space of budgeted policies, where $\pi$ both depends on $\beta$ and also outputs a next budget $\beta_a$. Hence, the budget $\beta$ is neither fixed nor constant as in the CMDP setting but instead evolves as part of the dynamics.

For ease of notation, we consider the \emph{augmented} state and action spaces $\ocS = \cS\times \cB$ and $\ocA= \cA\times \cB$, and equip them with the augmented dynamics $\ov{P}\in \cM(\ocS)^{\ocS \times \ocA}$ defined as:
\begin{equation*}
    \ov{P}\left(\os' \condbar \os, \oa\right) = \ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) = P(s'|s, a)\delta(\beta' - \beta_a)
\end{equation*}
where $\delta$ is the Dirac distribution.

In other words, in these augmented dynamics, the output budget $\beta_a$ returned at time $t$ by a budgeted policy $\pi\in \Pi=\cM(\ocA)^{\ocS}$ will be used to condition the policy at the next timestep $t+1$.

We stack the rewards and cost functions in a single \emph{vectorial} signal $\ov{R} \in (\Real^2)^{{\ocS \times \ocA}}$.
Given an augmented transition $(\os, \oa) =((s,\beta), (a, \beta_a))$, we define:
\begin{equation}
     \ov{R}(\os, \oa) =  \begin{bmatrix}
     R(s, a)\\
     C(s, a)
     \end{bmatrix}\in\Real^2
\end{equation}

As earlier, the return $G^\pi = (G_r^\pi, G_c^\pi)$ of a budgeted policy $\pi\in\Pi$ refers to:
% \begin{equation*}
    $G^\pi = \sum_{t=0}^\infty \gamma^t \ov{R}(\os_t, \oa_t)$.
% \end{equation*}

We can introduce the value function $V^\pi$ and action-value function $Q^\pi$ of a budgeted policy $\pi\in\Pi$ as:
% \begin{align}
% \label{eq:value-function}
%     V^\pi(\os) &= \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \\
%     Q^\pi(\os, \oa) &= \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right]
% \end{align}
\begin{equation}
    \label{eq:value-function}
V^\pi(\os) = \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa) = \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right]
\end{equation}

Likewise, we note $V^\pi = (V_r^\pi, V_c^\pi)$ and $Q^\pi = (Q_r^\pi, Q_c^\pi)$.

\begin{proposition}[Budgeted Bellman Expectation]
\label{prop:bellman-expectation}
The value functions $V^\pi, Q^\pi$ of a budgeted policy $\pi$ verify:
\begin{align}
    V^\pi(\os) &= \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa)\\
    Q^\pi(\os, \oa) &= \ov{R}(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') \label{eq:bellman_expectation_Q}
\end{align}
\end{proposition}

\begin{definition}[Budgeted Optimality]
We now come to the definition of budgeted optimality. We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the rewards $R$, (iii)~in case of tie, minimise the costs $C$. To that end, we define the following sets for all $\os\in\ocS$:
\begin{itemize}
    \item Admissible policies: $\Pi_a(\os) = \{\pi\in\Pi: V_c^\pi(\os) \leq \beta\}$ where $\os = (s, \beta)$.
    \item Best admissible policies: $\Pi_r(\os) = \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)$
    \item Optimal policies: $\Pi^*(\os) = \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)$
\end{itemize}
\end{definition}

\begin{assumption}[Existence of optimal policy]
There exists one budgeted policy $\pi^*$ that uniformly maximises $V_r^\pi$ under constraint on $V_c^\pi$ for all augmented states $\os$.
\begin{align*}
    \exists \pi^*\in\Pi: \forall \os\in \ocS, \pi^*\in\Pi^*(\os)
\end{align*}
\end{assumption}

Let $\pi^*$ be one such optimal budgeted policy, we finally define the optimal budgeted value functions $V^*$, $Q^*$. For all $\os, \oa \in \ocS\times\ocA$:
\begin{equation*}
    V^*(\os) = V^{\pi^*}(\os), \qquad Q^*(\os, \oa) =  Q^{\pi^*}(\os, \oa),
\end{equation*}

\begin{remark}[Independence on the representative]
\label{rmk:independence}
Though they seemingly depend on the choice of $\pi^*$, the value functions $V^*$ and $Q^*$ are well-defined because all optimal policies in $\Pi^*$ share the same $V_r, Q_r$ as they belong to $\Pi_r$, and the same $V_c, Q_c$  as they belong to $\Pi^*$.
\end{remark}

\begin{theorem}[Budgeted Bellman Optimality]
\label{thm:bellman-optimality}
The optimal action-value of a budgeted policy verifies:
\begin{equation}
\label{eq:bellman-optimality}
    Q^{*}(\os, \oa) = \cT Q^{*}(\os, \oa) \eqdef \ov{R}(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'})
\end{equation}
where the greedy policy $\pi_\text{greedy}$ is defined by: $\forall \os=(s,\beta)\in \ocS, \oa\in 
\ocA, \forall Q\in(\Real^2)^{\ocA\times\ocS},$
\begin{subequations}
\label{eq:pi_greedy}
\begin{equation}
    \pi_\text{greedy}(\oa|\os; Q) \in \argmin_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa) \label{eq:pi_greedy_cost}
\end{equation}
\begin{align}
    \text{where }\quad\Pi_r^Q = &\argmax_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \label{eq:pi_greedy_reward}\\
    & \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \leq \beta \label{eq:pi_greedy_constraint}
\end{align}
\end{subequations}
\end{theorem}

\begin{remark}[Appearance of the greedy policy]
\label{rmk:greedy}
In classical Reinforcement Learning, the greedy policy takes a simple form $\pi_\text{greedy}(s; Q^*) = \argmax_{a\in\cA} Q^*(s, a)$, and the term $\pi_\text{greedy}(a'|s';Q^*) Q^{*}(s', a')$ in \eqref{eq:bellman-optimality} conveniently simplifies to $\max_{a'\in \cA} Q^*(s', a')$. Unfortunately, in a budgeted setting the greedy policy requires solving the nested constrained optimisation program \eqref{eq:pi_greedy} at each state and budget in order to apply this Budgeted Bellman Optimality operator.
\end{remark}

\begin{remark}[Notations]
Because of the budget dynamics $\beta' = \beta_a$, we can see in \eqref{eq:bellman_expectation_Q} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$. We will slightly abuse notations and denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
\end{remark}

\paragraph{Budgeted Value Iteration}

The Budgeted Bellman Optimality equation is a fixed-point equation, which motivates the introduction of a fixed-point iteration procedure. We introduce Algorithm~\ref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.

\section{Budgeted Reinforcement Learning}

In this section, we consider BMDPs with unknown parameters that must be solved by interaction with an environment.

\subsection{Budgeted Fitted-Q}

When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(s_i,\beta_i,a_i,R_i,C_i,s_i')\}_{i\in [0,N]}$ collected by interaction with the environment. Applying $\cT$ in \eqref{eq:bellman-optimality} requires computing an expectation $\expectedvalueover{\os'\sim \ov{P}}$ over next states $\os'$, which we can replace by an empirical average $\sum_{s'\in\cD}$.
We introduce in Algorithm~\ref{algo:bftq} the \emph{Budgeted-Fitted-Q} (BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs. 
Because we work on samples of a continuous state space $\cS$, we need to employ function-approximation in order to generalise to nearby states. Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $L(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, \ov{R}, \os')||_2^2$.
Any model can be used, such as linear models, regression trees, or neural networks.

\begin{minipage}[t]{0.5\textwidth}
\vspace{0pt}  
\begin{algorithm}[H]
\label{algo:bvi}
\DontPrintSemicolon
\KwData{$P, R, C$}
\KwResult{$Q^*$}
$Q_{0} \leftarrow 0$\;
\Repeat{convergence}{
    $Q_{k+1} \leftarrow \cT Q_k$\;
}
\caption{Budgeted Value Iteration}
\end{algorithm}
\end{minipage}%
\begin{minipage}[t]{0.5\textwidth}
\vspace{0pt}
\begin{algorithm}[H]
\label{algo:bftq}
\DontPrintSemicolon
\KwData{$\cD$}
\KwResult{$Q^*$}
$Q_{0} \leftarrow 0$\;
\Repeat{convergence}{
    $Q_{k+1} \leftarrow \argmin_\theta \cL(Q_\theta, \cT Q_k; \cD)$\;
}
\caption{Budgeted Fitted-Q}
\end{algorithm}
\end{minipage}

\subsection{Risk-sensitive exploration}

\label{sec:exploration}

In order to run Algorithm~\ref{algo:bftq}, we must first gather a batch of samples $\mathcal{D}$. Ideally we would need samples from the asymptotic state-budget distribution $\lim_{t\rightarrow\infty}\probability{\os_t}$ induced by an optimal policy $\pi^*$ given an initial distribution $\probability{\os_0}$, but as we are actually building this policy, it is not possible. Following the same idea of $\epsilon$-greedy exploration for \FTQ \citep{Ernst2005,Riedmiller2005}, we introduce an algorithm for risk-sensitive exploration. We follow an exploration policy: a mixture between a random budgeted policy $\pi_\text{rand}$ and the current greedy policy $\pi_\text{greedy}$. The batch $\cD$ is split into several mini-batches generated sequentially, and $\pi_\text{greedy}$ is updated by running Algorithm~\ref{algo:bftq} on $\cD$ upon mini-batch completion. $\pi_\text{rand}$ is designed to obtain trajectories that only explore feasible budgets: we impose that the joint distribution $\probability{a, \beta_a}$ verifies $\expectedvalue[\beta_a]\leq\beta$. This condition defines a probability simplex $\Delta_{\ocA}$ from which we sample uniformly. Finally, when interacting with an environment the initial state $s_0$ is usually sampled from a starting distribution $\probability{s_0}$. In the budgeted setting, we also need to sample the initial budget $\beta_0$. Importantly, we pick a uniform distribution $\probability{\beta_0} = \cU(\cB)$ so that the entire range of risk-level is explored, and not only reward-seeking behaviours as would be the case with a traditional risk-neutral $\epsilon$-greedy strategy. The pseudo-code of our exploration procedure is shown in Algorithm~\ref{algo:risk-sensitive-exploration} in the Supplementary Material.

\section{A Scalable Implementation}

\label{sec:scalable-bftq}
In this section, we introduce an implementation of the BFTQ algorithm designed to operate efficiently and handle large batches of experiences $\mathcal{D}$.

\subsection{How to compute the greedy policy?}

As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested non-linear constrained optimisation program \eqref{eq:pi_greedy}. However, it can be solved efficiently by exploiting the \emph{structure} of the optimal value function $V^*$. % In the following, we say that a vector-valued function is concave and non-decreasing with respect to a variable if all of its components are.

%\begin{lemma}[Concavity and monotony of $Q^*$]
%\label{lemma:concavity}
%$Q^*$ is non-decreasing concave with respect to $\beta_a$. 
%\end{lemma}
%\begin{lemma}[Conservation of concavity and monotony by $\cT$]
%\label{lemma:tau_concavity}
%For any $Q\in(\Real^2)^{\ocS\ocA}$ non-decreasing concave with respect to $\beta_a$, $\cT Q$ is also non-decreasing concave with respect to $\beta_a$.
%\end{lemma}

%These two Lemmas can be used to derive an efficient computation method for $\pi_\text{greedy}$.

%\begin{lemma}[Hull policy equality]
%\label{lemma:pi_hull}
%For any $Q\in(\Real^2)^{\ocS\ocA}$ non-decreasing concave with respect to $\beta_a$, $\pi_\text{greedy}$ in \eqref{eq:pi_greedy} can be computed by running Algorithm~\ref{algo:pi_hull}.
%\begin{equation*}
%    \pi_\text{greedy}(\oa|\os; Q) = \pi_\text{hull}(\oa|\os; Q)
%\end{equation*}
%\end{lemma}

\begin{proposition}[Equality of $\pi_\text{greedy}$ and $\pi_\text{hull}$]
\label{prop:bftq_pi_hull}
Algorithm~\ref{algo:bvi} and Algorithm~\ref{algo:bftq} can be run by replacing $\pi_\text{greedy}$ in the definition \eqref{eq:bellman-optimality} of $\cT$ with $\pi_\text{hull}$ as described in Algorithm~\ref{algo:pi_hull}.
\begin{equation*}
    \pi_\text{greedy}(\oa|\os; Q) = \pi_\text{hull}(\oa|\os; Q)
\end{equation*}
\end{proposition}

\begin{algorithm}
\DontPrintSemicolon
\KwData{$s$, $Q$}
$\cH \leftarrow \texttt{convex\_hull}(Q(s,\cA, \cB))$\;
denote $q^i = Q(s,a^i,\beta_a^i)\in\cH$ its $i^{\text{th}}$ point in clockwise order\;
find two successive points $((q_c^0, q_r^0), (q_c^1, q_r^1))$ of $\cH$ such that $q_c^0 \leq \beta < q_c^1$\;
\lIf{$q_r^0 = q_r^1$}{
\Return $\delta(a^0, \beta_a^0)$
}\Else{
$p \leftarrow (\beta - q_c^0) / (q_c^1 - q_c^0)$\;
\Return the mixture $(1-p)\delta(a^0, \beta_a^0) + p\delta(a^1, \beta_a^1)$\;
}
\caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
\label{algo:pi_hull}
\end{algorithm}

\begin{minipage}[t]{0.5\textwidth}
\vspace{0pt}
\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{source/img/qrqc}
    \caption{Computation of $\pi_\text{hull}$ for a concave non-decreasing Q-function $Q$. The dotted tangent on the right defines the greedy policy when the budget $\beta$ lies between $Q_c(s,a^0,\beta_a^0)$ and $Q_c(s,a^1,\beta_a^1)$, as mixtures of the two points $(0, 1)$ represented as green crosses}
    \label{fig:hull}
\end{figure}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
\begin{figure}[H]
    \centering
    \input{source/architecture.tex}
    \caption{Neural Network for $Q$-functions approximation when the state dimension is 2 and there are 2 actions.}
    \label{fig:architecture}
\end{figure}
\end{minipage}

\subsection{Function approximation}

Neural networks are well suited to model Q-functions in Reinforcement Learning algorithms \citep{Riedmiller2005,Mnih2015}.  We approximate $Q = (Q_r, Q_c)$ using one single neural network. Thus, the ywo components are jointly optimised which accelerates convergence and fosters learning of useful shared representations. Moreover, as in \citep{Mnih2015} we are dealing with a finite (categorical) action space $\cA$, instead of including the action in the input we add the output of the $Q$-function for each action to the last layer. Again, it provides a faster convergence toward useful shared representations and it only requires one forward pass to evaluate all action values. Finally, beside the state $s$ there is one more input to a budgeted $Q$-function:~the budget $\beta_a$. This budget is a scalar value whereas the state $s$ is a vector of potentially large size. To avoid a weak influence of $\beta$ compared to $s$ in the prediction, we include an additional encoder for the budget, whose width and depth may depend on the application. A straightforward choice is a single layer with the same width as the state. The overall architecture is shown in Fig.~\ref{fig:architecture}.

\subsection{Parallel computing}

In a simulated environment, a first process that can be distributed is the collection of samples in the exploration procedure of Algorithm~\ref{algo:risk-sensitive-exploration}, as $\pi_\text{greedy}$ stays constant within each mini-batch which avoids the need of synchronisation between workers. Second, the main bottleneck of \BFTQ is the computation of the target $\cT Q$. Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $O(|\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$. The resulting total time-complexity is $O(|\mathcal{D}|/(1-\gamma) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$. This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA,\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass. By using multiprocessing in the computations of $\pi_\text{hull}$, we obtain a linear speedup.% and reduce the time-complexity to $O(|\mathcal{D}|/(W(1-\gamma)) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$.
The full description of our scalable implementation of \BFTQ is recalled in Algorithm~\ref{algo:bftq_full} in Supplementary Material~\ref{sec:bftq-full}.

\section{Experiments}

There are two hypotheses we want to validate.

\paragraph{Exploration strategies} We claimed in Section~\ref{sec:exploration} that a risk-sensitive exploration was required in the setting of BMDPs. We test this hypotheses by confronting our strategy to a classical risk-neutral strategy. The latter is chosen to be a $\epsilon$-greedy policy slowly transitioning from a random to a greedy policy\footnote{We train this greedy policy using \FTQ.} that aims to maximise $\expectedvalue_{\pi} G_r^\pi$ regardless of $\expectedvalue_{\pi} G_c^\pi$. The quality of the resulting batches $\cD$ is assessed by training a BFTQ policy and comparing the resulting performance.

\paragraph{Budgeted algorithms} We compare our  scalable \BFTQ algorithm described in Section~\ref{sec:scalable-bftq} to an \FTQl baseline. This baseline consists in approximating the BMDP by a finite set of CMDPs problems. We solve each of these CMDP using the standard technique of Lagrangian Relaxation: the cost constraint is converted to a soft penalty weighted by a Lagrangian multiplier $\lambda$ in a surrogate reward function: $\max_{\pi} \expectedvalue_{\pi} G_r^\pi - \lambda G_c^\pi\label{eq:lagrangian}$. The resulting MDP can be solved by any RL algorithm, and we chose $\FTQ$ for being closest to $\BFTQ$.
In these experiments, each run is composed of a training and a test phase and repeated $N_{\text{seeds}}$ times. Parameters of the algorithms can be found in Supplementary Material~\ref{sec:algorithms-parameters}

\subsection{Environments}

We evaluate our method on three different environments involving reward-cost trade-offs. Their parameters can be found in Supplementary Material~\ref{sec:env-parameters}

\paragraph{Corridors}
This simplistic environment is only meant to highlight clearly the specificity of exploration in a budgeted setting. It is a continuous gridworld with Gaussian perturbations, consisting in a maze composed of two corridors: a risky one with high rewards and costs, and a safe one with low rewards and no cost. In both corridors the outermost cell is the one yielding the most reward, which motivates a deep exploration.

\paragraph{Spoken dialogue system}
Our second application is a slot-filling simulation \citep{Li2009ReinforcementLF,chandramohan2010optimizing}. The system fills in a form of slot-values by interacting with a user through speech, before sending them a response. For example, in the restaurant domain, it may ask for three slots : the area of the restaurant, the price-range and the food type. The user could respectively provide those three slot-values : \texttt{Cambridge}, \texttt{Cheap} and \texttt{Indian-food}. In this application, we do not focus on how to extract such information from the user utterances, we rather focus on decision-making for filling in the form. To that end, the system can choose among a set of generic actions. As in \citep{carrara2018safe} there are two ways of asking for a slot value: a slot value can be either be provided with an utterance, which may to speech recognition errors with some probability, or by requiring the user to fill-in the slots by using a numeric pad. In this case, there are no recognition errors but a counterparty risk of hang-up: we assume that manually filling a key-value form is time-consuming and annoying. The environment yields a reward if all slots are filled without errors, and a constraint if the user hang-ups. Thus, there is a clear trade-off between using utterances and potentially committing a mistake, or using the numeric pad and risking a premature hang-up.
% Note that batch reinforcement learning has already been applied to spoken dialogue systems in \citep{Li2009ReinforcementLF,chandramohan2010optimizing} which makes this task a realistic candidate for optimisation with our method.

\paragraph{Autonomous driving}
In our third application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{Leurent2018} for simulated highway driving and behavioural decision-making.
We propose a novel task that displays a clear trade-off between safety and efficiency. The agent controls a vehicle with a finite set of manoeuvres implemented by low-lever controllers: $\mathcal{A}$ = \{\text{no-op}, \text{right-lane}, \text{left-lane}, \text{faster}, \text{slower}\}. It is driving on a two-lane road populated with other traffic participants: the vehicles in front of the agent drive slowly, and there are incoming vehicles on the opposite lane. Their behaviours are randomised, which introduces some uncertainty with respect to their possible future trajectories.
The task consists in driving as fast as possible, which is modelled by a reward proportional to the velocity: $R(s_t, a_t) \propto v_t$. This motivates the agent to try and overtake its preceding vehicles by driving fast on the opposite lane. This optimal but overly aggressive behaviour can be tempered through a cost function that embodies a safety objective: $C(s_t, a_t)$ is set to $1/H$ whenever the ego-vehicle is driving on the opposite lane, where $H$ is the episode horizon. Thus, the constrained signal $G_c^\pi$ is the maximum proportion of time that the agent is allowed to drive on the wrong side of the road.

\subsection{Results}

In the following figures, each patch represents the mean and 95\% confidence interval over $N_{\text{seeds}}$ seeds of the means of $(G_r^\pi,G_c^\pi)$ over $N_\text{trajs}$ trajectories\footnote{Examples of policy execution can be found in Supplementary Material~\ref{sec:bftq-executions}}. That way, we display the variation related to learning (and batches) rather than the variation in the execution of the policies.

We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: Figure~\ref{fig:exploration} shows the set of trajectories collected by each exploration strategy, and the resulting performance of a budgeted policy trained on each batch. The trajectories (purple) in the risk-neutral batch (left) are concentrated along the risky corridor (orange) and ignore the safe corridor (green), which results in bad performances in the low-risk regime. Conversely, trajectories in the risk-sensitive batch (right) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.
\begin{figure}[tp]
    \centering
    \includegraphics[width=0.25\textwidth]{source/img/risk-neutral.png}
    \includegraphics[width=0.25\textwidth]{source/img/risk-sensitive.png}
    \includegraphics[page=1, width=0.4\textwidth]{source/img/corridors}
    \caption{Trajectories (left) and performances (right) of two exploration strategies in the \text{corridors} environment. }
    \label{fig:exploration}
\end{figure}
% \begin{figure}[tp]
%     \begin{center}
%         \subfloat[\texttt{slot-filling}]{
%         \includegraphics[width=0.5\textwidth]{source/img/slot-filling}
%         \label{fig:slot-filling}
%         }
%         \subfloat[\texttt{highway-env}]{
%         \includegraphics[width=0.5\textwidth]{source/img/highway}
%         \label{fig:highway-env}
%         }
%         \caption{Performance comparison of \FTQl and BFTQ}
%         \label{fig:results}
%     \end{center}
% \end{figure}
\begin{figure}[tp]
    \begin{center}
    \includegraphics[width=0.45\linewidth]{source/img/slot-filling}
    \includegraphics[width=0.45\linewidth]{source/img/highway}
    \caption{Performance comparison of \FTQl and BFTQ on \text{slot-filling} (left) and \text{highway-env}(right) }
    \label{fig:results}
    \end{center}
\end{figure}

In a second experiment displayed in Fig.~\ref{fig:results}, we compare the performance of \FTQl to that of \BFTQ in the slot-filling and autonomous driving tasks. For each algorithm, we plot the reward-cost trade-off curve. In both cases, BFTQ performs almost as well as \FTQl despite only requiring a single model. All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can notice in the bottom-left corner an underestimation of $Q^c$, since $\expectedvalue[G_c|\beta=0] \simeq 0.1 $. However, these estimates still provide a reasonable prior approximation of the expected cost of the policy, which cannot be obtained with \FTQl. In addition, BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance. Qualitatively, the budgeted agents display a wide variety of behaviours. In the \texttt{highway-env} experiment, when $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic. On the contrary when $\beta = 0$, the ego-vehicle is conservative: it stays on its lane and drives at a low velocity. With intermediate budgets such as $\beta = 0.2$, the agent sometimes decides to overtake its front vehicle but promptly steers back to its original lane afterwards.

\section{Discussion}

Algorithm~\ref{algo:bftq} is an algorithm for solving large unknown BMDPs with continuous states. To the best of our knowledge, there is no algorithm in the current literature that combines all those features.

%As a side note , utiliser Survey ~\citep{Roijers2013ASO} et  reformuler ça qui vient du papier  \citep{le2019batch} : Multi-objective Reinforcement Learning. Another related area is multi-objective reinforcement learning (MORL)(Van Moffaert & Nowe´, 2014; Roijers et al., 2013). Generally, research in MORL has largely focused on approximating the Pareto frontier that trades-off competing objectives (Van Moffaert & Nowe´, 2014; Roijers et al., 2013).The underlying approach to MORL frequently relies on linear or non-linear scalarization of rewards to heuristicallyturns the problem into a standard RL problem. Our proposed approach represents another systematic paradigm to solve MORL, whether in batch or online settings.}
%The literature of Multi-objective Reinforcement Learning mainly focuses on recovering the Pareto-optimal frontier,

Algorithms have been proposed for CMDPs, which are less flexible sub-problems of the more general BMDP. When the environment parameters ($P$, $R$, $C$) are known but not tractable, solutions relying on function approximation~\citep{Undurti} or approximate linear programming~\citep{Poupart2015} have been proposed. For unknown environments, batch algorithms exist in the tabular setting \citep{Thomas2015, Petrik2016, Laroche2017}. Online algorithms \citep{Geibel2005, Abe2010,ChowGJP15,AchiamHTA17} and a batch algorithm \citep{le2019batch} combine the best of both world by solving intractable unknown CMDPs.

A first way of solving a BMDP is to approximate it with a finite set of CMDPs. The solutions of these CMDPs take the form of mixtures between two deterministic policies\citep[Theorem 4.4][]{BEUTLER1985236}. To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material.}. Our solution not only requires one single model but also avoids any supplementary interaction.

The closest work to ours is the Dynamic Programming algorithm proposed by \citet{Boutilier_Lu:uai16} for solving BDMPs. However, their work was established for finite state spaces only, and their solution relies heavily on this property. For instance, they enumerate and sort the next states $s'\in\cS$ by their expected value-by-cost, which could not be performed in a continuous state space $\cS$. Moreover, they rely on the knowledge of the model ($P$, $R$, $C$), and do not address the question of learning from interaction data.

\section{Conclusion}
The BMDP framework is a principled framework for safe decision making under uncertainty, which could be beneficial to the diffusion of Reinforcement Learning in industrial applications. However, BMDPs could so far only be solved in finite state spaces which limits their interest in many use-cases. We extend their definition to continuous states by introducing of a novel Dynamic Programming operator, that we build upon to propose a Reinforcement Learning algorithm. In order to scale to large problems, we provide an efficient implementation that exploits the structure of the value function and leverages tools from Deep Distributed Reinforcement Learning. We show that on two practical tasks, our solution performs similarly to a baseline Lagrangian relaxation method while only requiring a single model to train.

%    \section*{Acknowledgments}
%    %
%    This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
%
%    Thanks to Guillaume Gautier, Fabrice Clerot, Xuedong Shang

\bibliographystyle{named}
\bibliography{budgeted_rl}

\clearpage
\input{supplementary.tex}

\end{document}

### 2019-05-19 source/dialogues.tex
Initial Overleaf Import

\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}[]{lll}
\toprule

turn&$\beta=0$&$\beta=0.5$\tabularnewline
\midrule
turn 0& \makecell[l]{valid slots : [0, 0, 0]\\ srs : [ None None None ]\\ system says ASK\_ORAL(1) \\ user says INFORM} &\makecell[l]{ valid slots : [0, 0, 0]\\ srs : [ None None None ]\\ system says ASK\_ORAL(2)\\ user says INFORM}\tabularnewline\midrule
turn 1&\makecell[l]{valid slots : [0, 0, 0]\\ srs : [ None 0.48 None ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1]\\ srs : [ None None 0.56 ]\\ system says ASK\_ORAL(0)\\ user says INFORM}\tabularnewline\midrule
turn 2&\makecell[l]{valid slots : [0, 0, 0]\\ srs : [ None 0.48 0.22 ]\\ system says ASK\_ORAL(0) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1] \\ srs : [ 0.30 None 0.56 ]\\ system says ASK\_ORAL(1)\\ user says INFORM}\tabularnewline\midrule
turn 3&\makecell[l]{valid slots : [0, 0, 0]\\ srs : [ 0.62 0.48 0.22 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1]\\ srs : [ 0.30 0.54 0.56 ]\\ system says ASK\_ORAL(0)\\ user says INFORM}\tabularnewline\midrule
turn 4&\makecell[l]{valid slots : [0, 0, 0] \\ srs : [ 0.62 0.48 0.66 ]\\ system says ASK\_ORAL(1) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1]\\ srs : [ 0.68 0.54 0.56 ]\\ system says ASK\_NUM\_PAD(1)\\ user says INFORM}\tabularnewline\midrule
turn 5&\makecell[l]{valid slots : [0, 1, 0]\\ srs : [ 0.62 0.56 0.66 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 0.56 ]\\ system says SUMMARIZE\_AND\_INFORM\\ user says DENY\_SUMMARIZE}\tabularnewline\midrule
turn 6&\makecell[l]{valid slots : [0, 1, 0]\\ srs : [ 0.62 0.56 0.14 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 0.56 ]\\ system says ASK\_NUM\_PAD(2)\\ user says INFORM}\tabularnewline\midrule
turn 7&\makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.62 0.56 0.30 ]\\ system says ASK\_ORAL(2) \\ user says INFORM }& \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 1.00 ]\\ system says SUMMARIZE\_AND\_INFORM\\ user says DENY\_SUMMARIZE}\tabularnewline\midrule
turn 8&\makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.62 0.56 0.49 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 1.00 ]\\ system says ASK\_NUM\_PAD(0)\\ user hangs up !}\tabularnewline\midrule
turn 9&\makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.62 0.56 0.65 ]\\ system says SUMMARIZE\_AND\_INFORM \\ max size reached !}& \makecell[l]{ }\tabularnewline\bottomrule
\end{tabular}
}
    \caption{Two dialogues generated by a safe policy ($\beta=0$) on the left and a risky one ($\beta=0.5$) on the right.}
    \label{table:dialogues}
\end{table}

### 2019-05-19 supplementary.tex
Initial Overleaf Import

% \documentclass{article}

% % if you need to pass options to natbib, use, e.g.:
% %     \PassOptionsToPackage{numbers, compress}{natbib}
% % before loading neurips_2019

% % ready for submission
% % \usepackage{neurips_2019}

% % to compile a preprint version, e.g., for submission to arXiv, add add the
% % [preprint] option:
% %     \usepackage[preprint]{neurips_2019}

% % to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage{neurips_2019}

% % to avoid loading the natbib package, add option nonatbib:
% %     \usepackage[nonatbib]{neurips_2019}

% \usepackage[utf8]{inputenc} % allow utf-8 input
% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
% \usepackage{url}            % simple URL typesetting
% \usepackage{booktabs}       % professional-quality tables
% \usepackage{amsfonts}       % blackboard math symbols
% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
% \usepackage{microtype}      % microtypography

% % Paper specific includes and commands
% \usepackage{times}
% \usepackage{soul}
% \usepackage[small]{caption}
% \usepackage{graphicx}
% \usepackage{amsmath}
% \usepackage{amssymb}
% \usepackage{amsfonts}
% \usepackage{amsthm}
% \usepackage[shortlabels]{enumitem}
% \usepackage{algorithm}
% \usepackage{algorithmic}
% \usepackage{tabularx}
% \usepackage{makecell}
% \urlstyle{same}
% \usepackage{xargs}
% \usepackage{mathtools}
% \usepackage{dsfont}
% \usepackage{import}
% \usepackage{svg}
% \usepackage{pdfpages}

% \input{mathdef}

% \makeatletter
% \providecommand*{\input@path}{}
% \g@addto@macro\input@path{{./source/}{./source/ressources/}{./ressources/}}
% \makeatother

% \urlstyle{same}

% \usepackage{wrapfig}
% \usepackage{xr-hyper}
% \usepackage{hyperref}
% \usepackage[capitalise,noabbrev]{cleveref}

% % the following package is optional:
% %\usepackage{latexsym}

% \title{Supplementary material}

% \author{
% Nicolas Carrara$^{1,}$\thanks{equal contribution}\and
% Edouard Leurent$^{1,3,}$\footnotemark[1]\and
% Romain Laroche$^{4}$\and
% Tanguy Urvoy$^{2}$\and
% Odalric Maillard$^{1}$\and
% Olivier Pietquin$^{1,2,}$\thanks{now with Google Brain, Paris.}\\
% \affiliations $^1$SequeL team, INRIA Lille -- Nord Europe, France\\
% $^2$Orange Labs, Lannion, France\\
% $^3$Renault Group, France\\
% $^4$Microsoft Research , Montreal, Canada
% \emails
% \{nicolas.carrara, edouard.leurent, odalric.maillard, olivier.pietquin\}@inria.fr,
% romain.laroche@microsoft.com, tanguy.urvoy@orange.com
% }

% \begin{document}
% \maketitle

\begin{center}
\LARGE Supplementary material
\end{center}
\appendix

\section{Proof of Main Results}
\subsection{Proposition \ref{prop:bellman-expectation}}

\begin{proof}
\begin{align*}
    V^\pi(\os) &= \expectedvalue\left[ J^\pi \condbar \ov{s_0} = \os\right] \\
    &=\sum_{\oa\in\ocA} \probability{\oa_0 = \oa \condbar\ov{s_0} = \os} \expectedvalue\left[ J^\pi \condbar \ov{s_0} = \os, \oa_0 = \oa\right]\\
    &= \sum_{\oa\in\ocA} \pi(\oa | \os) Q^\pi(\os,\oa)
\end{align*}
\begin{align*}
    Q^\pi(\os, \oa) &= \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \ov{R}(\os_t, \oa_t)\condbar \ov{s_0} = \os, \ov{a_0} = \oa\right] \\
    &= \ov{R}(\os, \oa) + \sum_{\os'\in\ocS}\probability{\os_1 = \os' \condbar\ov{s_0} = \os, \ov{a_0} = \oa}\expectedvalue\left[\sum_{t=1}^\infty \gamma^t \ov{R}(\os_t, \oa_t)\condbar \ov{s_1} = \os'\right] \\
    &= \ov{R}(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar\os, \oa\right) \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \ov{R}(\os_t, \oa_t) \condbar \ov{s_0} = \os'\right] \\
    &=  \ov{R}(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar\os, \oa\right) V^\pi(\os')
\end{align*}
\end{proof}

\subsection{Theorem \ref{thm:bellman-optimality}}

\begin{proof}
Let $\os, \oa \in \ocA\times\ocS$. For this proof, we consider  potentially non-stationary policies $\pi=(\rho, \pi')$, with $\rho\in\cM(\ocA)$, $\pi'\in\cM(\ocA)^\Natural$. The results will apply to the particular case of stationary optimal policies, when they exist.

\begin{align}
    Q_r^*(\os, \oa) &=  r(\os, \oa) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa) V_r^{\pi*}(\os') \label{eq:pthm_def}\\
    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa)  \max_{\rho, \pi'} V_r^{\rho, \pi'}(\os') \label{eq:pthm_opt}\\
    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
\end{align}
where $\pi = (\rho, \pi')\in\Pi_a(\os)$ and $\pi'\in\Pi_a(\os')$.

This follows from:
\begin{enumerate}
\item[\eqref{eq:pthm_def}.] Definition of $Q^*$ and Bellman Expectation expansion from Proposition \ref{prop:bellman-expectation}. 
\item[\eqref{eq:pthm_opt}.] Definition of optimality.
\item[\eqref{eq:pthm_marg}.] Marginalisation on $\oa'$.
\item[\eqref{eq:pthm_max}.] \begin{itemize}
    \item Trivially $\max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\cA} \cdot \leq \sum_{\oa'\in\cA} \max_{ \pi'\in\Pi_a(\os)} \cdot$
    \item Let $\ov{\pi}\in\argmax_{\pi'\in\Pi_a(\os')} Q_r^{\pi'}(\os', \oa')$, then:
    \begin{align*}
        \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') &= \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')Q_r^{\ov{\pi}}(\os', \oa') \\
        &\leq  \max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\ov{A}}\rho(\oa'|\os')Q_r^{\pi'}(\os', \oa')
    \end{align*}
\end{itemize}
\item[\eqref{eq:pthm_marg_def2}.] Definition of $Q^*$.
\end{enumerate}

Moreover, the condition $\pi=(\rho, \pi')\in\Pi_a(\os)$ gives
\begin{equation*}
   \expectedvalueover{\oa'\sim\rho} Q_c^{*}(\os, \oa) = \expectedvalueover{\oa'\sim\rho} Q_c^{\pi'}(\os, \oa) = V_c^{\pi}(\os) \leq \beta
\end{equation*}

Consequently, $\pi_\text{greedy}(\cdot; Q^*)$ belongs to the $\argmax$ of \eqref{eq:pthm_marg_def2}, and in particular:
\begin{equation*}
     Q_r^*(\os, \oa) = r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \expectedvalueover{\oa'\sim\pi_\text{greedy}(\os', Q^*)} Q_r^*(\os', \oa')
\end{equation*}

The same reasoning can be made for $Q_c^*$ by replacing $\max$ operators by $\min$, and $\Pi_a$ by $\Pi_r$.
\end{proof}

%\subsection{Lemma \ref{lemma:concavity}}

%\begin{proof}. Let $s,s'\in\cS, a\in\cA$.
%We first prove those results for $V_r^*(s', \cdot)$

%\textbf{Non-decreasing}

%Consider $\beta_a^1 \leq \beta_a^2 \in \cB$.
%Any policy that satisfies the budget $\beta_a^1$ in $s'$ also satisfies $\beta_a^2$, so $\Pi_c(s', \beta_a^1) \subset \Pi_c(s', \beta_a^2)$. Hence, by taking the max over policies, $V_r^*(s', \beta_a^1) \leq V_r^*(s', \beta_a^2)$.
%Hence, $V_r^*(s', \cdot)$ is non-decreasing.

%\textbf{Concave}

%By contradiction: assume that $V_r^*(s', \cdot)$ is not concave, i.e. there exist $\beta^1 < \beta^2\in \cB$ and $p\in(0, 1)$ such that $\beta^3 = (1-p)\beta^1 + p\beta^2$ verifies: $V_r^*(s', \beta^3) < (1-p)V_r^*(s', \beta^1) + pV_r^*(s',\beta^2)$. By definition of $V^*$, there must be $\pi_1,\pi_2\in\Pi^*$ such that $V^*(s', \beta^1) = V^{\pi_1}(s', \beta^1)$ and $V^*(s', \beta^2) = V^{\pi_2}(s', \beta^2)$.

%Define $\pi = (1-p)(\pi_1(\cdot, \beta^1), \pi_1) + p(\pi_2(\cdot, \beta^2), \pi_2)$. By linearity of $V^\pi$ with respect to $\pi$, we have that $V_c^\pi(s', \beta^3) = (1-p)V_c^{\pi_1}(s', \beta^1) + pV_c^{\pi_2}(s', \beta^2) \leq (1-p)\beta^1 + p\beta^2 = \beta^3$ since $\pi_1, \pi_2\in\Pi^*(s')\subset\Pi_a(s')$, so $\pi$ respects the budget $\beta^3$. Moreover, we also have $V_r^\pi(s', \beta^3) = (1-p)V_r^{\pi_1}(s', \beta^1) + pV_r^{\pi_2}(s', \beta^2) > V_r^*(s', \beta^3)$, which contradicts the definition of $V_r^*$.

%Consequently, $V_r^*(s', \cdot)$ is non-decreasing and concave. By \eqref{eq:bellman_expectation_Q} we see that $Q_r^*(s,a,\cdot) = R(s,a) + \gamma\expectedvalueover{s'}V_r^*(s', \cdot)$  is too.

%\end{proof}

%\subsection{Lemma \ref{lemma:tau_concavity}}

%\subsection{Lemma \ref{lemma:pi_hull}}

%\td

%\begin{proof}
%If the estimates $q^c_0, q^c_1$ are accurate, then by construction and linearity of the expectation, the returned mixture policy has an expected total cost of $\expectedvalueover{a, \beta_a \sim\pi_\text{greedy}}Q_c(s, a, \beta_a) = \beta$ as desired in \eqref{eq:pi_greedy_constraint}. Because the $Q_r(s,a,\cdot)$ is concave and under its tangents, this mixture must have the largest $Q_r$ possible as required in \eqref{eq:pi_greedy_reward}. The special case of a tie $q_r^0 = q_r^1$ is considered, where we do minimise $Q_c$ as required in \eqref{eq:pi_greedy_cost}.
%\end{proof}

\subsection{Proposition \ref{prop:bftq_pi_hull}}
\begin{definition}
Let $A$ be a set, and $f$ a function defined on $A$. We define:

\begin{itemize}
    \item Convex hull of $A$: $\cC(A) = \{\sum_{i=1}^p \lambda_i a_i: a_i\in A, \lambda_i\in\Real^+, \sum_{i=1}^p \lambda_i = 1, p\in\Natural\}$
    \item Convex edges of $A$: $\cC^2(A) = \{\lambda a_1 + (1-\lambda)a_2: a_1, a_2\in A, \lambda\in[0, 1]\}$
    \item Dirac distributions of $A$: $\delta(A) = \{\delta(a-a_0): a_0\in A\}$ 
    \item Image of $A$ by $f$: $f(A) = \{f(a): a\in A\}$
\end{itemize}
\end{definition}

\begin{proof}
Let $\os=(s,\beta)\in\ocS$ and $Q\in(\Real^2)^{\ocS\ocA}$. We recall the definition of $\pi_\text{greedy}$:
\begin{subequations}
\label{eq:pi_greedy}
\begin{equation}
    \pi_\text{greedy}(\oa|\os; Q) \in \argmin_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa) \tag{\ref{eq:pi_greedy_cost}}
\end{equation}
\begin{align}
    \text{where }\quad\Pi_r^Q = &\argmax_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \tag{\ref{eq:pi_greedy_reward}}\\
    & \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \leq \beta \tag{\ref{eq:pi_greedy_constraint}}
\end{align}
\end{subequations}

Note that any policy in the $\argmin$ in \eqref{eq:pi_greedy_cost} is suitable to compute $\cT$.
We first reduce the set of candidate optimal policies.
Consider the problem described in \eqref{eq:pi_greedy_reward},\eqref{eq:pi_greedy_constraint}: it can be seen as a single-step CMDP problem with reward $R=Q_r$ and cost $C=Q_c$. By \citep[Theorem 4.4][]{BEUTLER1985236}, we know that the solutions are mixtures of two deterministic policies. Hence, we can replace $\cM(\cA)$ by $\cC^2(\delta(\ocA))$ in \eqref{eq:pi_greedy_reward}.

Moreover, remark that:
\begin{align*}
    \{\expectedvalueover{\oa\sim\rho} Q(\os,\oa): \rho\in \cC^2(\delta(\ocA))\} &= \{\expectedvalueover{\oa\sim\rho} Q(\os,\oa): \rho=(1-\lambda)\delta(\oa-\oa_1)+\lambda\delta(\oa-\oa_2), \oa_1,\oa_2\in\ocA, \lambda\in[0,1]\} \\
    &= \{(1-\lambda)Q(\os, \oa_1)+\lambda Q(\os, \oa_2), \oa_1,\oa_2\in\ocA, \lambda\in[0,1]\} \\
    &= \cC^2(Q(\os,\ocA))\}
\end{align*}

Hence, the problem \eqref{eq:pi_greedy_reward}, \eqref{eq:pi_greedy_constraint} has become:
\begin{equation*}
    \tilde{\Pi}^Q_r = \argmax_{(q_r, q_c)\in\cC^2(Q(\os, \ocA))} q_r \quad\text{ s.t. }\quad q_c \leq \beta 
\end{equation*}
and the solution space for $\pi_\text{greedy}$ is $q^*=\argmin_{q\in\tilde{\Pi}^Q_r} q_c$.

The original problem in the space of actions $\ocA$ is now expressed in the space of values $Q(\os, \ocA)$ (which is why we use $=$ instead of $\in$ before $\argmin$ here, see Remark \ref{rmk:independence}).

We further restrict the search space of $q^*$ following two observations:
\begin{enumerate}
    \item $q^*$ belongs to the \emph{undominated} points $\cC^2(Q(\os,\ocA)^-)$:
    \begin{equation*}
        Q(\os,\ocA)^- = \{q\in Q(\os,\ocA): q_c \leq q_c^{\pm} = \min_{q^+\in Q^+} q_c^+\}\text{ with }Q^+ = \argmax_{q\in Q(\os,\ocA)} q_r
    \end{equation*}
    Denote $q^*$ = $(1-\lambda) q^1 + \lambda q^2$, with $q^1, q^2\in Q(\os,\ocA)$. There are three possible cases:
    \begin{enumerate}
        \item $q^1, q^2 \not\in Q(\os,\ocA)^-$. Then $q_c^* = (1-\lambda) q^1_c + \lambda q^2_c > q_c^{\pm}$. But then $q_c^{\pm} < q_c^* \leq \beta$ so $q^{\pm}\in\tilde{\Pi}^Q_r$ with a strictly lower $q_c$ than $q^*$, which contradicts the $\argmin$.
        \item $q^1\in Q(\os,\ocA)^-, q^2 \not\in Q(\os,\ocA)^-$. But then consider the mixture $q^\top = (1-\lambda) q^1 + \lambda q^\pm$. Since $q_r^{\pm} \geq q_r^{2}$ and $q_r^{\pm} < q_r^{2}$, we also have $q^\top_r \geq q_r^*$ and $q^\top_c < q_c^*$, which also contradicts the $\argmin$.
        \item $q^1\in Q(\os,\ocA)^-$ is the only remaning possibility.
    \end{enumerate}
    \item $q^*$ belongs to the \emph{top frontier} $\cC^2(\cF_Q)$:
    \begin{equation*}
        \cF_Q = \{q\in Q(\os, \ocA)^-: \not\exists q'\in Q(\os, \ocA)^-: q_c=q_c'\text{ and }q_r<q_r'\}
    \end{equation*}
    Again, denote $q^*$ = $(1-\lambda) q^1 + \lambda q^2$, with $q^1, q^2\in Q(\os,\ocA)^-$.
    If either $q^1$ or $q^2$ do not belong to $\cF_Q$, they can be replaced in $q^\top = (1-\lambda) q^{1'} + \lambda q^{2'}$ by a $q^{i'}$ with equal $q_c$ and strictly higher $q_r$, which makes $q^\top$ a strictly better candidate than $q^*$ and contradicts the $\argmax$.
\end{enumerate}

TODO : $\cF_Q$ is "strictly increasing" (quel sens pour un set ??)

TODO : $\cF_Q$ is "concave" (quel sens pour un set ??)

We construct the optimal mixture as follow :

\textbf{Regular case} If it exists $\ox$ such that $Q_c(\ov{s},\ox) \geq \beta$ then $q^* = \lambda Q(\ov{s},\oa_0) + (1-\lambda) Q(\ov{s},\oa_1)$ with $(Q(\ov{s},\oa_0),Q(\ov{s},\oa_1)) \in \cF_Q\times \cF_Q$ .

The constraint value of Q for $\oa_0$ and $\oa_1$ must flank the budget: $Q_c(\ov{s},\oa_0) \leq \beta \leq Q_c(\ov{s},\oa_1)$. This condition  is verified by contradiction :

\begin{enumerate}
    \item if $Q_c(\ov{s},\oa_0) > \beta$ and $Q_c(\ov{s},\oa_1) > \beta$ then the constraint is violated.
    \item If $Q_c(\ov{s},\oa_0) < \beta$ and $Q_c(\ov{s},\oa_1) < \beta$ then $q^*$ is sub-optimal as it exists $\oa_\top$ such that $Q(\ov{s},\oa_\top) \in \cF_Q$ dominates $Q(\ov{s},\oa_0)$ and $Q(\ov{s},\oa_1)$ as $\cF_Q$ is "strictly increasing".
\end{enumerate}

The $\lambda$ value is designed such that the budget is respected. Then  $\lambda = \frac{\beta - Q_c(\ov{s},\oa_0) }{Q_c(\ov{s},\oa_1) - Q_c(\ov{s},\oa_0)}$ because the line $(Q(\ov{s},\oa_0),Q(\ov{s},\oa_1))$ is strictly increasing as  $Q_c(\ov{s},\oa_0) \leq \beta \leq Q_c(\ov{s},\oa_1)$ and $Q_r(\ov{s},\oa_0) \leq Q_r(\ov{s},\oa_1)$ as $\cF_Q$ is "strictly increasing". Then,  $\lambda$ is the unique candidate such that $q_c = \beta$.

The 2 candidates $\oa_0$ and $\oa_1$ are necessarily :

\begin{align}
\ov{a_0} = \argmin\limits_{\oz \in \ocA} (\beta - Q^c(\os,\ox))
 \text{ s.t. } Q^c(\os,\ox) \leq \beta
\text{ and } Q(\os,\ox) \in\cF_Q\\
\ov{a_1} = \argmin\limits_{\oz \in \ocA} (Q^c(\os,\ox) - \beta)
 \text{ s.t. } Q^c(\os,\ox) \geq \beta
\text{ and } Q(\os,\ox) \in\cF_Q
\end{align}

Indeed, let $(\ox,\oy)$ two points such that $Q^c(\os,\ox) \leq \beta \leq Q^c(\os,\oy)$ , then it exists $(\ox',\oy')$ such that $Q^c(\os,\ox) \leq Q^c(\os,\ox') \leq \beta \leq Q^c(\os,\oy) \leq Q^c(\os,\oy)$. As $\cF_Q$ is "concave", the segment $[Q(\os,\ox'),Q(\os,\oy')]$ is above the segment  $[Q(\os,\ox),Q(\os,\oy)]$ for any abscisse between $Q_c(\os,\ox)$ and $Q(\os,\oy)$ and in particular when the abscisse is $\beta$.

\textbf{Degenerate case} If it doesn't exists $\ox$ such that $Q_c(\ov{s},\ox) \geq \beta$ then  $q^* =  Q(\ov{s},\oa_0)$ with $\ov{a_0} = \argmax\limits_{\oz \in \ocA} Q^r(\os,\ox)$ such that $Q(\os,\ox) \in \cF_Q$

\end{proof}

%\begin{proof}
%First, a straightforward proof by induction shows that for all $k\in\Natural$, $Q_k$ computed at iteration $k$ of either Algorithm \ref{algo:bvi} or Algorithm \ref{algo:bftq} is concave non-decreasing with respect to $\beta_a$: the initialisation is trivial from $Q_0 = 0$, and the heredity stems from Lemma \ref{lemma:tau_concavity}.
%\end{proof}

% \subsection{Decomposition Lemma}

% \begin{lemma}
%     For any sequence real valued functions $f_1,\ldots,f_n$ and any real number $c$, we have:
%     \[
%         \begin{array}{lcl}
%             \underbrace{\max\limits_{\sum_i x_i \leq c}\sum_j f_j(x_j)}_{(a)} & \quad{}=\quad{} & \underbrace{\max\limits_{\sum_i c_i \leq c}\left(\sum_j\max\limits_{x\leq c_j} f_j(x)\right)}_{(b)}\\
%         \end{array}
%     \]
% \end{lemma}

% \begin{proof}
%     Let us first show that $(a)\leq(b)$.
%     By definition of the maximum on a set, for any $f_j$ and any $c_j$ we have:
%     $\max\limits_{x\leq c_j} f_j(x) \geq f_j(c_j)$.
%     Hence, by replacing these terms in $(b)$ we get:
%       \[
%     \begin{array}{lcl}
%         \max\limits_{\sum_i c_i \leq c} \sum_j f_j(c_j) & \quad{}\leq\quad{} & \max\limits_{\sum_i c_i\leq c}\left(\sum_j \max\limits_{x_j\leq c_j} f_j(x_j)\right)\\
%     \end{array}
%     \]
%     The left hand side of this inequality is just a rewriting of $(a)$ with different dummy variables names.

%     Let us show now that $(a) \geq (b)$.
%     Let $\hat{x}_1,\ldots,\hat{x}_n, \hat{c}_1, \ldots \hat{c}_n$ be a realisation (argmax) of $(b)$.
%     By definition of $(b)$'s feasible set, we have $\sum_i\hat{c}_i \leq c$ and for any $i$: $\hat{x}_i\leq \hat{c}_i$.
%     Because $\sum_i\hat{x}_i\leq \sum_i\hat{c}_i \leq c$, the tuple $(\hat{x}_1, \ldots \hat{x}_n)$ is also a feasible value for $(a)$. And, by definition of the maximum on a set: $(a) = \max\limits_{\sum_i x_i \leq c} \sum_j f_j(x_j) \geq \sum_j f_j(\hat{x}_j) = (b)$.
% \end{proof}

\section{Risk-Sensitive Exploration}

\input{source/risk-sensitive-explo-pseudo-code.tex}

\section{Scalable Implementation of BFTQ}
\label{sec:bftq-full}

\input{source/bftq-pseudo-code.tex}

\section{Experiments}

\subsection{The Lagrangian Relaxation Baseline}

As explained on \cref{fig:Lagrangian}, the optimal deterministic policy can be obtained by a line-search on the Lagrange multiplier values $\lambda$.

Then, according to \citet[Theorem 4.4]{BEUTLER1985236}, the optimal policy is a randomised mixture of two deterministic policies: the safest deterministic policy that violates the constraint $\pi_{\lambda-}$ and the riskier of the feasible ones $\pi_{\lambda+}$.

Fitted-Q (FTQ)~\citep{Ernst2005,Riedmiller2005} can be easily adapted for continuous states CMDP and BMDP through this methodology, but given the high variance it requires a lot of simulations to get a proper estimate of the calibration curve. Our purpose is to avoid this calibration phase.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.5\textwidth]{source/img/CalibrationExample}
    \caption{Calibration of a penalty multiplier according to the budget $\beta$. The optimal multiplier $\lambda^*_{\text{avg}}$ is the smallest one to satisfy the budget constraint on average. Safer policies can also be selected according to the largest deviation from this mean cost.}
    \label{fig:Lagrangian}
\end{figure}

\subsection{Environments Parameters}
\label{sec:env-parameters}

\paragraph{slot-filling} The system can either understand $(\mu=\mu_u)$ or misunderstand $(\mu=\mu_m)$ with a fixed probability called the sentence error rate $ser$. Then, the speech recognition score is simulated \citep{Khouzaimi2015}: $srs = (1+\exp(-x))^{-1}$ with $x\sim N(\mu, \sigma)$. It's the confidence score of the natural language understanding module about the last utterance. On the other hand, there are no recognition errors ($ser=0$ and $srs=1$) when the user provides information using the numeric pad.

\begin{table}[ht!]
    \centering
    \begin{tabularx}{1.0\textwidth}{lll}
        \toprule
        Parameter & Description & Value\tabularnewline
        \midrule
        - & Size of the environment & 7 x 6\tabularnewline
        - & \makecell[l]{Standard deviation of the Gaussian \\noise applied to actions} & (0.25,0.25)\tabularnewline
        H & Episode duration & 9\tabularnewline
        \bottomrule
    \end{tabularx}
    \caption{Parameters of \texttt{Corridors}}
\end{table}

\begin{table}[ht!]
    \centering
    \begin{tabularx}{1.0\textwidth}{lll}
        \toprule
        Parameter & Description & Value\tabularnewline
        \midrule
        ser & Sentence Error Rate & 0.6\tabularnewline
        $\mu_m$& Gaussian mean for misunderstanding & -0.25\tabularnewline
        $\mu_u$& Gaussian mean for understanding & 0.25\tabularnewline
        $\sigma$& Gaussian standard deviation & 0.6\tabularnewline
        $p$& Probability of hang-up & 0.25\tabularnewline
        H & Episode duration & 10\tabularnewline
        - & Number of slots & 3\tabularnewline
        \bottomrule
    \end{tabularx}
    \caption{Parameters of \texttt{Slot-Filling}}
\end{table}

\begin{table}[ht!]
    \centering
    \begin{tabularx}{1.0\textwidth}{lll}
        \toprule
        Parameter & Description & Value\tabularnewline
        \midrule
        $N_v$& Number of vehicles & 2 - 6\tabularnewline
        $\sigma_p$& Standard deviation of vehicles initial positions & 100 m\tabularnewline
        $\sigma_v$& Standard deviation of vehicles initial velocities & 3 m/s\tabularnewline
        H & Episode duration & 15 s\tabularnewline
        \bottomrule
    \end{tabularx}

    \caption{Parameters of \texttt{highway-env}}
\end{table}

\subsection{Algorithm parameters}
\label{sec:algorithms-parameters}

\begin{table}[H]
    \centering
    \begin{tabularx}{1.0\textwidth}{lll}
        \toprule
        Parameters & BFTQ(risk-sensitive) & BFTQ(risk-neutral)\tabularnewline
        \midrule
        architecture & 256x128x64 & 256x128x64\tabularnewline
        regularisation & 0.001 & 0.001\tabularnewline
        activation & relu & relu\tabularnewline
        size beta encoder & 3 & 3\tabularnewline
        initialisation & xavier & xavier\tabularnewline
        loss function & L2 & L2\tabularnewline
        optimizer & adam & adam\tabularnewline
        learning rate & 0.001 & 0.001\tabularnewline
        epoch (NN) & 1000 & 5000\tabularnewline
        normalize reward & true & true\tabularnewline
        epoch (FTQ) & 12 & 12\tabularnewline
        & 0:0.01:1 & -\tabularnewline
        beta for duplication & 0:0.1:1 & 0:0.1:1\tabularnewline
        & (1,1) & (1,1)\tabularnewline
        & 5000 & 5000\tabularnewline
        & 10 & 10\tabularnewline
        & 4 & 4\tabularnewline
        & 1000 & 1000\tabularnewline
        decay epsilon scheduling & 0.001 & 0.001\tabularnewline
        \bottomrule

    \end{tabularx}
    \caption{Algorithms parameters for \texttt{Corridors}}
\end{table}

\begin{table}[H]
    \centering
    \begin{tabularx}{1.0\textwidth}{lll}
        \toprule
        Parameters & BFTQ & FTQ\tabularnewline
        \midrule
        architecture & 256x128x64 & 128x64x32\tabularnewline
        regularisation & 0.0005 & 0.0005\tabularnewline
        activation & relu & relu\tabularnewline
        size beta encoder & 50 & -\tabularnewline
        initialisation & xavier & xavier\tabularnewline
        loss function & L2 & L2\tabularnewline
        optimizer & adam & adam\tabularnewline
        learning rate & 0.001 & 0.001\tabularnewline
        epoch (NN) & 5000 & 5000\tabularnewline
        normalize reward & true & true\tabularnewline
        epoch (FTQ) & 11 & 11\tabularnewline
        & 0:0.01:1 & -\tabularnewline
        beta for duplication & - & -\tabularnewline
        & (1,1) & (1,1)\tabularnewline
        & 5000 & 5000\tabularnewline
        & 10 & 10\tabularnewline
        & 6 & 6\tabularnewline
        & 1000 & 1000\tabularnewline
        decay epsilon scheduling & 0.001 & 0.001\tabularnewline
        \bottomrule
    \end{tabularx}
    \caption{Algorithms parameters for \texttt{Slot-Filling}}
\end{table}
%
\begin{table}[H]
    \centering
    \begin{tabularx}{1.0\textwidth}{lll}
        \toprule
        Parameters & BFTQ & FTQ\tabularnewline
        \midrule
        architecture & 256x128x64 & 128x64x32\tabularnewline
        regularisation & 0.0005 & 0\tabularnewline
        activation & relu & relu\tabularnewline
        size beta encoder & 50 & -\tabularnewline
        initialisation & xavier & xavier\tabularnewline
        loss function & L2 & L2\tabularnewline
        optimizer & adam & adam\tabularnewline
        learning rate & 0.001 & 0.01\tabularnewline
        epoch (NN) & 5000 & 400\tabularnewline
        normalize reward & true & true\tabularnewline
        epoch (FTQ) & 15 & 15\tabularnewline
        & 0:0.01:1 & -\tabularnewline
        beta for duplication & - & -\tabularnewline
        & (0.9, 0.9) & (0.9, 0.9)\tabularnewline
        & 10000 & 10000\tabularnewline
        & 10 & 10\tabularnewline
        & 10 & 10\tabularnewline
        & 150 & 150\tabularnewline
        decay epsilon scheduling & 0.0003 & 0.0003\tabularnewline
        \bottomrule
    \end{tabularx}
    \caption{Algorithms parameters for \texttt{Highway-Env}}
\end{table}

\subsection{Examples of BFTQ policies executions}
\label{sec:bftq-executions}

In \cref{table:dialogues}, we display two dialogues done with the same BFTQ policy. The policy is given two budgets to respect in expectation,$\beta=0$ and $\beta=0.5$. For budget 0, one can see that the system never uses the \texttt{ask\_num\_pad} action. Instead, it uses \texttt{ask\_oral} , an action subject to recognition errors. The system keeps asking for the same slot 2, because it has the lowest speech recognition score. It eventually summarizes the form to the user, but then reaches the maximum dialogue length and thus faces a dialogue failure. For budget 0.5, the system first asks in a safe way, with \texttt{ask\_oral}. It may want to \texttt{ask\_num\_pad} if one of the speech recognition score is low. Then, the system proceeds to a confirmation of the slot values. If it is incorrect, the system continues the dialogue using unsafe the \texttt{ask\_num\_pad} action to be certain of the slot values.

\input{source/dialogues.tex}

% \clearpage
% \bibliographystyle{named}
% \bibliography{budgeted_rl}
% \end{document}

### 2019-05-19 clean.py

@@ -1,0 +1,3 @@
+from subprocess import call
+
+call(['sh','compile.sh','clean'])

### 2019-05-19 main.py

@@ -1,0 +1,12 @@
+from subprocess import call
+
+call(['sh','compile.sh','compile','annexes'])
+
+call(['sh','compile.sh','compile','main'])
+call(['git','commit','-a','-m','\"update on pycharm\"'])
+call(['git','push'])
+
+# call(['sh','compile.sh','compile','analysis'])
+# call(['sh','compile.sh','compile','analysis_2'])
+
+

### 2019-05-21 main.tex
Updates from Overleaf

@@
 Odalric Maillard$^{1}$\and
-Olivier Pietquin$^{1,2,}$\thanks{now with Google Brain, Paris.}\\
+Olivier Pietquin$^{1,2,}$\thanks{now with Google Research - Brain Team, Paris, France.}\\
 \affiliations $^1$SequeL team, INRIA Lille -- Nord Europe, France\\
@@
 % Section 3: OK
-% Section 4.1: OK dans un premier temps, mais on va peut-être modifier.
-En fait, on remplace pi_greedy(Q) par pi_hull(Q), et on ne peut faire cela que lorsque Q est concave.
-Deux possibilité: soit on dit juste que Tau_greedy est remplacé par Tau_hull, qui admet également Q* comme point fixe (Tau_greedy(Q*) = Tau_hull(Q*) puisque Q* concave d'après le lemme).
-Soit on est intéressé par ce qui se passe le long de la trajectoire Q_k (par exemple si on suppose Tau_greedy contractant), et pour conserver cette propriété il faudrait alors montrer que Tau conserve la concavité et que donc on a bien Tau_hull = Tau_greedy tout le long de la trajectoire.
-Il faut aussi reprendre les notations des graphiques.
-% Section 4.2 et 4.3: KO. Pas encore retravaillé, mais ça ne devrait pas trop bougé
-% Section 5: OK pour environnements et results, KO pour discussion
-% Section 6: KO
-% Section 7: KO
+% Section 4: OK. Il faut aussi reprendre les notations des graphiques.
+% Section 5: OK
+% Section 6: OK
+% Section 7: OK
 
@@
 Reinforcement Learning (RL) is a general framework for decision-making under uncertainty.
-It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R, \gamma)$ with continuous state space $\cS$, discrete actions $\cA$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ and rewards $R\in\Real^{\cS \times \cA}$, where $\mathcal{M}(\mathcal{X})$ denotes the set of probability measures over a measurable space $\mathcal{X}$.
-Formally, we seek a policy $\pi\in\cM(A)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R(s_t, a_t)$.
+It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R_r, \gamma)$ with continuous state space $\cS$, discrete actions $\cA$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ and rewards $R_r\in\Real^{\cS \times \cA}$, where $\mathcal{M}(\mathcal{X})$ denotes the set of probability measures over a measurable space $\mathcal{X}$.
+Formally, we seek a policy $\pi\in\cM(A)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$.
 
@@
 
-Reinforcement Learning also assumes that the performance can be described solely by a reward function $R$.
+Reinforcement Learning also assumes that the performance can be described solely by a reward function $R_r$.
 Conversely, real problems typically involve many aspects, some of which can be contradictory \citep{ChunmingLiu2014}.
@@
 In this multi-objective formulation, task completion and safety are considered separately.
-We equip the MDP with a cost signal $C \in \Real^{\cS\times \cA}$ and a cost budget $\beta\in\Real$.
-Similarly to $G_r$, we define the return of costs $G_c^\pi = \sum_{t=0}^\infty \gamma^t C(s_t, a_t)$ and the new cost-constrained objective:
+We equip the MDP with a cost signal $R_c \in \Real^{\cS\times \cA}$ and a cost budget $\beta\in\Real$.
+Similarly to $G_r$, we define the return of costs $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ and the new cost-constrained objective:
 \begin{equation}
@@
 To address this concern, the \emph{Budgeted Markov Decision Process} (BMDP) was introduced in \citep{Boutilier_Lu:uai16} as an extension of CMDPs to enable the online control over the budget $\beta$ within an interval $\cB \subset \Real$ of admissible budgets.
-Instead of fixing the budget prior to training, the objective is now to find a generic optimal policy $\pi^*$ takes $\beta$ as input so as to solve the corresponding CMDP \eqref{eq:cmdp} for all $\beta\in\cB$.
+Instead of fixing the budget prior to training, the objective is now to find a generic optimal policy $\pi^*$ that takes $\beta$ as input so as to solve the corresponding CMDP \eqref{eq:cmdp} for all $\beta\in\cB$.
 This gives the system designer the ability to move the optimal policy $\pi^*$ in real-time along the Pareto-optimal curve of the different reward-cost trade-offs.
@@
 We then propose a novel Budgeted Bellman Optimality Operator and prove the optimal value function to be a fixed point of this operator.
-Second, we use this operator in \BFTQ, a Reinforcement Learning algorithm for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure.
+Second, we use this operator in a Reinforcement Learning algorithm for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure.
 Third, we scale this algorithms to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex computing, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
 Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.
+
+The proofs of our main results are provided in Supplementary Material \ref{sec:proofs}.
 
@@
 \begin{equation*}
-    \ov{P}\left(\os' \condbar \os, \oa\right) = \ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) = P(s'|s, a)\delta(\beta' - \beta_a)
+    \ov{P}\left(\os' \condbar \os, \oa\right) = \ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) \eqdef P(s'|s, a)\delta(\beta' - \beta_a),
 \end{equation*}
@@
 
-We stack the rewards and cost functions in a single \emph{vectorial} signal $\ov{R} \in (\Real^2)^{{\ocS \times \ocA}}$.
+We stack the rewards and cost functions in a single \emph{vectorial} signal $R \in (\Real^2)^{{\ocS \times \ocA}}$.
 Given an augmented transition $(\os, \oa) =((s,\beta), (a, \beta_a))$, we define:
 \begin{equation}
-     \ov{R}(\os, \oa) =  \begin{bmatrix}
-     R(s, a)\\
-     C(s, a)
-     \end{bmatrix}\in\Real^2
+     R(\os, \oa) \eqdef  \begin{bmatrix}
+     R_r(s, a)\\
+     R_c(s, a)
+     \end{bmatrix}\in\Real^2.
 \end{equation}
@@
 As earlier, the return $G^\pi = (G_r^\pi, G_c^\pi)$ of a budgeted policy $\pi\in\Pi$ refers to:
-% \begin{equation*}
-    $G^\pi = \sum_{t=0}^\infty \gamma^t \ov{R}(\os_t, \oa_t)$.
-% \end{equation*}
+    $G^\pi \eqdef \sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t)$.
 
 We can introduce the value function $V^\pi$ and action-value function $Q^\pi$ of a budgeted policy $\pi\in\Pi$ as:
-% \begin{align}
-% \label{eq:value-function}
-%     V^\pi(\os) &= \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \\
-%     Q^\pi(\os, \oa) &= \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right]
-% \end{align}
 \begin{equation}
     \label{eq:value-function}
-V^\pi(\os) = \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa) = \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right]
+V^\pi(\os) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right].
 \end{equation}
@@
 \begin{align}
-    V^\pi(\os) &= \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa)\\
-    Q^\pi(\os, \oa) &= \ov{R}(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') \label{eq:bellman_expectation_Q}
+    V^\pi(\os) &= \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa),\\
+    Q^\pi(\os, \oa) &= R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') \label{eq:bellman_expectation_Q}.
 \end{align}
@@
 We now come to the definition of budgeted optimality.
-We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the rewards $R$, (iii)~in case of tie, minimise the costs $C$.
+We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the rewards $R_r$, (iii)~in case of tie, minimise the costs $R_c$.
 To that end, we define the following sets for all $\os\in\ocS$:
 \begin{itemize}
-    \item Admissible policies: $\Pi_a(\os) = \{\pi\in\Pi: V_c^\pi(\os) \leq \beta\}$ where $\os = (s, \beta)$.
-    \item Best admissible policies: $\Pi_r(\os) = \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)$
-    \item Optimal policies: $\Pi^*(\os) = \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)$
+    \item Admissible policies: $\Pi_a(\os) \eqdef \{\pi\in\Pi: V_c^\pi(\os) \leq \beta\}$ where $\os = (s, \beta)$.
+    \item Best admissible policies: $\Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)$.
+    \item Optimal policies: $\Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)$.
 \end{itemize}
@@
 \begin{equation*}
-    V^*(\os) = V^{\pi^*}(\os), \qquad Q^*(\os, \oa) =  Q^{\pi^*}(\os, \oa),
+    V^*(\os) \eqdef V^{\pi^*}(\os), \qquad Q^*(\os, \oa) \eqdef  Q^{\pi^*}(\os, \oa),
 \end{equation*}
@@
 \label{eq:bellman-optimality}
-    Q^{*}(\os, \oa) = \cT Q^{*}(\os, \oa) \eqdef \ov{R}(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'})
+    Q^{*}(\os, \oa) = \cT Q^{*}(\os, \oa) \eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'}),
 \end{equation}
@@
 \begin{equation}
-    \pi_\text{greedy}(\oa|\os; Q) \in \argmin_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa) \label{eq:pi_greedy_cost}
+    \pi_\text{greedy}(\oa|\os; Q) \in \argmin_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa), \label{eq:pi_greedy_cost}
 \end{equation}
 \begin{align}
-    \text{where }\quad\Pi_r^Q = &\argmax_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \label{eq:pi_greedy_reward}\\
-    & \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \leq \beta \label{eq:pi_greedy_constraint}
+    \text{where }\quad\Pi_r^Q \eqdef &\argmax_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \label{eq:pi_greedy_reward}\\
+    & \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \leq \beta. \label{eq:pi_greedy_constraint}
 \end{align}
@@
 Because of the budget dynamics $\beta' = \beta_a$, we can see in \eqref{eq:bellman_expectation_Q} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
-We will slightly abuse notations and denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
+We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
+\label{remark:notations}
 \end{remark}
@@
 
-When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(s_i,\beta_i,a_i,R_i,C_i,s_i')\}_{i\in [0,N]}$ collected by interaction with the environment.
-Applying $\cT$ in \eqref{eq:bellman-optimality} requires computing an expectation $\expectedvalueover{\os'\sim \ov{P}}$ over next states $\os'$, which we can replace by an empirical average $\sum_{s'\in\cD}$.
+When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i=(s_i,\beta_i),\oa_i=(a_i,\beta_{a_i}),R_i=({R_r}_i,{R_c}_i),\os_i'=(s_i',\beta_{a_i})\}_{i\in [0,N]}$ collected by interaction with the environment.
+Following \cref{remark:notations}, we rewrite this batch as  $\cD=\{s_i,a_i,\beta_{a_i},r_i,c_i,s_i'\}_{i\in [0,N]}$.
+Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$.
+We instead use $\hat{\cT}$ in which this expectation is replaced by a sample average:
+\begin{equation*}
+    \hat{\cT} Q(\os, \oa, R, \os') \eqdef R + \gamma \sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q) Q(\ov{s'}, \ov{a'}).
+\end{equation*}
 We introduce in Algorithm~\ref{algo:bftq} the \emph{Budgeted-Fitted-Q} (BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs.
-Because we work on samples of a continuous state space $\cS$, we need to employ function-approximation in order to generalise to nearby states.
-Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $L(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, \ov{R}, \os')||_2^2$.
+Because we work with  continuous state space $\cS$ and budget space $\cB$, we need to employ function-approximation in order to generalise to nearby states and budgets.
+Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, R, \os')||_2^2$.
 Any model can be used, such as linear models, regression trees, or neural networks.
 
-\begin{minipage}[t]{0.5\textwidth}
+\begin{minipage}[t]{0.47\textwidth}
 \vspace{0pt}  
@@
 \DontPrintSemicolon
-\KwData{$P, R, C$}
+\KwData{$P, R_r, R_c$}
 \KwResult{$Q^*$}
@@
 \end{minipage}%
-\begin{minipage}[t]{0.5\textwidth}
+\hfill
+\begin{minipage}[t]{0.47\textwidth}
 \vspace{0pt}
@@
 \KwResult{$Q^*$}
-$Q_{0} \leftarrow 0$\;
+$Q_{\theta_0} \leftarrow 0$\;
 \Repeat{convergence}{
-    $Q_{k+1} \leftarrow \argmin_\theta \cL(Q_\theta, \cT Q_k; \cD)$\;
+    $\theta_{k+1} \leftarrow \argmin_\theta \cL(Q_\theta, \hat{\cT} Q_{\theta_{k}}; \cD)$\;
 }
@@
 As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested non-linear constrained optimisation program \eqref{eq:pi_greedy}.
-However, it can be solved efficiently by exploiting the \emph{structure} of the optimal value function $V^*$.
-% In the following, we say that a vector-valued function is concave and non-decreasing with respect to a variable if all of its components are.
-
-%\begin{lemma}[Concavity and monotony of $Q^*$]
-%\label{lemma:concavity}
-%$Q^*$ is non-decreasing concave with respect to $\beta_a$. 
-%\end{lemma}
-%\begin{lemma}[Conservation of concavity and monotony by $\cT$]
-%\label{lemma:tau_concavity}
-%For any $Q\in(\Real^2)^{\ocS\ocA}$ non-decreasing concave with respect to $\beta_a$, $\cT Q$ is also non-decreasing concave with respect to $\beta_a$.
-%\end{lemma}
-
-%These two Lemmas can be used to derive an efficient computation method for $\pi_\text{greedy}$.
-
-%\begin{lemma}[Hull policy equality]
-%\label{lemma:pi_hull}
-%For any $Q\in(\Real^2)^{\ocS\ocA}$ non-decreasing concave with respect to $\beta_a$, $\pi_\text{greedy}$ in \eqref{eq:pi_greedy} can be computed by running Algorithm~\ref{algo:pi_hull}.
-%\begin{equation*}
-%    \pi_\text{greedy}(\oa|\os; Q) = \pi_\text{hull}(\oa|\os; Q)
-%\end{equation*}
-%\end{lemma}
+However, it can be solved efficiently by exploiting the \emph{structure} of the set of solutions with respect to $\beta$, which is concave and increasing.
 
@@
 \KwData{$s$, $Q$}
-$\cH \leftarrow \texttt{convex\_hull}(Q(s,\cA, \cB))$\;
-denote $q^i = Q(s,a^i,\beta_a^i)\in\cH$ its $i^{\text{th}}$ point in clockwise order\;
-find two successive points $((q_c^0, q_r^0), (q_c^1, q_r^1))$ of $\cH$ such that $q_c^0 \leq \beta < q_c^1$\;
-\lIf{$q_r^0 = q_r^1$}{
-\Return $\delta(a^0, \beta_a^0)$
-}\Else{
-$p \leftarrow (\beta - q_c^0) / (q_c^1 - q_c^0)$\;
-\Return the mixture $(1-p)\delta(a^0, \beta_a^0) + p\delta(a^1, \beta_a^1)$\;
-}
+$Q^+\leftarrow \{Q_c > \min \{Q_c(s,\oa) \text{ s.t. }\oa\in\argmax_{\oa} Q_r(s,\oa)\} \}$\tcp*[f]{dominated points}\;
+$\cF \leftarrow \text{top frontier of }\texttt{convex\_hull}(Q(s,\ocA) \setminus Q^+)$\tcp*[f]{candidate mixtures}\;
+$\cF_Q \leftarrow \cF\cap Q(s,\ocA)$\;
+\For{points $q = Q(s,\oa)\in\cF_Q$ in clockwise order}{
+\uIf{find two successive points $((q_c^1, q_r^1), (q_c^2, q_r^2))$ of $\cF_Q$ such that $q_c^1 \leq \beta < q_c^2$}{
+$p \leftarrow (\beta - q_c^1) / (q_c^2 - q_c^1)$\;
+\Return the mixture $(1-p)\delta(\oa-\oa^1) + p\delta(\oa-\oa^2)$\;
+}}
+\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(s,\oa))$\tcp*[f]{Budget $\beta$ always respected}}
 \caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
@@
 \end{algorithm}
+
+The computation of $\pi_\text{hull}$ in \cref{algo:pi_hull} is illustrated in \cref{fig:hull}.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{source/img/qrqc}
-    \caption{Computation of $\pi_\text{hull}$ for a concave non-decreasing Q-function $Q$.
-The dotted tangent on the right defines the greedy policy when the budget $\beta$ lies between $Q_c(s,a^0,\beta_a^0)$ and $Q_c(s,a^1,\beta_a^1)$, as mixtures of the two points $(0, 1)$ represented as green crosses}
+    \includegraphics[width=\linewidth]{source/img/pi.pdf}
+    \caption{Representation of $\pi_\text{hull}$.
+When the budget lies between $Q(\os,\oa_1)$ and $Q(\os,\oa_2)$, two points of the top frontier of the convex hull, then the policy is a mixture of these two points.}
     \label{fig:hull}
@@
     \input{source/architecture.tex}
-    \caption{Neural Network for $Q$-functions approximation when the state dimension is 2 and there are 2 actions.}
+    \caption{Neural Network for $Q$-functions approximation when $\cS=\Real^2$ and $|\cA| = 2$.}
     \label{fig:architecture}
@@
 We approximate $Q = (Q_r, Q_c)$ using one single neural network.
-Thus, the ywo components are jointly optimised which accelerates convergence and fosters learning of useful shared representations.
+Thus, the two components are jointly optimised which accelerates convergence and fosters learning of useful shared representations.
 Moreover, as in \citep{Mnih2015} we are dealing with a finite (categorical) action space $\cA$, instead of including the action in the input we add the output of the $Q$-function for each action to the last layer.
@@
 This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA,\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass.
-By using multiprocessing in the computations of $\pi_\text{hull}$, we obtain a linear speedup.% and reduce the time-complexity to $O(|\mathcal{D}|/(W(1-\gamma)) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$.
+By using multiprocessing in the computations of $\pi_\text{hull}$, we enjoy a linear speedup.% and reduce the time-complexity to $O(|\mathcal{D}|/(W(1-\gamma)) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$.
 The full description of our scalable implementation of \BFTQ is recalled in Algorithm~\ref{algo:bftq_full} in Supplementary Material~\ref{sec:bftq-full}.
@@
 Their behaviours are randomised, which introduces some uncertainty with respect to their possible future trajectories.
-The task consists in driving as fast as possible, which is modelled by a reward proportional to the velocity: $R(s_t, a_t) \propto v_t$.
+The task consists in driving as fast as possible, which is modelled by a reward proportional to the velocity: $R_r(s_t, a_t) \propto v_t$.
 This motivates the agent to try and overtake its preceding vehicles by driving fast on the opposite lane.
-This optimal but overly aggressive behaviour can be tempered through a cost function that embodies a safety objective: $C(s_t, a_t)$ is set to $1/H$ whenever the ego-vehicle is driving on the opposite lane, where $H$ is the episode horizon.
+This optimal but overly aggressive behaviour can be tempered through a cost function that embodies a safety objective: $R_c(s_t, a_t)$ is set to $1/H$ whenever the ego-vehicle is driving on the opposite lane, where $H$ is the episode horizon.
 Thus, the constrained signal $G_c^\pi$ is the maximum proportion of time that the agent is allowed to drive on the wrong side of the road.
@@
     \centering
-    \includegraphics[width=0.25\textwidth]{source/img/risk-neutral.png}
-    \includegraphics[width=0.25\textwidth]{source/img/risk-sensitive.png}
-    \includegraphics[page=1, width=0.4\textwidth]{source/img/corridors}
+    \includegraphics[width=0.23\textwidth]{source/img/risk-neutral.png}
+    \includegraphics[width=0.23\textwidth]{source/img/risk-sensitive.png}
+    \includegraphics[page=1, width=0.45\textwidth]{source/img/corridors}
     \caption{Trajectories (left) and performances (right) of two exploration strategies in the \text{corridors} environment. }
@@
     \begin{center}
-    \includegraphics[width=0.45\linewidth]{source/img/slot-filling}
-    \includegraphics[width=0.45\linewidth]{source/img/highway}
+    \includegraphics[width=0.49\linewidth]{source/img/slot-filling}
+    \includegraphics[width=0.49\linewidth]{source/img/highway}
     \caption{Performance comparison of \FTQl and BFTQ on \text{slot-filling} (left) and \text{highway-env}(right) }
@@
 In both cases, BFTQ performs almost as well as \FTQl despite only requiring a single model.
-All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can notice in the bottom-left corner an underestimation of $Q^c$, since $\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
+All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can notice in the bottom-left corner an underestimation of $Q_c$, since $\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
 However, these estimates still provide a reasonable prior approximation of the expected cost of the policy, which cannot be obtained with \FTQl.
@@
 Another related area is multi-objective reinforcement learning (MORL)(Van Moffaert & Nowe´, 2014; Roijers et al., 2013).
-Generally, research in MORL has largely focused on approximating the Pareto frontier that trades-off competing objectives (Van Moffaert & Nowe´, 2014; Roijers et al., 2013).The underlying approach to MORL frequently relies on linear or non-linear scalarization of rewards to heuristicallyturns the problem into a standard RL problem.
+Generally, research in MORL has largely focused on approximating the Pareto frontier that trades-off competing objectives (Van Moffaert & Nowe´, 2014; Roijers et al., 2013).The underlying approach to MORL frequently relies on linear or non-linear scalarization of rewards to heuristically turns the problem into a standard RL problem.
 Our proposed approach represents another systematic paradigm to solve MORL, whether in batch or online settings.}
@@
 Algorithms have been proposed for CMDPs, which are less flexible sub-problems of the more general BMDP.
-When the environment parameters ($P$, $R$, $C$) are known but not tractable, solutions relying on function approximation~\citep{Undurti} or approximate linear programming~\citep{Poupart2015} have been proposed.
+When the environment parameters ($P$, $R_r$, $R_c$) are known but not tractable, solutions relying on function approximation~\citep{Undurti} or approximate linear programming~\citep{Poupart2015} have been proposed.
 For unknown environments, batch algorithms exist in the tabular setting \citep{Thomas2015, Petrik2016, Laroche2017}.
@@
 For instance, they enumerate and sort the next states $s'\in\cS$ by their expected value-by-cost, which could not be performed in a continuous state space $\cS$.
-Moreover, they rely on the knowledge of the model ($P$, $R$, $C$), and do not address the question of learning from interaction data.
+Moreover, they rely on the knowledge of the model ($P$, $R_r$, $R_c$), and do not address the question of learning from interaction data.
 
@@
 %    Thanks to Guillaume Gautier, Fabrice Clerot, Xuedong Shang
-
+\clearpage
 \bibliographystyle{named}

### 2019-05-21 supplementary.tex
Updates from Overleaf

@@
 
-\section{Proof of Main Results}
+\section{Proofs of Main Results}
+\label{sec:proofs}
 \subsection{Proposition \ref{prop:bellman-expectation}}
@@
 \begin{align*}
-    Q^\pi(\os, \oa) &= \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \ov{R}(\os_t, \oa_t)\condbar \ov{s_0} = \os, \ov{a_0} = \oa\right] \\
-    &= \ov{R}(\os, \oa) + \sum_{\os'\in\ocS}\probability{\os_1 = \os' \condbar\ov{s_0} = \os, \ov{a_0} = \oa}\expectedvalue\left[\sum_{t=1}^\infty \gamma^t \ov{R}(\os_t, \oa_t)\condbar \ov{s_1} = \os'\right] \\
-    &= \ov{R}(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar\os, \oa\right) \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \ov{R}(\os_t, \oa_t) \condbar \ov{s_0} = \os'\right] \\
-    &=  \ov{R}(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar\os, \oa\right) V^\pi(\os')
+    Q^\pi(\os, \oa) &= \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t)\condbar \ov{s_0} = \os, \ov{a_0} = \oa\right] \\
+    &= R(\os, \oa) + \sum_{\os'\in\ocS}\probability{\os_1 = \os' \condbar\ov{s_0} = \os, \ov{a_0} = \oa}\expectedvalue\left[\sum_{t=1}^\infty \gamma^t R(\os_t, \oa_t)\condbar \ov{s_1} = \os'\right] \\
+    &= R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar\os, \oa\right) \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t) \condbar \ov{s_0} = \os'\right] \\
+    &=  R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar\os, \oa\right) V^\pi(\os')
 \end{align*}
@@
 
-%Consequently, $V_r^*(s', \cdot)$ is non-decreasing and concave. By \eqref{eq:bellman_expectation_Q} we see that $Q_r^*(s,a,\cdot) = R(s,a) + \gamma\expectedvalueover{s'}V_r^*(s', \cdot)$  is too.
+%Consequently, $V_r^*(s', \cdot)$ is non-decreasing and concave. By \eqref{eq:bellman_expectation_Q} we see that $Q_r^*(s,a,\cdot) = R_r(s,a) + \gamma\expectedvalueover{s'}V_r^*(s', \cdot)$  is too.
 
@@
 \begin{subequations}
-\label{eq:pi_greedy}
 \begin{equation}
@@
 We first reduce the set of candidate optimal policies.
-Consider the problem described in \eqref{eq:pi_greedy_reward},\eqref{eq:pi_greedy_constraint}: it can be seen as a single-step CMDP problem with reward $R=Q_r$ and cost $C=Q_c$.
+Consider the problem described in \eqref{eq:pi_greedy_reward},\eqref{eq:pi_greedy_constraint}: it can be seen as a single-step CMDP problem with reward $R_r=Q_r$ and cost $R_c=Q_c$.
 By \citep[Theorem 4.4][]{BEUTLER1985236}, we know that the solutions are mixtures of two deterministic policies.
@@
 \end{equation*}
-and the solution space for $\pi_\text{greedy}$ is $q^*=\argmin_{q\in\tilde{\Pi}^Q_r} q_c$.
+and the solution of $\pi_\text{greedy}$ is $q^*=\argmin_{q\in\tilde{\Pi}^Q_r} q_c$.
 
@@
 \begin{enumerate}
-    \item $q^*$ belongs to the \emph{undominated} points $\cC^2(Q(\os,\ocA)^-)$:
-    \begin{equation*}
-        Q(\os,\ocA)^- = \{q\in Q(\os,\ocA): q_c \leq q_c^{\pm} = \min_{q^+\in Q^+} q_c^+\}\text{ with }Q^+ = \argmax_{q\in Q(\os,\ocA)} q_r
-    \end{equation*}
+    \item $q^*$ belongs to the \emph{undominated} points $\cC^2(Q^-)$:
+    \begin{align}
+        \label{eq:q_minus_undominated}
+        Q^+ &= \{(q_c, q_r): q_c > q_c^{\pm} = \min_{q^+} q_c^+\text{ s.t. }q^+\in\argmax_{q\in Q(\os,\ocA)} q_r\}\\
+        Q^- &= Q(\os,\ocA) \setminus Q^+
+    \end{align}
     Denote $q^*$ = $(1-\lambda) q^1 + \lambda q^2$, with $q^1, q^2\in Q(\os,\ocA)$. There are three possible cases:
     \begin{enumerate}
-        \item $q^1, q^2 \not\in Q(\os,\ocA)^-$.
+        \item $q^1, q^2 \not\in Q^-$.
 Then $q_c^* = (1-\lambda) q^1_c + \lambda q^2_c > q_c^{\pm}$.
 But then $q_c^{\pm} < q_c^* \leq \beta$ so $q^{\pm}\in\tilde{\Pi}^Q_r$ with a strictly lower $q_c$ than $q^*$, which contradicts the $\argmin$.
-        \item $q^1\in Q(\os,\ocA)^-, q^2 \not\in Q(\os,\ocA)^-$.
+        \item $q^1\in Q^-, q^2 \not\in Q^-$.
 But then consider the mixture $q^\top = (1-\lambda) q^1 + \lambda q^\pm$.
 Since $q_r^{\pm} \geq q_r^{2}$ and $q_r^{\pm} < q_r^{2}$, we also have $q^\top_r \geq q_r^*$ and $q^\top_c < q_c^*$, which also contradicts the $\argmin$.
-        \item $q^1\in Q(\os,\ocA)^-$ is the only remaning possibility.
+        \item $q^1,q^2\in Q^-$ is the only remaining possibility.
     \end{enumerate}
-    \item $q^*$ belongs to the \emph{top frontier} $\cC^2(\cF_Q)$:
+    \item $q^*$ belongs to the \emph{top frontier} $\cF$:
     \begin{equation*}
-        \cF_Q = \{q\in Q(\os, \ocA)^-: \not\exists q'\in Q(\os, \ocA)^-: q_c=q_c'\text{ and }q_r<q_r'\}
+        \cF_Q = \{q\in \cC^2(Q^-): \not\exists q'\in \cC^2(Q^-): q_c=q_c'\text{ and }q_r<q_r'\}
     \end{equation*}
-    Again, denote $q^*$ = $(1-\lambda) q^1 + \lambda q^2$, with $q^1, q^2\in Q(\os,\ocA)^-$.
-    If either $q^1$ or $q^2$ do not belong to $\cF_Q$, they can be replaced in $q^\top = (1-\lambda) q^{1'} + \lambda q^{2'}$ by a $q^{i'}$ with equal $q_c$ and strictly higher $q_r$, which makes $q^\top$ a strictly better candidate than $q^*$ and contradicts the $\argmax$.
+    Trivially, otherwise q' would be a better candidate than $q^*$.
 \end{enumerate}
 
-TODO : $\cF_Q$ is "strictly increasing" (quel sens pour un set ??)
-
-TODO : $\cF_Q$ is "concave" (quel sens pour un set ??)
-
-We construct the optimal mixture as follow :
-
-\textbf{Regular case} If it exists $\ox$ such that $Q_c(\ov{s},\ox) \geq \beta$ then $q^* = \lambda Q(\ov{s},\oa_0) + (1-\lambda) Q(\ov{s},\oa_1)$ with $(Q(\ov{s},\oa_0),Q(\ov{s},\oa_1)) \in \cF_Q\times \cF_Q$ .
-
-The constraint value of Q for $\oa_0$ and $\oa_1$ must flank the budget: $Q_c(\ov{s},\oa_0) \leq \beta \leq Q_c(\ov{s},\oa_1)$. This condition  is verified by contradiction :
-
+Let us characterise this frontier $\cF$. It is both:
 \begin{enumerate}
-    \item if $Q_c(\ov{s},\oa_0) > \beta$ and $Q_c(\ov{s},\oa_1) > \beta$ then the constraint is violated.
-    \item If $Q_c(\ov{s},\oa_0) < \beta$ and $Q_c(\ov{s},\oa_1) < \beta$ then $q^*$ is sub-optimal as it exists $\oa_\top$ such that $Q(\ov{s},\oa_\top) \in \cF_Q$ dominates $Q(\ov{s},\oa_0)$ and $Q(\ov{s},\oa_1)$ as $\cF_Q$ is "strictly increasing".
+    \item the \emph{graph of a non-decreasing function}: $\forall q^1, q^2\in\cF$ such that $q_c^1\leq q_c^2$ then $q_r^1\leq q_r^2$.\\
+    By contradiction, if we had $q_r^1 > q_r^2$, we could define $q^\top = (1-\lambda)q^1 + \lambda q^\pm$ where $q^\pm$ is the dominant point as defined in \eqref{eq:q_minus_undominated}.
+By choosing $\lambda=(q^2_c-q^1_c)/(q^\pm_c-q^1_c)$ such that $q^\top_c = q_c^2$, then since $q_r^\pm \geq q_r^1 > q_r 2$ we also have $q^\top_r > q_r^2$ which contradicts $q^2\in\cF$.
+    \item the \emph{graph of a concave function}: $\forall q^1, q^2, q^3\in\cF$ such that $q_c^1\leq q_c^2 \leq q_c^3$ with $\lambda$ such that $q^2_c = (1-\lambda)q^1_c + \lambda q^3_c$, then $q_r^2 \geq (1-\lambda)q_r^1 + \lambda q_r^3$.\\
+    Trivially, otherwise the point $q^\top = (1-\lambda)q^1 + \lambda q^3$ would verify $q^\top_c=q^2_c$ and $q^\top_r > q^2_r$, which would contradict $q^2 \in\cF$.
 \end{enumerate}
 
-The $\lambda$ value is designed such that the budget is respected.
-Then  $\lambda = \frac{\beta - Q_c(\ov{s},\oa_0) }{Q_c(\ov{s},\oa_1) - Q_c(\ov{s},\oa_0)}$ because the line $(Q(\ov{s},\oa_0),Q(\ov{s},\oa_1))$ is strictly increasing as  $Q_c(\ov{s},\oa_0) \leq \beta \leq Q_c(\ov{s},\oa_1)$ and $Q_r(\ov{s},\oa_0) \leq Q_r(\ov{s},\oa_1)$ as $\cF_Q$ is "strictly increasing".
-Then,  $\lambda$ is the unique candidate such that $q_c = \beta$.
+We denote $\cF_Q = \cF \cap Q$.
+Clearly, $q^*\in\cC^2(\cF_Q)$: let $q^1, q^2\in Q^-$ such that $q^* = (1-\lambda)q^1  + \lambda q^2$.
+First, $q^1, q^2\in Q^-\subset\cC^2(Q^-)$.
+Then, by contradiction, if there existed $q^{1'}$ or $q^{2'}$ with equal $q_c$ and strictly higher $q_r$, again we could build an admissible mixture $q^{\top}=(1-\lambda)q^{1'}  + \lambda q^{2'}$ strictly better than $q^*$.
 
-The 2 candidates $\oa_0$ and $\oa_1$ are necessarily :
+$q^*$ can be written as $q^* = (1-\lambda)q^1  + \lambda q^2$ with $q^1, q^2\in\cF_Q$ and, without loss of generality, $q^1_c \leq q^2_c$.
 
-\begin{align}
-\ov{a_0} = \argmin\limits_{\oz \in \ocA} (\beta - Q^c(\os,\ox))
- \text{ s.t. } Q^c(\os,\ox) \leq \beta
-\text{ and } Q(\os,\ox) \in\cF_Q\\
-\ov{a_1} = \argmin\limits_{\oz \in \ocA} (Q^c(\os,\ox) - \beta)
- \text{ s.t. } Q^c(\os,\ox) \geq \beta
-\text{ and } Q(\os,\ox) \in\cF_Q
-\end{align}
+\textbf{Regular case:} there exists $q^0\in\cF_Q$ such that $q^0_c \geq \beta$.
 
-Indeed, let $(\ox,\oy)$ two points such that $Q^c(\os,\ox) \leq \beta \leq Q^c(\os,\oy)$ , then it exists $(\ox',\oy')$ such that $Q^c(\os,\ox) \leq Q^c(\os,\ox') \leq \beta \leq Q^c(\os,\oy) \leq Q^c(\os,\oy)$.
-As $\cF_Q$ is "concave", the segment $[Q(\os,\ox'),Q(\os,\oy')]$ is above the segment  $[Q(\os,\ox),Q(\os,\oy)]$ for any abscisse between $Q_c(\os,\ox)$ and $Q(\os,\oy)$ and in particular when the abscisse is $\beta$.
+Then $q^1$ and $q^2$ must flank the budget: $q_c^1 \leq \beta \leq q_c^2$.
+Indeed, by contradiction, if $q_c^2 \geq q_c^1 > \beta$ then $q_c^* > \beta$ which contradicts $\Pi_r^Q$.
+Conversely, if $q_c^1 \leq q_c^2 < \beta$ then $q^* < \beta \leq q^0_c$, which would make $q^*$ a worse candidate than $q^\top=(1-\lambda)q^* + \lambda q^0$ when $\lambda$ is chosen such that $q_c^\top=\beta$, and contradict $\Pi_r^Q$ again.
 
-\textbf{Degenerate case} If it doesn't exists $\ox$ such that $Q_c(\ov{s},\ox) \geq \beta$ then  $q^* =  Q(\ov{s},\oa_0)$ with $\ov{a_0} = \argmax\limits_{\oz \in \ocA} Q^r(\os,\ox)$ such that $Q(\os,\ox) \in \cF_Q$
+Because $\cF$ is the graph of a non-decreasing function, $\lambda$ should be as high as possible, as long as the budget $q^*\leq\beta$ is respected.
+We reach the highest $q_r^*$ when $q^*_c=\beta$, that is: $\lambda=(\beta-q_c^1)/(q_c^2-q_c^1)$.
+
+It remains to show that $q^1$ and $q^2$ are two successive points in $\cF_Q$: $\not\exists q\in\cF_Q\setminus\{q^1, q^2\}: q^1_c \leq q_c \leq q^2_c$.
+Otherwise, as $\cF$ is the graph of a concave function, we would have $q_r \geq (1-\mu)q_r^1 + \mu q_r^2$.
+$q_r$ cannot be strictly greater than $(1-\mu)q_r^1 + \mu q_r^2$ which would contradict $q^*$, but it can still be equal, which means the tree points $q, q^1, q^2$ are aligned.
+In fact, every points aligned with $q^1$ and $q^2$ can also be used to construct mixtures resulting in $q^*$, but among these solutions we can still choose $q^1$ and $q^2$ as the two points in $\cF_Q$ closest to $q^*$.
+
+\textbf{Edge case:} $\forall q\in\cF_Q, q_c < \beta$. Then  $q^* =  \argmax_{q\in\cF} q_r = q^\pm =  \argmax_{q\in Q^-} q_r$
 
@@
 
+We recall the Risk-Sensitive Exploration algorithm in \cref{algo:risk-sensitive-exploration}
+
 \input{source/risk-sensitive-explo-pseudo-code.tex}
@@
 \label{sec:bftq-full}
+
+We recall the scalable version of BFTQ in \cref{algo:bftq_full}
 
@@
 
-\paragraph{slot-filling} The system can either understand $(\mu=\mu_u)$ or misunderstand $(\mu=\mu_m)$ with a fixed probability called the sentence error rate $ser$.
+All environments parameters are displayed in \cref{tab:param-corridors},\cref{tab:param-slot-filling} and \cref{tab:param-highway-env}.
+
+\paragraph{Remark on the slot-filling environment} When receiving an utterance, the system can either understand it $(\mu=\mu_u)$ or misunderstand it $(\mu=\mu_m)$ with a fixed probability called the sentence error rate $ser$.
 Then, the speech recognition score is simulated \citep{Khouzaimi2015}: $srs = (1+\exp(-x))^{-1}$ with $x\sim N(\mu, \sigma)$.
 It's the confidence score of the natural language understanding module about the last utterance.
-On the other hand, there are no recognition errors ($ser=0$ and $srs=1$) when the user provides information using the numeric pad.
+Note that here are no recognition errors ($ser=0$ and $srs=1$) when the user provides information using the numeric pad.
 
@@
     \caption{Parameters of \texttt{Corridors}}
+    \label{tab:param-corridors}
 \end{table}
@@
     \caption{Parameters of \texttt{Slot-Filling}}
+    \label{tab:param-slot-filling}
 \end{table}
@@
     \caption{Parameters of \texttt{highway-env}}
+    \label{tab:param-highway-env}
 \end{table}
@@
 
-\begin{table}[H]
+All algorithm parameters are displayed in \cref{tab:param-algo-corridors},\cref{tab:param-algo-slot-filling} and \cref{tab:param-algo-highway-env}.
+
+\begin{table}[tp]
     \centering
@@
     \caption{Algorithms parameters for \texttt{Corridors}}
+    \label{tab:param-algo-corridors}
 \end{table}
 
-\begin{table}[H]
+\begin{table}[tp]
     \centering
@@
     \caption{Algorithms parameters for \texttt{Slot-Filling}}
+    \label{tab:param-algo-slot-filling}
 \end{table}
 %
-\begin{table}[H]
+\begin{table}[tp]
     \centering
@@
     \caption{Algorithms parameters for \texttt{Highway-Env}}
+    \label{tab:param-algo-highway-env}
 \end{table}

### 2019-05-21 main.tex
Updates from Overleaf

@@
 
-% @Nicolas/Edouard: faites ici la liste des sections qui sont pretes a etre relues.
-% Section 1: il manque les contributions
-% Section 2: OK. On pense éventuellement rajouter une hypothèse de contractance de Tau (pas sûr que ce soit le cas)
-% Section 3: OK
-% Section 4: OK. Il faut aussi reprendre les notations des graphiques.
-% Section 5: OK
-% Section 6: OK
-% Section 7: OK
-
 \section{Introduction}
@@
 When aggregating several objectives in a single scalar signal, as often in Multi-Objectives RL~\citep{Roijers2013ASO}, no control is given over their relative ratios, as high rewards can compensate high penalties.
-For instance, if a weighted sum is used to balance velocity $v$ and accidents $a$, then for any given choice of weights $\omega$ the optimality equation $\omega_v\expectedvalue[\sum\gamma^t v_t] + \omega_a\expectedvalue[\sum\gamma^t a_t] = G_r^* = \max_\pi G^\pi_r$ is the equation of a line in $(\expectedvalue[\sum\gamma^t v_t], \expectedvalue[\sum\gamma^t a_t])$, and the automotive company cannot control where its optimal policy $\pi^*$ lies on that line.
+For instance, if a weighted sum is used to balance velocity $v$ and crashes $c$, then for any given choice of weights $\omega$ the optimality equation $\omega_v\expectedvalue[\sum\gamma^t v_t] + \omega_a\expectedvalue[\sum\gamma^t c_t] = G_r^* = \max_\pi G^\pi_r$ is the equation of a line in $(\expectedvalue[\sum\gamma^t v_t], \expectedvalue[\sum\gamma^t c_t])$, and the automotive company cannot control where its optimal policy $\pi^*$ lies on that line.
 
@@
 Second, we use this operator in a Reinforcement Learning algorithm for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure.
-Third, we scale this algorithms to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex computing, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
+Third, we scale this algorithms to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
 Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.
@@
 \begin{align}
-    V^\pi(\os) &= \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa),\\
+    V^\pi(\os) &= \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa),\label{eq:bellman_expectation_V}\\
     Q^\pi(\os, \oa) &= R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') \label{eq:bellman_expectation_Q}.
@@
 We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the rewards $R_r$, (iii)~in case of tie, minimise the costs $R_c$.
-To that end, we define the following sets for all $\os\in\ocS$:
+To that end, we define for all $\os\in\ocS$:
 \begin{itemize}
-    \item Admissible policies: $\Pi_a(\os) \eqdef \{\pi\in\Pi: V_c^\pi(\os) \leq \beta\}$ where $\os = (s, \beta)$.
-    \item Best admissible policies: $\Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)$.
-    \item Optimal policies: $\Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)$.
+    \item Admissible policies:
+    \begin{equation}
+        \Pi_a(\os) \eqdef \{\pi\in\Pi: V_c^\pi(\os) \leq \beta\}\text{ where }\os = (s, \beta)
+    \end{equation}
+    \item Optimal value function for rewards $ V_r^*$ and candidate policies $\Pi_r$:
+    \begin{equation}
+        V_r^*(\os) \eqdef \max_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\qquad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
+    \end{equation}
+    \item Optimal value function for costs $V_c^*$ and optimal policies $\Pi^*$:
+    \begin{equation}
+        V_c^*(\os) \eqdef \min_{\pi\in\Pi_r(\os)}  V_c^\pi(\os) \qquad\qquad \Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
+    \end{equation}
 \end{itemize}
+We define the budgeted action-value function $Q^*$ similarly:
+\begin{equation}
+    Q_r^*(\os, \oa) \eqdef \max_{\pi\in\Pi_a(\os)}  Q_r^\pi(\os, \oa) \qquad\qquad Q_c^*(\os, \oa) \eqdef \min_{\pi\in\Pi_r(\os)}  Q_c^\pi(\os, \oa) 
+\end{equation}
+and denote $V^* = (V_r^*, V_c^*)$, $Q = (Q_r^*, Q_c^*)$.
 \end{definition}
-
-\begin{assumption}[Existence of optimal policy]
-There exists one budgeted policy $\pi^*$ that uniformly maximises $V_r^\pi$ under constraint on $V_c^\pi$ for all augmented states $\os$.
-\begin{align*}
-    \exists \pi^*\in\Pi: \forall \os\in \ocS, \pi^*\in\Pi^*(\os)
-\end{align*}
-\end{assumption}
-
-Let $\pi^*$ be one such optimal budgeted policy, we finally define the optimal budgeted value functions $V^*$, $Q^*$. For all $\os, \oa \in \ocS\times\ocA$:
-\begin{equation*}
-    V^*(\os) \eqdef V^{\pi^*}(\os), \qquad Q^*(\os, \oa) \eqdef  Q^{\pi^*}(\os, \oa),
-\end{equation*}
-
-\begin{remark}[Independence on the representative]
-\label{rmk:independence}
-Though they seemingly depend on the choice of $\pi^*$, the value functions $V^*$ and $Q^*$ are well-defined because all optimal policies in $\Pi^*$ share the same $V_r, Q_r$ as they belong to $\Pi_r$, and the same $V_c, Q_c$  as they belong to $\Pi^*$.
-\end{remark}
 
@@
 \label{thm:bellman-optimality}
-The optimal action-value of a budgeted policy verifies:
+The optimal budgeted action-value function $Q^*$ verifies:
 \begin{equation}
@@
 
-\begin{remark}[Appearance of the greedy policy]
-\label{rmk:greedy}
-In classical Reinforcement Learning, the greedy policy takes a simple form $\pi_\text{greedy}(s; Q^*) = \argmax_{a\in\cA} Q^*(s, a)$, and the term $\pi_\text{greedy}(a'|s';Q^*) Q^{*}(s', a')$ in \eqref{eq:bellman-optimality} conveniently simplifies to $\max_{a'\in \cA} Q^*(s', a')$.
+% \begin{remark}[Appearance of the greedy policy]
+% \label{rmk:greedy}
+% In classical Reinforcement Learning, the greedy policy takes a simple form $\pi_\text{greedy}(s; Q^*) = \argmax_{a\in\cA} Q^*(s, a)$, and the term $\pi_\text{greedy}(a'|s';Q^*) Q^{*}(s', a')$ in \eqref{eq:bellman-optimality} conveniently simplifies to $\max_{a'\in \cA} Q^*(s', a')$.
 Unfortunately, in a budgeted setting the greedy policy requires solving the nested constrained optimisation program \eqref{eq:pi_greedy} at each state and budget in order to apply this Budgeted Bellman Optimality operator.
-\end{remark}
+% \end{remark}
+
+\begin{proposition}[Optimality of the greedy policy]
+\label{prop:greedy_optimal}
+The greedy policy $\pi_\text{greedy}(\cdot~; Q^*)$ is optimal, in the sense that for all $\os\in\ocS, \oa\in\ocA$:
+\begin{equation}
+    V^{\pi_\text{greedy}(\cdot; Q^*)}(\os) = V^*(\os) \qquad\qquad Q^{\pi_\text{greedy}(\cdot; Q^*)}(\os, \oa) = Q^*(\os, \oa) 
+\end{equation}
+\end{proposition}
 
@@
 When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i=(s_i,\beta_i),\oa_i=(a_i,\beta_{a_i}),R_i=({R_r}_i,{R_c}_i),\os_i'=(s_i',\beta_{a_i})\}_{i\in [0,N]}$ collected by interaction with the environment.
-Following \cref{remark:notations}, we rewrite this batch as  $\cD=\{s_i,a_i,\beta_{a_i},r_i,c_i,s_i'\}_{i\in [0,N]}$.
+Following \cref{remark:notations}, we rewrite this batch as  $\cD=\{s_i,a_i,\beta_{a_i},{R_r}_i,{R_c}_i,s_i'\}_{i\in [0,N]}$.
 Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$.
@@
 
-As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested non-linear constrained optimisation program \eqref{eq:pi_greedy}.
-However, it can be solved efficiently by exploiting the \emph{structure} of the set of solutions with respect to $\beta$, which is concave and increasing.
+%As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested non-linear constrained optimisation program \eqref{eq:pi_greedy}.
+In order to apply $\cT$ in \eqref{eq:bellman-optimality}, it is necessary to compute the greedy policy $\pi_\text{greedy}(s; Q)$.
+Contrary to classical Dynamic Programming where it takes the simple form $\argmax_{a\in\cA} Q(s, a)$, in a budgeted setting is not trivial since it requires solving the nested constrained optimisation program \eqref{eq:pi_greedy}.
+However, it can be solved efficiently by exploiting the \emph{structure} of the set of solutions with respect to $\beta$, that is, concave and increasing.
 
@@
 
-\begin{minipage}[t]{0.5\textwidth}
-\vspace{0pt}
-\begin{figure}[H]
+%\begin{minipage}[t]{0.5\textwidth}
+%\vspace{0pt}
+\begin{wrapfigure}{r}{0.5\textwidth}
     \centering
@@
     \label{fig:hull}
-\end{figure}
-\end{minipage}
-\begin{minipage}[t]{0.5\textwidth}
-\begin{figure}[H]
-    \centering
-    \input{source/architecture.tex}
-    \caption{Neural Network for $Q$-functions approximation when $\cS=\Real^2$ and $|\cA| = 2$.}
-    \label{fig:architecture}
-\end{figure}
-\end{minipage}
+\end{wrapfigure}
+%\end{minipage}
 
@@
 A straightforward choice is a single layer with the same width as the state.
-The overall architecture is shown in Fig.~\ref{fig:architecture}.
+The overall architecture is shown in Supplementary Material Fig.~\ref{fig:architecture}.
 
@@
 A first way of solving a BMDP is to approximate it with a finite set of CMDPs.
-The solutions of these CMDPs take the form of mixtures between two deterministic policies\citep[Theorem 4.4][]{BEUTLER1985236}.
+The solutions of these CMDPs take the form of mixtures between two deterministic policies\citep[Theorem 4.4,][]{BEUTLER1985236}.
 To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material.}.

### 2019-05-21 supplementary.tex
Updates from Overleaf

@@
 \begin{align}
-    Q_r^*(\os, \oa) &=  r(\os, \oa) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa) V_r^{\pi*}(\os') \label{eq:pthm_def}\\
-    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa)  \max_{\rho, \pi'} V_r^{\rho, \pi'}(\os') \label{eq:pthm_opt}\\
-    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
-    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
-    &= r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
+    Q_r^*(\os, \oa) &=  \max_{\rho, \pi'} Q_r^{\rho, \pi'}(\os', \oa') \label{eq:pthm_def}\\
+    &= \max_{\rho, \pi'} R_r(\os, \oa) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa) V_r^{\rho, \pi'}(\os') \label{eq:pthm_exp}\\
+    &= R_r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
+    &= R_r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
+    &= R_r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
 \end{align}
@@
 \begin{enumerate}
-\item[\eqref{eq:pthm_def}.] Definition of $Q^*$ and Bellman Expectation expansion from Proposition \ref{prop:bellman-expectation}. 
-\item[\eqref{eq:pthm_opt}.] Definition of optimality.
+\item[\eqref{eq:pthm_def}.] Definition of $Q^*$. 
+\item[\eqref{eq:pthm_exp}.] Bellman Expectation expansion from Proposition \ref{prop:bellman-expectation}.
 \item[\eqref{eq:pthm_marg}.] Marginalisation on $\oa'$.
@@
 The same reasoning can be made for $Q_c^*$ by replacing $\max$ operators by $\min$, and $\Pi_a$ by $\Pi_r$.
+\end{proof}
+
+\subsection{Proposition \ref{prop:greedy_optimal}}
+
+\begin{definition}
+We define the Bellman Expectation Operator $\cT^\pi$:\\
+$\forall \pi\in\Pi, Q\in(\Real^2)^{\ocS\ocA}, \os\in\ocS, \oa\in\ocA$,
+\begin{align}
+\label{eq:bellman_exepctation_operator}
+    \cT^\pi Q(\os, \oa) &= R(\os, \oa) + \gamma \sum_{\os'\in\ocS}\sum_{a'\in\ocA}\ov{P}(\os'|\os, \oa)\pi(\oa'|\os') Q(\os')
+\end{align}
+\end{definition}
+
+\begin{lemma}[Properties of $\cT^\pi$]
+\label{lemma:bellman_expectation_contraction}
+$\cT^\pi$ is a $\gamma$-contraction and $Q^\pi$ is its unique fixed point.
+\end{lemma}
+\begin{proof}
+The proof is the same as in the MDP case: let $\pi\in\Pi, Q_1, Q_2\in(\Real^2)^{\ocS\ocA}$.
+\begin{align*}
+    \forall \os\in\ocS, \oa\in\ocA,\quad \left|\cT^\pi Q_1(\os,\oa) - \cT^\pi Q_2(\os,\oa)\right| &= \left|\gamma\expectedvalueover{\substack{\os'\sim\ov{P}(\os'|\os,\oa) \\ \oa'\sim\pi(\oa'|\os')}} Q_1(\os',\oa') - Q_2(\os',\oa')\right|\\
+    &\leq \gamma\left\|Q_1-Q_2\right\|_\infty
+\end{align*}
+Hence, $\left\|\cT^\pi Q_1  - \cT^\pi Q_2 \right\|_\infty \leq \gamma\left\|Q_1-Q_2\right\|_\infty$
+
+By Banach fixed point theorem, $\cT^\pi$ admits a unique fixed point.
+It can be easily verified that $Q^\pi$ is indeed this fixed point by combining the two Bellman Expectation equations \eqref{eq:bellman_expectation_V} and \eqref{eq:bellman_expectation_Q}.
+\end{proof}
+
+We now prove the result of \Cref{prop:greedy_optimal}.
+\begin{proof}
+Notice from the definitions of $\cT$ and $\cT^\pi$ in \eqref{eq:bellman-optimality} and \eqref{eq:bellman_exepctation_operator} that $\cT$ and $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$ coincide on $Q^*$.
+Moreover, since $Q^* = \cT Q^*$ by \cref{thm:bellman-optimality}, we have: $    \cT^{\pi_\text{greedy}(\cdot;Q^*)} Q^* = \cT Q^* = Q^*
+$.
+Hence, $Q^*$ is a fixed point of $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$, and by \cref{lemma:bellman_expectation_contraction} it must be equal to $Q^{\pi_\text{greedy}(\cdot;Q^*)}$
+
+To show the same result for $V^*$, notice that 
+\begin{equation*}
+    V^{\pi_\text{greedy}(Q^*)}(\os) = \expectedvalueover{\oa\sim\pi_\text{greedy}(Q^*)}Q^{\pi_\text{greedy}(Q^*)}(\os,\oa) = \expectedvalueover{\oa\sim\pi_\text{greedy}(Q^*)}Q^*(\os,\oa)
+\end{equation*}
+By applying the definitions of $Q^*$ and $\pi_\text{greedy}$, we recover the definition of $V^*$.
 \end{proof}
@@
 
-The original problem in the space of actions $\ocA$ is now expressed in the space of values $Q(\os, \ocA)$ (which is why we use $=$ instead of $\in$ before $\argmin$ here, see Remark \ref{rmk:independence}).
+The original problem in the space of actions $\ocA$ is now expressed in the space of values $Q(\os, \ocA)$ (which is why we use $=$ instead of $\in$ before $\argmin$ here).
 
@@
 
-We recall the scalable version of BFTQ in \cref{algo:bftq_full}
+We recall the scalable version of BFTQ in \cref{algo:bftq_full} and the architecture of the neural network \cref{fig:architecture}.
+
+\begin{minipage}[t]{0.5\textwidth}
+\begin{figure}[H]
+    \centering
+    \input{source/architecture.tex}
+    \caption{Neural Network for $Q$-functions approximation when $\cS=\Real^2$ and $|\cA| = 2$.}
+    \label{fig:architecture}
+\end{figure}
+\end{minipage}
 

### 2019-05-21 main.tex
Updates from Overleaf

@@
 \usepackage[lofdepth,lotdepth]{subfig}
+\usepackage{wrapfig}
+\usepackage{enumerate}
 
@@
 
-However, this modelling assumption comes at a price: no control is given over the spread of the performance distribution.
+However, this modelling assumption comes at a price: no control is given over the spread of the performance distribution \citep{Dann2018}.
 In many critical real-world applications where failures may turn out very costly, this is an issue as most decision-makers would rather give away some amount of expected optimality to increase the performances in the lower-tail of the distribution.
@@
 
-Reinforcement Learning also assumes that the performance can be described solely by a reward function $R_r$.
-Conversely, real problems typically involve many aspects, some of which can be contradictory \citep{ChunmingLiu2014}.
+Reinforcement Learning also assumes that the performance can be described by a single reward function $R_r$.
+Conversely, real problems typically involve many aspects, some of which can be contradictory \citep{Liu2014}.
 For instance, a self-driving car needs to balance between progressing quickly on the road and avoiding collisions.
@@
 \begin{array}{lcr}
- \max_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s, \pi] & \text{ s.t. } & \expectedvalue[G_c^\pi | s_0=s, \pi] \leq \beta
+ \displaystyle \max_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s, \pi] & \text{ s.t. } & \expectedvalue[G_c^\pi | s_0=s, \pi] \leq \beta
 \end{array}
@@
 To address this concern, the \emph{Budgeted Markov Decision Process} (BMDP) was introduced in \citep{Boutilier_Lu:uai16} as an extension of CMDPs to enable the online control over the budget $\beta$ within an interval $\cB \subset \Real$ of admissible budgets.
-Instead of fixing the budget prior to training, the objective is now to find a generic optimal policy $\pi^*$ that takes $\beta$ as input so as to solve the corresponding CMDP \eqref{eq:cmdp} for all $\beta\in\cB$.
+Instead of fixing the budget prior to training, the objective is now to find a generic optimal policy $\pi^*$ that takes $\beta$ as input so as to solve the corresponding CMDP (Eq.
+\eqref{eq:cmdp}) for all $\beta\in\cB$.
 This gives the system designer the ability to move the optimal policy $\pi^*$ in real-time along the Pareto-optimal curve of the different reward-cost trade-offs.
@@
 Second, we use this operator in a Reinforcement Learning algorithm for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure.
-Third, we scale this algorithms to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
+Third, we scale this algorithm to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
 Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.
-
 The proofs of our main results are provided in Supplementary Material \ref{sec:proofs}.
@@
 
-We will work in the space of budgeted policies, where $\pi$ both depends on $\beta$ and also outputs a next budget $\beta_a$.
+We work in the space of budgeted policies, where $\pi$ both depends on $\beta$ and also outputs a next budget $\beta_a$.
 Hence, the budget $\beta$ is neither fixed nor constant as in the CMDP setting but instead evolves as part of the dynamics.
@@
 For ease of notation, we consider the \emph{augmented} state and action spaces $\ocS = \cS\times \cB$ and $\ocA= \cA\times \cB$, and equip them with the augmented dynamics $\ov{P}\in \cM(\ocS)^{\ocS \times \ocA}$ defined as:
-\begin{equation*}
+\begin{equation}
+    \label{eq:dynamics}
     \ov{P}\left(\os' \condbar \os, \oa\right) = \ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) \eqdef P(s'|s, a)\delta(\beta' - \beta_a),
-\end{equation*}
+\end{equation}
 where $\delta$ is the Dirac distribution.
@@
 
-As earlier, the return $G^\pi = (G_r^\pi, G_c^\pi)$ of a budgeted policy $\pi\in\Pi$ refers to:
-    $G^\pi \eqdef \sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t)$.
-
-We can introduce the value function $V^\pi$ and action-value function $Q^\pi$ of a budgeted policy $\pi\in\Pi$ as:
+Likewise, the return $G^\pi = (G_r^\pi, G_c^\pi)$ of a budgeted policy $\pi\in\Pi$ refers to:
+$G^\pi \eqdef \sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t)$,
+and the value functions $V^\pi$, $Q^\pi$ of a budgeted policy $\pi\in\Pi$ are defined as:
 \begin{equation}
     \label{eq:value-function}
-V^\pi(\os) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right].
+V^\pi(\os) = (V_r^\pi, V_c^\pi) = \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa)= (Q_r^\pi, Q_c^\pi) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right].
 \end{equation}
-
-Likewise, we note $V^\pi = (V_r^\pi, V_c^\pi)$ and $Q^\pi = (Q_r^\pi, Q_c^\pi)$.
 
@@
 \label{prop:bellman-expectation}
-The value functions $V^\pi, Q^\pi$ of a budgeted policy $\pi$ verify:
+The value functions $V^\pi and Q^\pi$ verify:
 \begin{align}
-    V^\pi(\os) &= \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa),\label{eq:bellman_expectation_V}\\
-    Q^\pi(\os, \oa) &= R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') \label{eq:bellman_expectation_Q}.
+    V^\pi(\os) = \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa) \qquad Q^\pi(\os, \oa) = R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') \label{eq:bellman_expectation}
 \end{align}
+Moreover, consider the Budgeted Bellman Expectation operator $\cT^\pi$:
+$\forall Q\in(\Real^2)^{\ocS\ocA}, \os\in\ocS, \oa\in\ocA$,
+\begin{align}
+\label{eq:bellman_expectation_operator}
+    \cT^\pi Q(\os, \oa) &\eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS}\sum_{\oa'\in\ocA}\ov{P}(\os'|\os, \oa)\pi(\oa'|\os') Q(\os', \oa')
+\end{align}
+Then $\cT^\pi$ is a $\gamma$-contraction and $Q^\pi$ is its unique fixed point.
 \end{proposition}
@@
 We now come to the definition of budgeted optimality.
-We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the rewards $R_r$, (iii)~in case of tie, minimise the costs $R_c$.
+We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the $\gamma$-discounted return of rewards $G_r$, (iii)~in case of tie, minimise the $\gamma$-discounted return of costs $G_c$.
 To that end, we define for all $\os\in\ocS$:
-\begin{itemize}
-    \item Admissible policies:
+\begin{enumerate}[(i)]
+    \item Admissible policies $\Pi_a$: 
     \begin{equation}
@@
     \end{equation}
-    \item Optimal value function for rewards $ V_r^*$ and candidate policies $\Pi_r$:
+    \item Optimal value function for rewards $V_r^*$ and candidate policies $\Pi_r$: 
     \begin{equation}
-        V_r^*(\os) \eqdef \max_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\qquad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
+        V_r^*(\os), \eqdef \max_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\qquad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
     \end{equation}
@@
     \begin{equation}
-        V_c^*(\os) \eqdef \min_{\pi\in\Pi_r(\os)}  V_c^\pi(\os) \qquad\qquad \Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
+        V_c^*(\os) \eqdef \min_{\pi\in\Pi_r(\os)}  V_c^\pi(\os), \qquad\qquad \Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
     \end{equation}
-\end{itemize}
+\end{enumerate}
 We define the budgeted action-value function $Q^*$ similarly:
@@
 \end{equation}
-and denote $V^* = (V_r^*, V_c^*)$, $Q = (Q_r^*, Q_c^*)$.
+and denote $V^* = (V_r^*, V_c^*)$, $Q^* = (Q_r^*, Q_c^*)$.
+\end{definition}
+
+\begin{definition}[Budgeted Optimality (tentative de Romain pour remplacer la precedente definition)]
+We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the expected $\gamma$-discounted return of rewards, (iii)~in case of tie, minimise the expected $\gamma$-discounted return of costs.
+Let $V_c^\bot = \min_{\pi\in\cM(\cA)^{\cS}} V_c^\pi$ denote the minimal cost return.
+Let $\overline{S}_b = \{s,\beta\}_{s\in\cS, \beta\geq V_c^\bot(s)}$ denote the feasible augmented state space.
+To that end, we define the following:
+\begin{enumerate}[(i)]
+    \item Admissible policies set: $\Pi_a \eqdef \{\pi\in\Pi: V_c^\pi(\os=(s, \beta)) \leq \beta ;\;\; \forall \os\in\overline{S}_b\}$
+    \item Candidate policies set: $\Pi_r \eqdef \argmax_{\pi\in\Pi_a}  V_r^\pi$
+    \item Optimal policies set: $\Pi^* \eqdef \argmin_{\pi\in\Pi_r}  V_c^\pi$
+\end{enumerate}
+Additionally, let $V^* = (V_r^*, V_c^*)$ and $Q^* = (Q_r^*, Q_c^*)$ denote the values of the optimal policies in $\Pi^*$.
 \end{definition}
@@
 \label{eq:pi_greedy}
-\begin{equation}
-    \pi_\text{greedy}(\oa|\os; Q) \in \argmin_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa), \label{eq:pi_greedy_cost}
-\end{equation}
 \begin{align}
+    \pi_\text{greedy}(\oa|\os; Q) \in &\argmin_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa), \label{eq:pi_greedy_cost}\\
     \text{where }\quad\Pi_r^Q \eqdef &\argmax_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \label{eq:pi_greedy_reward}\\
@@
 
-% \begin{remark}[Appearance of the greedy policy]
-% \label{rmk:greedy}
-% In classical Reinforcement Learning, the greedy policy takes a simple form $\pi_\text{greedy}(s; Q^*) = \argmax_{a\in\cA} Q^*(s, a)$, and the term $\pi_\text{greedy}(a'|s';Q^*) Q^{*}(s', a')$ in \eqref{eq:bellman-optimality} conveniently simplifies to $\max_{a'\in \cA} Q^*(s', a')$.
+\begin{remark}[Appearance of the greedy policy]
+In classical Reinforcement Learning, the greedy policy takes a simple form $\pi_\text{greedy}(s; Q^*) = \argmax_{a\in\cA} Q^*(s, a)$, and the term $\pi_\text{greedy}(a'|s';Q^*) Q^{*}(s', a')$ in \eqref{eq:bellman-optimality} conveniently simplifies to $\max_{a'\in \cA} Q^*(s', a')$.
 Unfortunately, in a budgeted setting the greedy policy requires solving the nested constrained optimisation program \eqref{eq:pi_greedy} at each state and budget in order to apply this Budgeted Bellman Optimality operator.
-% \end{remark}
+\end{remark}
 
@@
 \label{prop:greedy_optimal}
-The greedy policy $\pi_\text{greedy}(\cdot~; Q^*)$ is optimal, in the sense that for all $\os\in\ocS, \oa\in\ocA$:
-\begin{equation}
-    V^{\pi_\text{greedy}(\cdot; Q^*)}(\os) = V^*(\os) \qquad\qquad Q^{\pi_\text{greedy}(\cdot; Q^*)}(\os, \oa) = Q^*(\os, \oa) 
-\end{equation}
+The greedy policy $\pi_\text{greedy}(\cdot~; Q^*)$ is \emph{uniformly} optimal: for all $\os\in\ocS$, $\pi_\text{greedy}(\cdot~; Q^*)\in\Pi^*(\os)$.
+In particular, $V^{\pi_\text{greedy}(\cdot; Q^*)} = V^*$ and $Q^{\pi_\text{greedy}(\cdot; Q^*)}= Q^*$.
 \end{proposition}
 
-\begin{remark}[Notations]
-Because of the budget dynamics $\beta' = \beta_a$, we can see in \eqref{eq:bellman_expectation_Q} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
-We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
-\label{remark:notations}
-\end{remark}
+\begin{conjecture}[Contractivity of $\cT$]
+We conjecture that $\cT$ is a $\gamma$-contraction.
+\end{conjecture}
 
@@
 We introduce Algorithm~\ref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.
+
+Because of the budget dynamics $\beta' = \beta_a$ in \eqref{eq:dynamics}, we can see in \eqref{eq:bellman_expectation} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
+We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
 
@@
 When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i=(s_i,\beta_i),\oa_i=(a_i,\beta_{a_i}),R_i=({R_r}_i,{R_c}_i),\os_i'=(s_i',\beta_{a_i})\}_{i\in [0,N]}$ collected by interaction with the environment.
-Following \cref{remark:notations}, we rewrite this batch as  $\cD=\{s_i,a_i,\beta_{a_i},{R_r}_i,{R_c}_i,s_i'\}_{i\in [0,N]}$.
 Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$.
-We instead use $\hat{\cT}$ in which this expectation is replaced by a sample average:
+We instead use $\hat{\cT}$, a sampling operator, in which this expectation is replaced by:
 \begin{equation*}
-    \hat{\cT} Q(\os, \oa, R, \os') \eqdef R + \gamma \sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q) Q(\ov{s'}, \ov{a'}).
+    \hat{\cT} Q(\os_i, \oa_i, R_i, \os'_i) \eqdef R_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
 \end{equation*}
@@
 
-%As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested non-linear constrained optimisation program \eqref{eq:pi_greedy}.
-In order to apply $\cT$ in \eqref{eq:bellman-optimality}, it is necessary to compute the greedy policy $\pi_\text{greedy}(s; Q)$.
+As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested constrained optimisation program \eqref{eq:pi_greedy}.
+% In order to apply $\cT$ in \eqref{eq:bellman-optimality}, it is necessary to compute the greedy policy $\pi_\text{greedy}(s; Q)$.
 Contrary to classical Dynamic Programming where it takes the simple form $\argmax_{a\in\cA} Q(s, a)$, in a budgeted setting is not trivial since it requires solving the nested constrained optimisation program \eqref{eq:pi_greedy}.
@@
 Second, the main bottleneck of \BFTQ is the computation of the target $\cT Q$.
-Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $O(|\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$.
-The resulting total time-complexity is $O(|\mathcal{D}|/(1-\gamma) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$.
+Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $\cO(|\cA||\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$.
+The resulting total time-complexity is $\cO(\frac{|\mathcal{D}||\cA||\tilde{\cB}|}{1-\gamma} \log |\cA||\tilde{\cB}|)$.
 This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA,\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass.
@@
 This baseline consists in approximating the BMDP by a finite set of CMDPs problems.
-We solve each of these CMDP using the standard technique of Lagrangian Relaxation: the cost constraint is converted to a soft penalty weighted by a Lagrangian multiplier $\lambda$ in a surrogate reward function: $\max_{\pi} \expectedvalue_{\pi} G_r^\pi - \lambda G_c^\pi\label{eq:lagrangian}$.
+We solve each of these CMDP using the standard technique of Lagrangian Relaxation: the cost constraint is converted to a soft penalty weighted by a Lagrangian multiplier $\lambda$ in a surrogate reward function: $\max_{\pi} \expectedvalue_{\pi}[G_r^\pi - \lambda G_c^\pi]$.
 The resulting MDP can be solved by any RL algorithm, and we chose $\FTQ$ for being closest to $\BFTQ$.
@@
 When the environment parameters ($P$, $R_r$, $R_c$) are known but not tractable, solutions relying on function approximation~\citep{Undurti} or approximate linear programming~\citep{Poupart2015} have been proposed.
-For unknown environments, batch algorithms exist in the tabular setting \citep{Thomas2015, Petrik2016, Laroche2017}.
+For unknown environments, batch algorithms exist in the tabular setting \citep{Thomas2015, Petrik2016, Laroche2019}.
 Online algorithms \citep{Geibel2005, Abe2010,ChowGJP15,AchiamHTA17} and a batch algorithm \citep{le2019batch} combine the best of both world by solving intractable unknown CMDPs.
@@
 The solutions of these CMDPs take the form of mixtures between two deterministic policies\citep[Theorem 4.4,][]{BEUTLER1985236}.
-To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material.}.
+To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material \ref{sec:algorithms-parameters}}.
 Our solution not only requires one single model but also avoids any supplementary interaction.

### 2019-05-21 supplementary.tex
Updates from Overleaf

@@
 \end{align*}
+
+\textbf{Contraction of $\cT^\pi$:}
+Let $\pi\in\Pi, Q_1, Q_2\in(\Real^2)^{\ocS\ocA}$.
+\begin{align*}
+    \forall \os\in\ocS, \oa\in\ocA,\quad \left|\cT^\pi Q_1(\os,\oa) - \cT^\pi Q_2(\os,\oa)\right| &= \left|\gamma\expectedvalueover{\substack{\os'\sim\ov{P}(\os'|\os,\oa) \\ \oa'\sim\pi(\oa'|\os')}} Q_1(\os',\oa') - Q_2(\os',\oa')\right|\\
+    &\leq \gamma\left\|Q_1-Q_2\right\|_\infty
+\end{align*}
+Hence, $\left\|\cT^\pi Q_1  - \cT^\pi Q_2 \right\|_\infty \leq \gamma\left\|Q_1-Q_2\right\|_\infty$
+
+According to the Banach fixed point theorem, $\cT^\pi$ admits a unique fixed point.
+It can be easily verified that $Q^\pi$ is indeed this fixed point by combining the two Bellman Expectation equations \eqref{eq:bellman_expectation}.
+
 \end{proof}
@@
 \subsection{Proposition \ref{prop:greedy_optimal}}
-
-\begin{definition}
-We define the Bellman Expectation Operator $\cT^\pi$:\\
-$\forall \pi\in\Pi, Q\in(\Real^2)^{\ocS\ocA}, \os\in\ocS, \oa\in\ocA$,
-\begin{align}
-\label{eq:bellman_exepctation_operator}
-    \cT^\pi Q(\os, \oa) &= R(\os, \oa) + \gamma \sum_{\os'\in\ocS}\sum_{a'\in\ocA}\ov{P}(\os'|\os, \oa)\pi(\oa'|\os') Q(\os')
-\end{align}
-\end{definition}
-
-\begin{lemma}[Properties of $\cT^\pi$]
-\label{lemma:bellman_expectation_contraction}
-$\cT^\pi$ is a $\gamma$-contraction and $Q^\pi$ is its unique fixed point.
-\end{lemma}
 \begin{proof}
-The proof is the same as in the MDP case: let $\pi\in\Pi, Q_1, Q_2\in(\Real^2)^{\ocS\ocA}$.
-\begin{align*}
-    \forall \os\in\ocS, \oa\in\ocA,\quad \left|\cT^\pi Q_1(\os,\oa) - \cT^\pi Q_2(\os,\oa)\right| &= \left|\gamma\expectedvalueover{\substack{\os'\sim\ov{P}(\os'|\os,\oa) \\ \oa'\sim\pi(\oa'|\os')}} Q_1(\os',\oa') - Q_2(\os',\oa')\right|\\
-    &\leq \gamma\left\|Q_1-Q_2\right\|_\infty
-\end{align*}
-Hence, $\left\|\cT^\pi Q_1  - \cT^\pi Q_2 \right\|_\infty \leq \gamma\left\|Q_1-Q_2\right\|_\infty$
-
-By Banach fixed point theorem, $\cT^\pi$ admits a unique fixed point.
-It can be easily verified that $Q^\pi$ is indeed this fixed point by combining the two Bellman Expectation equations \eqref{eq:bellman_expectation_V} and \eqref{eq:bellman_expectation_Q}.
-\end{proof}
-
-We now prove the result of \Cref{prop:greedy_optimal}.
-\begin{proof}
-Notice from the definitions of $\cT$ and $\cT^\pi$ in \eqref{eq:bellman-optimality} and \eqref{eq:bellman_exepctation_operator} that $\cT$ and $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$ coincide on $Q^*$.
+Notice from the definitions of $\cT$ and $\cT^\pi$ in \eqref{eq:bellman-optimality} and \eqref{eq:bellman_expectation_operator} that $\cT$ and $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$ coincide on $Q^*$.
 Moreover, since $Q^* = \cT Q^*$ by \cref{thm:bellman-optimality}, we have: $    \cT^{\pi_\text{greedy}(\cdot;Q^*)} Q^* = \cT Q^* = Q^*
 $.
-Hence, $Q^*$ is a fixed point of $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$, and by \cref{lemma:bellman_expectation_contraction} it must be equal to $Q^{\pi_\text{greedy}(\cdot;Q^*)}$
+Hence, $Q^*$ is a fixed point of $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$, and by \cref{prop:bellman-expectation} it must be equal to $Q^{\pi_\text{greedy}(\cdot;Q^*)}$
 
@@
 
-\begin{minipage}[t]{0.5\textwidth}
-\begin{figure}[H]
+%\begin{minipage}[t]{0.5\textwidth}
+\begin{figure}[tp]
     \centering
@@
 \end{figure}
-\end{minipage}
+%\end{minipage}
 

### 2019-05-23 main.tex
Updates from Overleaf

@@
 \usepackage{enumerate}
+\usepackage{minted}
 
@@
 
-%---------------------------------------------
-% Cross Referencing of Supplementary Material %---------------------------------------------
-% https://tex.stackexchange.com/questions/14364/cross-referencing-between-different-files
-% https://tex.stackexchange.com/questions/244239/correct-ordering-of-xr-hyper-hyperref-and-cleveref
-% \usepackage{xr}
-% \usepackage{xr-hyper}
 \usepackage{hyperref}
-\usepackage[capitalise,noabbrev]{cleveref}
-% \makeatletter
-% \newcommand*{\addFileDependency}[1]{% argument=file name and extension
-%   \typeout{(#1)}
-%   \@addtofilelist{#1}
-%   \IfFileExists{#1}{}{\typeout{No file #1.}}
-% }
-% \makeatother
-% \newcommand*{\myexternaldocument}[1]{%
-%     \externaldocument{#1}%
-%     \addFileDependency{#1.tex}%
-%     \addFileDependency{#1.aux}%
-% }
-% \myexternaldocument{supplementary}
-
+\usepackage{cleveref}
 \title{Budgeted Reinforcement Learning in Continuous State Space}
@@
 In many critical real-world applications where failures may turn out very costly, this is an issue as most decision-makers would rather give away some amount of expected optimality to increase the performances in the lower-tail of the distribution.
-This has led to the development of several risk-averse variants where the optimisation criteria include other statistics of the performance, such as the worst-case realisation \citep{Iyengar2005,Nilim2005,Wiesemann2013}, the variance-penalised expectation \citep{Garcia2015,Tamar2012}, or the Conditional Value-At-Risk (CVaR) \citep{Chow2014,DBLP:journals/corr/ChowGJP15}.
+This has led to the development of several risk-averse variants where the optimisation criteria include other statistics of the performance, such as the worst-case realisation \citep{Iyengar2005,Nilim2005,Wiesemann2013}, the variance-penalised expectation \citep{Garcia2015,Tamar2012}, or the Conditional Value-At-Risk (CVaR) \citep{Chow2014,ChowGJP15}.
 
@@
 \end{equation}
+We restrict $\ocS$ to feasible budgets only: $\ocS_f \eqdef \{(s,\beta)\in\ocS: \exists \pi\in\Pi, V_c^\pi(s) \geq \beta\}$, but still write $\ocS$ in place of $\ocS_f$ for brevity of notations.
 
@@
 \label{prop:bellman-expectation}
-The value functions $V^\pi and Q^\pi$ verify:
+The value functions $V^\pi$ and $Q^\pi$ verify:
 \begin{align}
@@
 
-\begin{definition}[Budgeted Optimality (tentative de Romain pour remplacer la precedente definition)]
-We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the expected $\gamma$-discounted return of rewards, (iii)~in case of tie, minimise the expected $\gamma$-discounted return of costs.
-Let $V_c^\bot = \min_{\pi\in\cM(\cA)^{\cS}} V_c^\pi$ denote the minimal cost return.
-Let $\overline{S}_b = \{s,\beta\}_{s\in\cS, \beta\geq V_c^\bot(s)}$ denote the feasible augmented state space.
-To that end, we define the following:
-\begin{enumerate}[(i)]
-    \item Admissible policies set: $\Pi_a \eqdef \{\pi\in\Pi: V_c^\pi(\os=(s, \beta)) \leq \beta ;\;\; \forall \os\in\overline{S}_b\}$
-    \item Candidate policies set: $\Pi_r \eqdef \argmax_{\pi\in\Pi_a}  V_r^\pi$
-    \item Optimal policies set: $\Pi^* \eqdef \argmin_{\pi\in\Pi_r}  V_c^\pi$
-\end{enumerate}
-Additionally, let $V^* = (V_r^*, V_c^*)$ and $Q^* = (Q_r^*, Q_c^*)$ denote the values of the optimal policies in $\Pi^*$.
-\end{definition}
-
 \begin{theorem}[Budgeted Bellman Optimality]
@@
 \begin{remark}[Appearance of the greedy policy]
+\label{rmk:greedy}
 In classical Reinforcement Learning, the greedy policy takes a simple form $\pi_\text{greedy}(s; Q^*) = \argmax_{a\in\cA} Q^*(s, a)$, and the term $\pi_\text{greedy}(a'|s';Q^*) Q^{*}(s', a')$ in \eqref{eq:bellman-optimality} conveniently simplifies to $\max_{a'\in \cA} Q^*(s', a')$.
@@
 
-\begin{conjecture}[Contractivity of $\cT$]
-We conjecture that $\cT$ is a $\gamma$-contraction.
-\end{conjecture}
+\begin{theorem}[Contractivity of $\cT$]
+\label{thm:contraction}
+If $|\cA| \geq 2$, then $\cT$ is not a contraction.\\
+However, we conjecture that $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
+We give some intuition of why that should be the case in Supplementary Material \ref{sec:proof_contraction}.
+\end{theorem}
 
 \paragraph{Budgeted Value Iteration}
-
+{
 The Budgeted Bellman Optimality equation is a fixed-point equation, which motivates the introduction of a fixed-point iteration procedure.
 We introduce Algorithm~\ref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.
+
+Unfortunately, as $\cT$ is not a contraction, we can guarantee neither the convergence of this procedure nor the unicity of its fixed points.
+However, we empirically observed the convergence in our experiments (\Cref{sec:experiements}).
+}
 
@@
 
-When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i=(s_i,\beta_i),\oa_i=(a_i,\beta_{a_i}),R_i=({R_r}_i,{R_c}_i),\os_i'=(s_i',\beta_{a_i})\}_{i\in [0,N]}$ collected by interaction with the environment.
+When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$ collected by interaction with the environment.
 Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$.
@@
 \begin{equation*}
-    \hat{\cT} Q(\os_i, \oa_i, R_i, \os'_i) \eqdef R_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
+    \hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
 \end{equation*}
@@
 \label{prop:bftq_pi_hull}
-Algorithm~\ref{algo:bvi} and Algorithm~\ref{algo:bftq} can be run by replacing $\pi_\text{greedy}$ in the definition \eqref{eq:bellman-optimality} of $\cT$ with $\pi_\text{hull}$ as described in Algorithm~\ref{algo:pi_hull}.
+Algorithm~\ref{algo:bvi} and Algorithm~\ref{algo:bftq} can be run by replacing $\pi_\text{greedy}$ in the equation \eqref{eq:bellman-optimality} of $\cT$ with $\pi_\text{hull}$ as described in Algorithm~\ref{algo:pi_hull}.
 \begin{equation*}
@@
 
-The computation of $\pi_\text{hull}$ in \cref{algo:pi_hull} is illustrated in \cref{fig:hull}.
+The computation of $\pi_\text{hull}$ in Alg.~\ref{algo:pi_hull} is illustrated in Fig.~\ref{fig:hull}.
 
@@
 Second, the main bottleneck of \BFTQ is the computation of the target $\cT Q$.
-Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $\cO(|\cA||\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$.
+Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $\cO(|\cA||\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$ (where $\tilde{\cB}$ is a finite discretisation of $\cB$).
 The resulting total time-complexity is $\cO(\frac{|\mathcal{D}||\cA||\tilde{\cB}|}{1-\gamma} \log |\cA||\tilde{\cB}|)$.
@@
 \section{Experiments}
-
+\label{sec:experiements}
 There are two hypotheses we want to validate.
@@
 The resulting MDP can be solved by any RL algorithm, and we chose $\FTQ$ for being closest to $\BFTQ$.
-In these experiments, each run is composed of a training and a test phase and repeated $N_{\text{seeds}}$ times.
+In our experiments, a single training of $\BFTQ$ corresponds to 10 trainings of $\FTQ_l$ policies.
+Each run was repeated $N_{\text{seeds}}$ times.
 Parameters of the algorithms can be found in Supplementary Material~\ref{sec:algorithms-parameters}
@@
 
-In the following figures, each patch represents the mean and 95\% confidence interval over $N_{\text{seeds}}$ seeds of the means of $(G_r^\pi,G_c^\pi)$ over $N_\text{trajs}$ trajectories\footnote{Examples of policy execution can be found in Supplementary Material~\ref{sec:bftq-executions}}.
+In the following figures, each patch represents the mean and 95\% confidence interval over $N_{\text{seeds}}$ seeds of the means of $(G_r^\pi,G_c^\pi)$ over $N_\text{trajs}$ trajectories.
 That way, we display the variation related to learning (and batches) rather than the variation in the execution of the policies.
 
-We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: Figure~\ref{fig:exploration} shows the set of trajectories collected by each exploration strategy, and the resulting performance of a budgeted policy trained on each batch.
+We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: Figure~\ref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in supplementary material ~\ref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch.
 The trajectories (purple) in the risk-neutral batch (left) are concentrated along the risky corridor (orange) and ignore the safe corridor (green), which results in bad performances in the low-risk regime.
 Conversely, trajectories in the risk-sensitive batch (right) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.
+
 \begin{figure}[tp]
     \centering
-    \includegraphics[width=0.23\textwidth]{source/img/risk-neutral.png}
-    \includegraphics[width=0.23\textwidth]{source/img/risk-sensitive.png}
+    \includegraphics[width=0.23\textwidth]{source/img/test.pdf}
+    %\includegraphics[width=0.23\textwidth]{source/img/risk-sensitive.png}
     \includegraphics[page=1, width=0.45\textwidth]{source/img/corridors}
@@
 \end{figure}
-% \begin{figure}[tp]
-%     \begin{center}
-%         \subfloat[\texttt{slot-filling}]{
-%         \includegraphics[width=0.5\textwidth]{source/img/slot-filling}
-%         \label{fig:slot-filling}
-%         }
-%         \subfloat[\texttt{highway-env}]{
-%         \includegraphics[width=0.5\textwidth]{source/img/highway}
-%         \label{fig:highway-env}
-%         }
-%         \caption{Performance comparison of \FTQl and BFTQ}
-%         \label{fig:results}
-%     \end{center}
-% \end{figure}
+
 \begin{figure}[tp]
@@
 In both cases, BFTQ performs almost as well as \FTQl despite only requiring a single model.
-All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can notice in the bottom-left corner an underestimation of $Q_c$, since $\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
-However, these estimates still provide a reasonable prior approximation of the expected cost of the policy, which cannot be obtained with \FTQl.
-In addition, BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance.
-Qualitatively, the budgeted agents display a wide variety of behaviours.
-In the \texttt{highway-env} experiment, when $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic.
-On the contrary when $\beta = 0$, the ego-vehicle is conservative: it stays on its lane and drives at a low velocity.
-With intermediate budgets such as $\beta = 0.2$, the agent sometimes decides to overtake its front vehicle but promptly steers back to its original lane afterwards.
+All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can observe in the bottom-left corner an underestimation of $Q_c$, since $\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
+%This underestimation can be a consequence of two approximations: the use of the sampling operator $\hat{\cT}$ instead of the true environmental operator $\cT$, and the use of the neural network function approximation $Q_\theta$ instead of $Q$.
+This is probably due to the fact that \text{highway-env} is actually infeasible for $\beta=0$, since we noticed that about $8\%$ of bad initialisations occur with the front vehicle driving slower than the minimum velocity of the controlled vehicle, leading to either a front collision or a forced change to the opposite lane, both yielding costs.
+Still, \BFTQ provides a better control on the expected cost of the policy, than \FTQl.
+In addition, \BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance.
+Examples of policy execution can be found in Supplementary Material~\ref{sec:bftq-executions}.
 
@@
 For unknown environments, batch algorithms exist in the tabular setting \citep{Thomas2015, Petrik2016, Laroche2019}.
-Online algorithms \citep{Geibel2005, Abe2010,ChowGJP15,AchiamHTA17} and a batch algorithm \citep{le2019batch} combine the best of both world by solving intractable unknown CMDPs.
+Online algorithms \citep{Geibel2005, Abe2010,ChowGJP15,AchiamHTA17} and a batch algorithm \citep{le2019batch} combine the best of both world by solving large unknown CMDPs.
 
@@
 The solutions of these CMDPs take the form of mixtures between two deterministic policies\citep[Theorem 4.4,][]{BEUTLER1985236}.
-To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material \ref{sec:algorithms-parameters}}.
+To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material ~\ref{sec:lagragian}}.
 Our solution not only requires one single model but also avoids any supplementary interaction.

### 2019-05-23 optimality_definition.tex
Updates from Overleaf

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{xargs}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{enumerate}
\input{mathdef.tex}

\begin{document}

\begin{definition}[Budgeted Optimality (tentative de Romain pour remplacer la precedente definition)]
We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the expected $\gamma$-discounted return of rewards, (iii)~in case of tie, minimise the expected $\gamma$-discounted return of costs. Let $V_c^\bot = \min_{\pi\in\cM(\cA)^{\cS}} V_c^\pi$ denote the minimal cost return. Let $\overline{S}_b = \{s,\beta\}_{s\in\cS, \beta\geq V_c^\bot(s)}$ denote the feasible augmented state space. To that end, we define the following:
\begin{enumerate}[(i)]
    \item Admissible policies set: $\Pi_a \eqdef \{\pi\in\Pi: V_c^\pi(\os=(s, \beta)) \leq \beta ;\;\; \forall \os\in\overline{S}_b\}$
    \item Candidate policies set: $\Pi_r \eqdef \argmax_{\pi\in\Pi_a}  V_r^\pi$
    \item Optimal policies set: $\Pi^* \eqdef \argmin_{\pi\in\Pi_r}  V_c^\pi$
\end{enumerate}
Additionally, let $V^* = (V_r^*, V_c^*)$ and $Q^* = (Q_r^*, Q_c^*)$ denote the values of the optimal policies in $\Pi^*$.
\end{definition}
\end{document}

### 2019-05-23 supplementary.tex
Updates from Overleaf

@@
 
+\subsection{Theorem \ref{thm:contraction}}
+\label{sec:proof_contraction}
+In the trivial case $|\cA| = 1$, there exits only one policy $\pi$ and $\cT = \cT^\pi$, which is a contraction by \cref{prop:bellman-expectation}.
+
+In the general case $|\cA| \geq 2$, we can build the following counter-example:
+
+Let $(\cS, \cA, P, R_r, R_c)$ be a BMDP.
+For any $0 < \epsilon < 1$, we define $Q_\epsilon^1$ and $Q_\epsilon^2$ as:
+
+\begin{align*}
+      Q_\epsilon^1(\os,\oa) =
+      \begin{cases}
+    (0, 1), & \text{if } a = a_0 \\
+    (\frac{1}{\epsilon}, 1+\epsilon), & \text{if } a \neq a_0
+  \end{cases}\\
+  Q_\epsilon^2(\os,\oa) =
+      \begin{cases}
+    (1, 0), & \text{if } a = a_0 \\
+    (1+\frac{1}{\epsilon}, \epsilon), & \text{if } a \neq a_0
+  \end{cases}
+\end{align*}
+Then, $\|Q_1-Q_2\|_\infty = 1$.
+ $Q_\epsilon^1$ and $Q_\epsilon^2$ are represented in \cref{fig:concavity_example}.
+
+\begin{figure}[tp]
+    \centering
+    \includegraphics[width=0.5\textwidth]{source/img/concavity_example.pdf}
+    \caption{Representation of $Q_\epsilon^1$ (blue) and $Q_\epsilon^2$ (yellow)}
+    \label{fig:concavity_example}
+\end{figure}
+
+But for $\oa=(a,\beta_a)$ with $\beta_a = 1+\epsilon/2$, we have:
+\begin{align*}
+    \|\cT Q_\epsilon^1(\os, \oa) - \cT Q_\epsilon^2(\os, \oa)\|_\infty &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)} \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^1_\epsilon)}Q^1_\epsilon(\os',\oa') - \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^2_\epsilon)}Q^2_\epsilon(\os',\oa')\right\|_\infty \\
+    &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)}(\frac{1}{2\epsilon}, 1+\frac{\epsilon}{2}) - (1+\frac{1}{\epsilon},\epsilon )\right\|_\infty \\
+    &= \gamma(1+\frac{1}{2\epsilon})
+\end{align*}
+Hence, $\|\cT Q_\epsilon^1 - \cT Q_\epsilon^2\|_\infty \geq \gamma(1+\frac{1}{2\epsilon})$
+
+As a conclusion, there does not exist $L>0$ such that:
+$$\forall Q_1,Q_2\in(\Real^2)^{\ocS\ocA}, \|\cT Q^1 - \cT Q^2\|_\infty \leq L \|Q^1 - Q^2\|_\infty$$
+In other words, $\cT$ is not a contraction for $\|\cdot\|_\infty$.
+
+\textbf{Restriction to smooth $Q$-functions.} We now study the contractivity of $\cT$ when restricted to the functions of $\cL_\gamma$ defined as follows:
+\begin{equation}
+    \cL_\gamma = \left\{\begin{array}{cc}
+         Q\in(\Real^2)^{\ocS\ocA}\text{ s.t. }\exists L<\frac{1}{\gamma}-1: \forall \os\in\ocS,\oa_1,\oa_2\in\ocA,   \\
+         |Q_r(\os,\oa_1) - Q_r(\os,\oa_2)| \leq L|Q_c(\os,\oa_1) - Q_c(\os,\oa_2)|
+    \end{array}\right\}
+\end{equation}
+That is, for all state $\os$, the set $Q(\os, \ocA)$ plot in the $(Q_c,Q_r)$ plane must be the \emph{graph} of a $L$-Lipschitz function, with $L<1/\gamma-1$.
+
+We impose such structure for the following reason: the counter-example presented above prevented contraction because it was a pathological case in which the slope of $Q$ can be arbitrary large.
+As a consequence, when solving $Q_r^*$ such that $Q_c^*=\beta$, a vertical slice of a $\|\cdot\|_\infty$ ball around $Q_1$ (which must contain $Q_2$) can be arbitrary large as well.
+
+This sketch of proof makes use of insights detailed in the proof of \cref{prop:bftq_pi_hull}, which we recommend the reader to consult first.
+
+We denote $\cB(Q,R)$ the ball of centre $Q$ and radius $R$ for the $\|\cdot\|_\infty$-norm:
+\begin{equation*}
+    \cB(Q,R) = \{Q'\in(R^2)^{\ocS\ocA}: \|Q-Q'\|_\infty \leq R\}
+\end{equation*}
+
+We give the three main steps required to show that $\cT$ restricted to $\cL_\gamma$ is a contraction. Given $Q^1, Q^2\in\cL_\gamma$, show that:
+\begin{enumerate}
+    \item $Q^2\in\cB(Q^1,R)\implies\cF^2\in\cB(\cF^1, R), \forall\os\in\ocS$, where $\cF$ is the top frontier of the convex hull of undominated points, as defined in \cref{sec:proof_pi_hull}.
+    \item $Q\in\cL_\gamma \implies \cF$ is the graph of a $L$-Lipschitz function, $\forall\os\in\ocS$.
+    \item taking the slice $Q_c=\beta$ of a ball $\cB(\cF,R)$ with $\cF$ $L$-Lipschitz results in an interval on $Q_r$ of range at most $(L+1)R$
+\end{enumerate}
+
+These three steps will allow us to control $Q_r^{2*} - Q_r^{1*}$ as a function of $R = \|Q^2-Q^1\|_\infty$.
+
+\textbf{Step 1:} we want to show that if $Q^1$ and $Q^2$ are close, then $\cF^1$ are $\cF^2$ are close as well in the following sense:
+\begin{align}
+    \cF^2\in\cB(\cF^1, R) &\iff d(\cF^1, \cF^2) \leq R \iff \max_{q^2\in\cF^2}\min_{q^1\in\cF^1}\|q^2-q^1\|_\infty \leq R
+    \label{eq:ball-set}
+\end{align}
+Assume $Q^2\in\cB(Q^1,R)$.
+We start by showing this result for $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$ as defined in \cref{sec:proof_pi_hull}:
+
+Let $\os\in\ocS$ and $q^2\in\cC^2(Q^{2-})$, $\exists\lambda\in[0,1], \oa_1,\oa_2\in\ocA: q^2 = (1-\lambda)Q^2(\os,\oa_1) + \lambda Q^2(\os,\oa_2)$.
+Define $q^1 = (1-\lambda)Q^1(\os,\oa_1) + \lambda Q^1(\os,\oa_2)$.
+Then
+\begin{align*}
+    \|q^2-q^1\|_\infty &= \|(1-\lambda)(Q^2(\os,\oa_1) - Q^1(\os,\oa_1)) + \lambda (Q^2(\os,\oa_2) - Q^1(\os,\oa_2))\|_\infty\\
+    &\leq  (1-\lambda)\|Q^2(\os,\oa_1) - Q^1(\os,\oa_1)\|_\infty + \lambda \|Q^2(\os,\oa_2) - Q^1(\os,\oa_2)\|_\infty\\
+    &\leq (1-\lambda)R+\lambda R = R
+\end{align*}
+
+It remains to show that when taking the top frontiers of the convex sets $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$, they remain at a distance of at most $R$.
+
+This is illustrated in \cref{fig:contraction_lips_hull}: given a function $Q^1$, we show the locus $\cB(Q_1,R)$ of $Q^2$.
+We then draw $\cF^1$ the top frontier of the convex hull of $Q^1$ and alongside the locus of all possible $\cF^2$, which belong to a ball $\cB(\cF^1, R)$.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[trim=7cm 4cm 7cm 4cm, clip, width=0.7\textwidth]{source/img/contraction_lipschitz.pdf}
+    \caption{We represent the range of possible solutions $Q_r^{2,*}$ for any $Q^2\in\cB(Q^1)$, given $Q_1\in\cL_\lambda$}
+    \label{fig:contraction_lips_hull}
+\end{figure}
+
+\textbf{Step 2:} We want to show that if $Q\in\cL_\gamma$, $\cF$ is the graph of an $L$-Lipschitz function:
+\begin{equation}
+\label{eq:L-lip-set}
+    \forall q^1,q^2\in\cF, |q_r^2-q_r^1| \leq |q_c^2-q_c^1|
+\end{equation}
+
+Let $Q\in\cL_\gamma$ and $\os\in\ocS$, $\cF$ the corresponding top frontier of convex hull.
+For all $q^1,q^2\in\cF, \exists \lambda,\mu\in[0,1], q^{11},q^{12},q^{21},q^{22}\in Q(\os,\ocA)$ such that $q^1 = (1-\lambda)q^{11} + \lambda q^{12}$ and $q^2 = (1-\mu)q^{21} + \mu q^{22}$.
+Without loss of generality, we can assume $q_c^{11}\leq q_c^{12}$ and $q_c^{21}\leq q_c^{22}$. We also consider the worst case in terms of maximum $q_r$ deviation: $q_c^{12} \leq q_c^{21}$.
+Then the maximum increment $q_r^2-q_r^{1}$ is:
+\begin{align*}
+    \|q^2_r-q^{1}_r\| &\leq \|q^{12}_r-q^{1}_r\| + \|q^{21}_r-q^{12}_r\| + \|q^{2}_r-q^{21}_r\| \\
+    &= (1-\lambda)\|q^{12}_r-q^{11}_r\| + \|q^{21}_r-q^{12}_r\| + \mu\|q^{22}_r-q^{21}_r\| \\ 
+    &\leq (1-\lambda)L\|q^{12}_c-q^{11}_c\| + L\|q^{21}_c-q^{12}_c\| + \mu L\|q^{22}_c-q^{21}_c\| \\
+    &= L\|q^{12}_c-q^{1}_c\| + L\|q^{21}_c-q^{12}_c\| + L\|q^{2}_c-q^{21}_c\|\\
+    &= L\|q^{2}_c-q^{1}_c\|
+\end{align*}
+
+This can also be seen in \cref{fig:contraction_lips_hull}: the maximum slope of the $\cF^1$ is lower than the maximum slope between two points of $Q^1$.
+
+\textbf{Step 3:} Let $\cF_1$ be a L-Lipschitz set as defined in \eqref{eq:L-lip-set}, and consider a ball $\cB(\cF_1,R)$ around it as defined in \eqref{eq:ball-set}.
+
+We want to bound the optimal reward value $Q_r^{2*}$ under constraint $Q_c^{2*} = \beta$ (regular case in \cref{sec:proof_pi_hull} where the constraint is saturated), for any $\cF^2\in\cB(\cF_1,R)$.
+This quantity is represented as a red double-ended arrow in \cref{fig:contraction_lips_hull}.
+
+Because we are only interested in what happens locally at $Q_c=\beta$, we can zoom in on \cref{fig:contraction_lips_hull} and only consider a thin $\epsilon$-section around $\beta$.
+In the limit $\epsilon\rightarrow 0$, this section becomes the tangent to $\cF^1$ at $Q_c^1=\beta$.
+It is represented in \cref{fig:contraction_lips_hull_slope}, from which we derive a geometrical proof:
+\begin{figure}[ht]
+    \centering
+    \includegraphics[trim=2cm 1cm 2cm 1cm, clip, width=0.7\textwidth]{source/img/contraction_lipschitz_slope.pdf}
+    \caption{We represent a section $[\beta-\epsilon, \beta+\epsilon]$ of $\cF^1$ and $B(\cF^1, R)$. We want to bound the range of $Q_r^{2*}.$}
+    \label{fig:contraction_lips_hull_slope}
+\end{figure}
+
+\begin{align*}
+    \Delta Q_r^{2*} &= b + c &\\
+    & \leq La + c & \text{($\cF^1$ $L$-Lipschitz)}\\
+    &= 2LR+2R = 2R(L+1)
+\end{align*}
+Hence,
+\begin{equation*}
+    | Q_r^{2*} - Q_r^{1*}| \leq \frac{\Delta Q_r^{2*}}{2} = R(L+1)
+\end{equation*}
+
+and $Q_c^{1*} = Q_c^{2*} = \beta$.
+Consequently, $ \|Q^{2*} - Q^{1*}\|_\infty \leq (L+1)R$
+
+For completeness, the edge case in \cref{sec:proof_pi_hull} should be considered as well.
+
+\textbf{Wrapping it up:}
+
+We've shown that for any $Q^1,Q^2\in\cL_\gamma$, and all $\os\in\ocS$, $\cF^2\in\cB(\cF^1,\|Q^2-Q^1\|_\infty)$ and $\cF^1$ is the graph of a $L$-Lipschitz function with $L<1/\gamma - 1$.
+Moreover, the solutions of $\pi_\text{greedy}(Q^1)$ and $\pi_\text{greedy}(Q^2)$ at $\os$ are such that $ \|Q^{2*} - Q^{1*}\|_\infty \leq (L+1)\|Q^2-Q^1\|_\infty$.
+
+Hence, for all $\oa$,
+\begin{align*}
+    \|\cT Q^1(\os, \oa) - \cT Q^2(\os, \oa)\|_\infty &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)} \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^1)}Q^1(\os',\oa') - \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^2)}Q^2(\os',\oa')\right\|_\infty \\
+    &= \gamma\left\|Q^{2*} - Q^{1*}\right\|_\infty \\
+    &\leq \gamma(L+1)\|Q^2-Q^1\|_\infty
+\end{align*}
+Taking the sup on $\ocS\ocA$,
+
+\begin{equation*}
+    \|\cT Q^1 - \cT Q^2\|_\infty \leq \gamma(L+1)\|Q^1-Q^2\|_\infty
+\end{equation*}
+with $\gamma(L+1) < 1$.
+As a conclusion, $\cT$ is a $\gamma(L+1)$-contraction on $\cL_\gamma$.
+
 %\subsection{Lemma \ref{lemma:concavity}}
@@
 \subsection{Proposition \ref{prop:bftq_pi_hull}}
+\label{sec:proof_pi_hull}
 \begin{definition}
@@
 
-\section{Experiments}
-
-\subsection{The Lagrangian Relaxation Baseline}
-
+\section{The Lagrangian Relaxation Baseline}
+\label{sec:lagragian}
 As explained on \cref{fig:Lagrangian}, the optimal deterministic policy can be obtained by a line-search on the Lagrange multiplier values $\lambda$.
@@
 
-\subsection{Environments Parameters}
+\section{Experiments}
+
+\subsection{Examples of different exploration strategies}
+\label{subsec:exploration-examples}
+We compare two approaches for constructing a batch of samples.
+The animations from \url{exploration.html} display the trajectories collected in each intermediate sub-batch.
+The first row corresponds to a classical  risk neutral epsilon-greedy exploration policy while the second row showcases a risk-sensitive exploration strategy introduced in the paper.
+Each animation corresponds to a different seed.
+
+\subsection{Examples of BFTQ policies executions}
+\label{sec:bftq-executions}
+
+We display the evolution in the budgeted policy behavior with respect to the budget on different environments. The policies have been learnt with a risk-sensitive exploration.
+
+\paragraph{Highway-Env}
+
+On the \texttt{highway-env} , the budgeted agents display a wide variety of behaviours.
+Animations are displayed on the html page \url{highway-env.html}.
+When $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic.
+On the contrary when $\beta = 0$, the ego-vehicle is conservative: it stays on its lane and drives at a low velocity.
+With intermediate budgets such as $\beta = 0.2$, the agent sometimes decides to overtake its front vehicle but promptly steers back to its original lane afterwards.
+
+\paragraph{Slot-filling}
+
+In \cref{table:dialogues}, we display two dialogues done with the same BFTQ policy on \texttt{Slot-Filling}.
+The policy is given two budgets to respect in expectation, $\beta=0$ and $\beta=0.5$.
+For budget 0, one can see that the system never uses the \texttt{ask\_num\_pad} action.
+Instead, it uses \texttt{ask\_oral} , an action subject to recognition errors.
+The system keeps asking for the same slot 2, because it has the lowest speech recognition score.
+It eventually summarizes the form to the user, but then reaches the maximum dialogue length and thus faces a dialogue failure.
+For budget 0.5, the system first asks in a safe way, with \texttt{ask\_oral}.
+It may want to \texttt{ask\_num\_pad} if one of the speech recognition score is low.
+Then, the system proceeds to a confirmation of the slot values.
+If it is incorrect, the system continues the dialogue using unsafe the \texttt{ask\_num\_pad} action to be certain of the slot values.
+
+\input{source/dialogues.tex}
+
+\paragraph{Corridors}
+
+Animations are displayed on the html page \url{corridors.html} for the \texttt{Corridors} environment.
+When the budget is low, the agent takes the safest path on the left.
+When the budget increases, it gradually switches to the other lane, earning higher rewards but also costs.
+This gradual process could not be achieved with a deterministic policy as it would chose either one path or the other.
+Each animation corresponds to a different seed.
+
+\subsection{Reproducibility}
+
+The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \cref{sec:experiements}.
+
+\subsubsection{Environments Parameters}
 \label{sec:env-parameters}
@@
 
-\subsection{Algorithm parameters}
+\subsubsection{Algorithm parameters}
 \label{sec:algorithms-parameters}
@@
 
-\subsection{Examples of BFTQ policies executions}
-\label{sec:bftq-executions}
+\subsubsection{Instructions for reproducibility}
 
-In \cref{table:dialogues}, we display two dialogues done with the same BFTQ policy.
-The policy is given two budgets to respect in expectation,$\beta=0$ and $\beta=0.5$.
-For budget 0, one can see that the system never uses the \texttt{ask\_num\_pad} action.
-Instead, it uses \texttt{ask\_oral} , an action subject to recognition errors.
-The system keeps asking for the same slot 2, because it has the lowest speech recognition score.
-It eventually summarizes the form to the user, but then reaches the maximum dialogue length and thus faces a dialogue failure.
-For budget 0.5, the system first asks in a safe way, with \texttt{ask\_oral}.
-It may want to \texttt{ask\_num\_pad} if one of the speech recognition score is low.
-Then, the system proceeds to a confirmation of the slot values.
-If it is incorrect, the system continues the dialogue using unsafe the \texttt{ask\_num\_pad} action to be certain of the slot values.
+To reproduce the result displayed in \cref{sec:experiements}, first install the following conventional libraries for python3: pycairo, numpy, scipy and pytorch.
+Then execute the commands recall in \cref{fig:instructions} on a Linux Operating System.
+The Graphic Processing Unit used for experiments is an NVDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}.
 
-\input{source/dialogues.tex}
+\begin{figure}
+    \centering
+
+\begin{minted}{sh}
+# Install highway-env
+pip3 install --user git+https://github.com/eleurent/rl-agents
+# Change python path to the path of this repository
+export PYTHONPATH="code/scaling-up-brl"
+# Navigate to budgeted-rl folder
+cd code/scaling-up-brl/budgeted-rl/
+# Run main script using any config file
+# Choose the range of seeds you want to test on
+python3 main/egreedy/main-egreedy.py config/slot-filling.json 0 6
+python3 main/egreedy/main-egreedy.py config/corridors.json 0 4
+python3 main/egreedy/main-egreedy.py config/highway-easy.json 0 10
+\end{minted}
+\caption{Insctructions to reproduce experiments}
+    \label{fig:instructions}
+\end{figure}
 

### 2019-05-23 source/counter_example.py

@@ -1,0 +1,50 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+# sns.set()
+# sns.set_style("white")
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif', size=15)
+
+eps = 0.7
+beta = 1+eps/2
+
+Q1 = np.array([[1, 1+eps], [0, 1/eps]])
+Q2 = np.array([[0, eps], [1, 1+1/eps]])
+sq1 = np.array([[0, 0], [1, 0], [1, 1], [0, 1], [0, 0]], dtype=np.float32)
+sq2 = sq1.copy()
+sq1[:, 0] += eps
+sq1[:, 1] += 1/eps
+beta_v = np.array([[beta, 0], [beta, 1+1/eps]])
+beta_p = np.array([[beta, 1/(2*eps)]])
+
+fig = plt.figure(figsize=(3, 3.5))
+ax = fig.add_subplot(1, 1, 1)
+
+plt.plot(Q1[0], Q1[1], marker='.', markersize=12)
+plt.plot(Q2[0], Q2[1], marker='.', markersize=12)
+plt.plot(sq1[:, 0], sq1[:, 1], color="grey", linestyle="dashed")
+plt.plot(sq2[:, 0], sq2[:, 1], color="grey", linestyle="dashed")
+# plt.plot(beta_v[:, 0], beta_v[:, 1], color="grey", linestyle="dashed")
+plt.plot(beta_p[:, 0], beta_p[:, 1], color="grey", marker='.', markersize=12)
+# plt.legend(["Q1", "Q2"])
+
+plt.axis('equal')
+
+xticks = [0, 1, beta, 1+eps]
+plt.xticks(xticks, ('0', '1', r'$\beta$', r'$1+\epsilon$'))
+yticks = [1, 1/eps, 1+1/eps]
+plt.yticks(yticks, ('1', r'$\frac{1}{\epsilon}$', r'$1+\frac{1}{\epsilon}$'))
+
+xticks_grid = [beta]
+ax.set_xticks(xticks_grid, minor=True)
+ax.grid(which='minor')
+plt.ylim([0, 1+1/eps])
+# plt.tight_layout(pad=0.4, w_pad=0.4, h_pad=0.)
+plt.subplots_adjust(left=0.3, bottom=0.15)
+
+plt.xlabel(r"$Q_c$")
+plt.ylabel(r"$Q_r$")
+
+plt.savefig("concavity_example.pdf")
+plt.show()

### 2019-05-24 main.tex
Updates from Overleaf

@@
 \usepackage[norelsize, linesnumbered, ruled, lined, boxed, commentsnumbered]{algorithm2e}
+\usepackage{hyperref}
+\usepackage{cleveref}
 \usepackage{float}
@@
 
-\usepackage{hyperref}
-\usepackage{cleveref}
+\crefname{algocf}{alg.}{algs.}
+\Crefname{algocf}{Algorithm}{Algorithms}
+
 \title{Budgeted Reinforcement Learning in Continuous State Space}
@@
     A Budgeted Markov Decision Process (BMDP) is an extension of a Markov Decision Process to critical applications requiring safety constraints.
-It relies on a notion of risk implemented in the shape of a cost signal constrained  to lie below a threshold that -- importantly -- can be modified in real-time.
+It relies on a notion of risk implemented in the shape of a cost signal constrained  to lie below an -- adjustable -- threshold.
+    %A key difference between Constrained Markov Decision Processes (CMDP) and
+    %BMDP lies in the ability to change this threshold in real-time.
+    %TU: on ne comprend pas cette subtilité à ce stade
+    % il vaut mieux développer plus loin.
+    % "that -- importantly -- can be modified in real-time."
     So far, BMDPs could only be solved in the case of finite state spaces with known dynamics.
@@
 \section{Introduction}
+\label{sec:intro}
 
 Reinforcement Learning (RL) is a general framework for decision-making under uncertainty.
-It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R_r, \gamma)$ with continuous state space $\cS$, discrete actions $\cA$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ and rewards $R_r\in\Real^{\cS \times \cA}$, where $\mathcal{M}(\mathcal{X})$ denotes the set of probability measures over a measurable space $\mathcal{X}$.
+It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R_r, \gamma)$ with measurable state space $\cS$, discrete actions $\cA$, unknown rewards $R_r\in\Real^{\cS \times \cA}$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ , where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$.
 Formally, we seek a policy $\pi\in\cM(A)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$.
@@
 In many critical real-world applications where failures may turn out very costly, this is an issue as most decision-makers would rather give away some amount of expected optimality to increase the performances in the lower-tail of the distribution.
-This has led to the development of several risk-averse variants where the optimisation criteria include other statistics of the performance, such as the worst-case realisation \citep{Iyengar2005,Nilim2005,Wiesemann2013}, the variance-penalised expectation \citep{Garcia2015,Tamar2012}, or the Conditional Value-At-Risk (CVaR) \citep{Chow2014,ChowGJP15}.
+This has led to the development of several risk-averse variants where the optimisation criteria include other statistics of the performance, such as the worst-case realisation \citep{Iyengar2005,Nilim2005,Wiesemann2013}, the variance-penalised expectation \citep{Garcia2015,Tamar2012}, the Value-At-Risk (VaR) \citep{Mausser2003,Luenberger2013}, or the Conditional Value-At-Risk (CVaR) \citep{Chow2014,ChowGJP15}.
 
@@
 We equip the MDP with a cost signal $R_c \in \Real^{\cS\times \cA}$ and a cost budget $\beta\in\Real$.
-Similarly to $G_r$, we define the return of costs $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ and the new cost-constrained objective:
+Similarly to $G_r^\pi$, we define the return of costs $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ and the new cost-constrained objective:
 \begin{equation}
@@
 \begin{array}{lcr}
- \displaystyle \max_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s, \pi] & \text{ s.t. } & \expectedvalue[G_c^\pi | s_0=s, \pi] \leq \beta
+ \displaystyle \max_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s] & \text{ s.t. } & \expectedvalue[G_c^\pi | s_0=s] \leq \beta
 \end{array}
@@
 We then propose a novel Budgeted Bellman Optimality Operator and prove the optimal value function to be a fixed point of this operator.
-Second, we use this operator in a Reinforcement Learning algorithm for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure.
+Second, we use this operator in \BFTQ, a batch Reinforcement Learning algorithm, for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure.
 Third, we scale this algorithm to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
 Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.
-The proofs of our main results are provided in Supplementary Material \ref{sec:proofs}.
+The proofs of our main results are provided \Cref{sec:proofs}.
 
 \section{Budgeted Dynamic Programming}
-
+\label{sec:bdp}
  We work in the space of budgeted policies, where $\pi$ both depends on $\beta$ and also outputs a next budget $\beta_a$.
@@
 
-For ease of notation, we consider the \emph{augmented} state and action spaces $\ocS = \cS\times \cB$ and $\ocA= \cA\times \cB$, and equip them with the augmented dynamics $\ov{P}\in \cM(\ocS)^{\ocS \times \ocA}$ defined as:
+We cast the BMDP problem as a multi-objective MDP problem \citep{Roijers2013ASO} by considering \emph{augmented} state and action spaces $\ocS = \cS\times \cB$ and $\ocA= \cA\times \cB$, and equip them with the augmented dynamics $\ov{P}\in \cM(\ocS)^{\ocS \times \ocA}$ defined as:
 \begin{equation}
@@
 \end{equation}
-where $\delta$ is the Dirac distribution.
+where $\delta$ is the Dirac indicator distribution.
+% TU: on a rappelé ce que c'est qu'un MDP, on pourrait peut-être rappeler ce qu'est un Dirac ?
+% TU: si je comprend bien \beta_a = \beta' c'est le budget restant et ça n'es tplus le delta de budget. La relation avec \beta reste très "implicite".
 
@@
     \label{eq:value-function}
-V^\pi(\os) = (V_r^\pi, V_c^\pi) = \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa)= (Q_r^\pi, Q_c^\pi) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right].
+V^\pi(\os) = (V_r^\pi, V_c^\pi) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa)= (Q_r^\pi, Q_c^\pi) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right].
 \end{equation}
-We restrict $\ocS$ to feasible budgets only: $\ocS_f \eqdef \{(s,\beta)\in\ocS: \exists \pi\in\Pi, V_c^\pi(s) \geq \beta\}$, but still write $\ocS$ in place of $\ocS_f$ for brevity of notations.
+We restrict $\ocS$ to feasible budgets only: $\ocS_f \eqdef \{(s,\beta)\in\ocS: \exists \pi\in\Pi, V_c^\pi(s) \geq \beta\}$ that we assume is non-empty for the BMDP to admit a solution.
+We still write $\ocS$ in place of $\ocS_f$ for brevity of notations.
 
@@
 \label{prop:bellman-expectation}
-The value functions $V^\pi$ and $Q^\pi$ verify:
+The value functions $V^\pi$ and $ Q^\pi$ verify:
 \begin{align}
@@
     \begin{equation}
-        V_r^*(\os), \eqdef \max_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\qquad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
+        V_r^*(\os) \eqdef \max_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\qquad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
     \end{equation}
@@
 \label{eq:bellman-optimality}
-    Q^{*}(\os, \oa) = \cT Q^{*}(\os, \oa) \eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'}),
+    Q^{*}(\os, \oa) = \cT Q^{*}(\os, \oa) \eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in \ocA} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'}),
 \end{equation}
@@
 \label{thm:contraction}
-If $|\cA| \geq 2$, then $\cT$ is not a contraction.\\
-However, we conjecture that $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
-We give some intuition of why that should be the case in Supplementary Material \ref{sec:proof_contraction}.
+If $|\cA| \geq 2$, then $\cT$ is not a contraction.
 \end{theorem}
@@
 \paragraph{Budgeted Value Iteration}
-{
+
 The Budgeted Bellman Optimality equation is a fixed-point equation, which motivates the introduction of a fixed-point iteration procedure.
-We introduce Algorithm~\ref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.
+We introduce \Cref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.
 
 Unfortunately, as $\cT$ is not a contraction, we can guarantee neither the convergence of this procedure nor the unicity of its fixed points.
-However, we empirically observed the convergence in our experiments (\Cref{sec:experiements}).
-}
+Despite those theoretical limitations, we empirically observed the convergence to a fixed point in our experiments (\Cref{sec:experiements}).
+We conjecture a possible explanation:
 
-Because of the budget dynamics $\beta' = \beta_a$ in \eqref{eq:dynamics}, we can see in \eqref{eq:bellman_expectation} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
+\begin{remark}[Contractivity of $\cT$ on smooth $Q$-functions]
+\label{rmk:contractivity-smooth}
+We conjecture that $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
+We lengthily discuss some intuition on why that should be the case in \Cref{proof:contraction-with-smooth}.
+\end{remark}
+
+% Because of the budget dynamics $\beta' = \beta_a$ in \eqref{eq:dynamics}, we can see in \eqref{eq:bellman_expectation} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
 We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
@@
 \section{Budgeted Reinforcement Learning}
+\label{sec:brl}
 
@@
 \subsection{Budgeted Fitted-Q}
+\label{subsec:bftq}
 
-When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$ collected by interaction with the environment.
+When the BMDP is unknown, we need to adapt \Cref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$ collected by interaction with the environment.
 Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$.
@@
 \end{equation*}
-We introduce in Algorithm~\ref{algo:bftq} the \emph{Budgeted-Fitted-Q} (BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs.
+We introduce in \Cref{algo:bftq} the \emph{Budgeted-Fitted-Q} (BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs.
 Because we work with  continuous state space $\cS$ and budget space $\cB$, we need to employ function-approximation in order to generalise to nearby states and budgets.
@@
 \vspace{0pt}
+
 \begin{algorithm}[H]
-\label{algo:bvi}
 \DontPrintSemicolon
@@
 \caption{Budgeted Value Iteration}
+\label{algo:bvi}
+
 \end{algorithm}
+
 \end{minipage}%
@@
 \vspace{0pt}
+
 \begin{algorithm}[H]
-\label{algo:bftq}
 \DontPrintSemicolon
@@
 \caption{Budgeted Fitted-Q}
+\label{algo:bftq}
+
 \end{algorithm}
+
 \end{minipage}
@@
 \subsection{Risk-sensitive exploration}
-
 \label{sec:exploration}
 
-In order to run Algorithm~\ref{algo:bftq}, we must first gather a batch of samples $\mathcal{D}$.
+In order to run \Cref{algo:bftq}, we must first gather a batch of samples $\mathcal{D}$.
 Ideally we would need samples from the asymptotic state-budget distribution $\lim_{t\rightarrow\infty}\probability{\os_t}$ induced by an optimal policy $\pi^*$ given an initial distribution $\probability{\os_0}$, but as we are actually building this policy, it is not possible.
@@
 We follow an exploration policy: a mixture between a random budgeted policy $\pi_\text{rand}$ and the current greedy policy $\pi_\text{greedy}$.
-The batch $\cD$ is split into several mini-batches generated sequentially, and $\pi_\text{greedy}$ is updated by running Algorithm~\ref{algo:bftq} on $\cD$ upon mini-batch completion.
-$\pi_\text{rand}$ is designed to obtain trajectories that only explore feasible budgets: we impose that the joint distribution $\probability{a, \beta_a}$ verifies $\expectedvalue[\beta_a]\leq\beta$.
+The batch $\cD$ is split into several mini-batches generated sequentially, and $\pi_\text{greedy}$ is updated by running \Cref{algo:bftq} on $\cD$ upon mini-batch completion.
+$\pi_\text{rand}$ is designed to obtain trajectories that only explore feasible budgets: we impose that the joint distribution $\probability{a, \beta_a|s, \beta}$ verifies $\expectedvalue[\beta_a]\leq\beta$.
 This condition defines a probability simplex $\Delta_{\ocA}$ from which we sample uniformly.
@@
 Importantly, we pick a uniform distribution $\probability{\beta_0} = \cU(\cB)$ so that the entire range of risk-level is explored, and not only reward-seeking behaviours as would be the case with a traditional risk-neutral $\epsilon$-greedy strategy.
-The pseudo-code of our exploration procedure is shown in Algorithm~\ref{algo:risk-sensitive-exploration} in the Supplementary Material.
+The pseudo-code of our exploration procedure is shown in \Cref{algo:risk-sensitive-exploration} in \Cref{sec:risk-sensitive-supp}.
 
 \section{A Scalable Implementation}
-
 \label{sec:scalable-bftq}
@@
 \subsection{How to compute the greedy policy?}
-
-As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested constrained optimisation program \eqref{eq:pi_greedy}.
+\label{subsec:compute-greedy-policy}
+As stated in \Cref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested constrained optimisation program \eqref{eq:pi_greedy}.
 % In order to apply $\cT$ in \eqref{eq:bellman-optimality}, it is necessary to compute the greedy policy $\pi_\text{greedy}(s; Q)$.
@@
 \label{prop:bftq_pi_hull}
-Algorithm~\ref{algo:bvi} and Algorithm~\ref{algo:bftq} can be run by replacing $\pi_\text{greedy}$ in the equation \eqref{eq:bellman-optimality} of $\cT$ with $\pi_\text{hull}$ as described in Algorithm~\ref{algo:pi_hull}.
+\Cref{algo:bvi} and \Cref{algo:bftq} can be run by replacing $\pi_\text{greedy}$ in the equation \eqref{eq:bellman-optimality} of $\cT$ with $\pi_\text{hull}$ as described in \Cref{algo:pi_hull}.
 \begin{equation*}
@@
 \DontPrintSemicolon
-\KwData{$s$, $Q$}
-$Q^+\leftarrow \{Q_c > \min \{Q_c(s,\oa) \text{ s.t. }\oa\in\argmax_{\oa} Q_r(s,\oa)\} \}$\tcp*[f]{dominated points}\;
-$\cF \leftarrow \text{top frontier of }\texttt{convex\_hull}(Q(s,\ocA) \setminus Q^+)$\tcp*[f]{candidate mixtures}\;
-$\cF_Q \leftarrow \cF\cap Q(s,\ocA)$\;
-\For{points $q = Q(s,\oa)\in\cF_Q$ in clockwise order}{
+\KwData{$\os=(s,\beta)$, $Q$}
+$Q^+\leftarrow \{Q_c > \min \{Q_c(\os,\oa) \text{ s.t. }\oa\in\argmax_{\oa} Q_r(\os,\oa)\} \}$\tcp*[f]{dominated points}\;
+$\cF \leftarrow \text{top frontier of }\texttt{convex\_hull}(Q(\os,\ocA) \setminus Q^+)$\tcp*[f]{candidate mixtures}\;
+$\cF_Q \leftarrow \cF\cap Q(\os,\ocA)$\;
+\For{points $q = Q(\os,\oa)\in\cF_Q$ in clockwise order}{
 \uIf{find two successive points $((q_c^1, q_r^1), (q_c^2, q_r^2))$ of $\cF_Q$ such that $q_c^1 \leq \beta < q_c^2$}{
@@
 }}
-\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(s,\oa))$\tcp*[f]{Budget $\beta$ always respected}}
+\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(\os,\oa))$\tcp*[f]{Budget $\beta$ always respected}}
 \caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
@@
 
-The computation of $\pi_\text{hull}$ in Alg.~\ref{algo:pi_hull} is illustrated in Fig.~\ref{fig:hull}.
+The computation of $\pi_\text{hull}$ in \Cref{algo:pi_hull} is illustrated in \Cref{fig:hull}.
 
@@
 A straightforward choice is a single layer with the same width as the state.
-The overall architecture is shown in Supplementary Material Fig.~\ref{fig:architecture}.
+The overall architecture is shown in \Cref{fig:architecture} in \Cref{sec:bftq-full}.
 
 \subsection{Parallel computing}
-
-In a simulated environment, a first process that can be distributed is the collection of samples in the exploration procedure of Algorithm~\ref{algo:risk-sensitive-exploration}, as $\pi_\text{greedy}$ stays constant within each mini-batch which avoids the need of synchronisation between workers.
+\label{subsec:parallel-computing}
+In a simulated environment, a first process that can be distributed is the collection of samples in the exploration procedure of \Cref{algo:risk-sensitive-exploration}, as $\pi_\text{greedy}$ stays constant within each mini-batch which avoids the need of synchronisation between workers.
 Second, the main bottleneck of \BFTQ is the computation of the target $\cT Q$.
@@
 This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA,\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass.
-By using multiprocessing in the computations of $\pi_\text{hull}$, we enjoy a linear speedup.% and reduce the time-complexity to $O(|\mathcal{D}|/(W(1-\gamma)) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$.
-The full description of our scalable implementation of \BFTQ is recalled in Algorithm~\ref{algo:bftq_full} in Supplementary Material~\ref{sec:bftq-full}.
+By using multiprocessing in the computations of $\pi_\text{hull}$, we enjoy a linear speedup.
+The full description of our scalable implementation of \BFTQ is recalled in \Cref{algo:bftq_full} in \Cref{sec:bftq-full}.
 
@@
 
-\paragraph{Exploration strategies} We claimed in Section~\ref{sec:exploration} that a risk-sensitive exploration was required in the setting of BMDPs.
+\paragraph{Exploration strategies}\label{par:ex-explo} We claimed in \Cref{sec:exploration} that a risk-sensitive exploration was required in the setting of BMDPs.
 We test this hypotheses by confronting our strategy to a classical risk-neutral strategy.
@@
 
-\paragraph{Budgeted algorithms} We compare our  scalable \BFTQ algorithm described in Section~\ref{sec:scalable-bftq} to an \FTQl baseline.
+\paragraph{Budgeted algorithms}\label{par:ex-brl} We compare our  scalable \BFTQ algorithm described in \Cref{sec:scalable-bftq} to an \FTQl baseline.
 This baseline consists in approximating the BMDP by a finite set of CMDPs problems.
@@
 Each run was repeated $N_{\text{seeds}}$ times.
-Parameters of the algorithms can be found in Supplementary Material~\ref{sec:algorithms-parameters}
+Parameters of the algorithms can be found in \Cref{sec:algorithms-parameters}
 
 \subsection{Environments}
-
-We evaluate our method on three different environments involving reward-cost trade-offs. Their parameters can be found in Supplementary Material~\ref{sec:env-parameters}
+\label{subsec:environments}
+We evaluate our method on three different environments involving reward-cost trade-offs. Their parameters can be found in \Cref{sec:env-parameters}
 
 \paragraph{Corridors}
-This simplistic environment is only meant to highlight clearly the specificity of exploration in a budgeted setting.
+This simple environment is only meant to highlight clearly the specificity of exploration in a budgeted setting.
 It is a continuous gridworld with Gaussian perturbations, consisting in a maze composed of two corridors: a risky one with high rewards and costs, and a safe one with low rewards and no cost.
@@
 \paragraph{Spoken dialogue system}
-Our second application is a slot-filling simulation \citep{Li2009ReinforcementLF,chandramohan2010optimizing}.
-The system fills in a form of slot-values by interacting with a user through speech, before sending them a response.
-For example, in the restaurant domain, it may ask for three slots : the area of the restaurant, the price-range and the food type.
+Our second application is a dialogue-based slot-filling simulation that has already benefited from batch RL optimisation in the past~\citep{Li2009ReinforcementLF,chandramohan2010optimizing,pietquin2011sample}.
+The system fills in a form of slot-values by interacting a user through speech, before sending them a response.
+For example, in a restaurant reservation domain, it may ask for three slots: the area of the restaurant, the price-range and the food type.
 The user could respectively provide those three slot-values : \texttt{Cambridge}, \texttt{Cheap} and \texttt{Indian-food}.
@@
 To that end, the system can choose among a set of generic actions.
-As in \citep{carrara2018safe} there are two ways of asking for a slot value: a slot value can be either be provided with an utterance, which may to speech recognition errors with some probability, or by requiring the user to fill-in the slots by using a numeric pad.
-In this case, there are no recognition errors but a counterparty risk of hang-up: we assume that manually filling a key-value form is time-consuming and annoying.
+As in \citep{carrara2018safe}, there are two ways of asking for a slot value: a slot value can be either be provided with an utterance, which may cause speech recognition errors with some probability, or by requiring the user to fill-in the slots by using a numeric pad.
+In this case, there are no recognition errors but a counterpart risk of hang-up: we assume that manually filling a key-value form is time-consuming and annoying.
 The environment yields a reward if all slots are filled without errors, and a constraint if the user hang-ups.
 Thus, there is a clear trade-off between using utterances and potentially committing a mistake, or using the numeric pad and risking a premature hang-up.
-% Note that batch reinforcement learning has already been applied to spoken dialogue systems in \citep{Li2009ReinforcementLF,chandramohan2010optimizing} which makes this task a realistic candidate for optimisation with our method.
 
@@
 In our third application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{Leurent2018} for simulated highway driving and behavioural decision-making.
-We propose a novel task that displays a clear trade-off between safety and efficiency.
+We define a task that displays a clear trade-off between safety and efficiency.
 The agent controls a vehicle with a finite set of manoeuvres implemented by low-lever controllers: $\mathcal{A}$ = \{\text{no-op}, \text{right-lane}, \text{left-lane}, \text{faster}, \text{slower}\}.
@@
 \subsection{Results}
-
+\label{subsec:results}
 In the following figures, each patch represents the mean and 95\% confidence interval over $N_{\text{seeds}}$ seeds of the means of $(G_r^\pi,G_c^\pi)$ over $N_\text{trajs}$ trajectories.
@@
 
-We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: Figure~\ref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in supplementary material ~\ref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch.
-The trajectories (purple) in the risk-neutral batch (left) are concentrated along the risky corridor (orange) and ignore the safe corridor (green), which results in bad performances in the low-risk regime.
-Conversely, trajectories in the risk-sensitive batch (right) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.
+We first bring to light the role of risk-sensitive exploration in the \texttt{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in \Cref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch.
+The trajectories (orange) in the risk-neutral batch are concentrated along the risky corridor (red) and ignore the safe corridor (green), which results in bad performances in the low-risk regime.
+Conversely, trajectories in the risk-sensitive batch (blue) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.
 
@@
     \includegraphics[page=1, width=0.45\textwidth]{source/img/corridors}
-    \caption{Trajectories (left) and performances (right) of two exploration strategies in the \text{corridors} environment. }
+    \caption{Trajectories (left) and performances (right) of two exploration strategies in the \texttt{corridors} environment. }
     \label{fig:exploration}
@@
     \includegraphics[width=0.49\linewidth]{source/img/highway}
-    \caption{Performance comparison of \FTQl and BFTQ on \text{slot-filling} (left) and \text{highway-env}(right) }
+    \caption{Performance comparison of \FTQl and BFTQ on \texttt{slot-filling} (left) and \texttt{highway-env}(right) }
     \label{fig:results}
@@
 
-In a second experiment displayed in Fig.~\ref{fig:results}, we compare the performance of \FTQl to that of \BFTQ in the slot-filling and autonomous driving tasks.
+In a second experiment displayed in \Cref{fig:results}, we compare the performance of \FTQl to that of \BFTQ in the dialogue and autonomous driving tasks.
 For each algorithm, we plot the reward-cost trade-off curve.
 In both cases, BFTQ performs almost as well as \FTQl despite only requiring a single model.
-All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can observe in the bottom-left corner an underestimation of $Q_c$, since $\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
-%This underestimation can be a consequence of two approximations: the use of the sampling operator $\hat{\cT}$ instead of the true environmental operator $\cT$, and the use of the neural network function approximation $Q_\theta$ instead of $Q$.
-This is probably due to the fact that \text{highway-env} is actually infeasible for $\beta=0$, since we noticed that about $8\%$ of bad initialisations occur with the front vehicle driving slower than the minimum velocity of the controlled vehicle, leading to either a front collision or a forced change to the opposite lane, both yielding costs.
+All budgets are well-respected on \texttt{slot-filling}, but on \texttt{highway-env} we can observe an underestimation of $Q_c$, since e.g.
+$\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
+This underestimation can be a consequence of two approximations: the use of the sampling operator $\hat{\cT}$ instead of the true environmental operator $\cT$, and the use of the neural network function approximation $Q_\theta$ instead of $Q$.
+% This is probably due to \text{highway-env} being actually infeasible for $\beta=0$, since we noticed that about $8\%$ of bad initialisations occur with the front vehicle driving slower than the minimum velocity of the controlled vehicle, leading to either a front collision or a forced change to the opposite lane, both yielding costs.
 Still, \BFTQ provides a better control on the expected cost of the policy, than \FTQl.
 In addition, \BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance.
-Examples of policy execution can be found in Supplementary Material~\ref{sec:bftq-executions}.
+Examples of policy execution can be found in \Cref{sec:bftq-executions}.
 
 \section{Discussion}
-
-Algorithm~\ref{algo:bftq} is an algorithm for solving large unknown BMDPs with continuous states.
-To the best of our knowledge, there is no algorithm in the current literature that combines all those features.
+\label{subsec:discussions}
+\Cref{algo:bftq} is an algorithm for solving large unknown BMDPs with continuous states. To the best of our knowledge, there is no algorithm in the current literature that combines all those features.
 
@@
 When the environment parameters ($P$, $R_r$, $R_c$) are known but not tractable, solutions relying on function approximation~\citep{Undurti} or approximate linear programming~\citep{Poupart2015} have been proposed.
-For unknown environments, batch algorithms exist in the tabular setting \citep{Thomas2015, Petrik2016, Laroche2019}.
-Online algorithms \citep{Geibel2005, Abe2010,ChowGJP15,AchiamHTA17} and a batch algorithm \citep{le2019batch} combine the best of both world by solving large unknown CMDPs.
+For unknown environments, online algorithms \citep{Geibel2005, Abe2010,ChowGJP15,AchiamHTA17} and a batch algorithm \citep{Thomas2015, Petrik2016, Laroche2019,le2019batch} can solve large unknown CMDPs.
+Nevertheless, these approaches are limited in that the constraints thresholds are fixed prior to training and cannot be updated in real-time at policy execution to select the desired level of risk.
 
-A first way of solving a BMDP is to approximate it with a finite set of CMDPs.
-The solutions of these CMDPs take the form of mixtures between two deterministic policies\citep[Theorem 4.4,][]{BEUTLER1985236}.
-To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material ~\ref{sec:lagragian}}.
+To our knowledge, there were only two ways of solving a BMDP.
+The first one is to approximate it with a finite set of CMDPs (e.g.
+see our \FTQl baseline).
+The solutions of these CMDPs take the form of mixtures between two deterministic policies \citep[Theorem 4.4,][]{BEUTLER1985236}.
+To obtain these policies, one needs to evaluate their expected cost by interacting with the environment\footnote{More details are provided in \Cref{sec:lagragian}}.
 Our solution not only requires one single model but also avoids any supplementary interaction.
 
-The closest work to ours is the Dynamic Programming algorithm proposed by \citet{Boutilier_Lu:uai16} for solving BDMPs.
+The only other existing BMDP algorithm, and closest work to ours, is the Dynamic Programming algorithm proposed by \citet{Boutilier_Lu:uai16}.
 However, their work was established for finite state spaces only, and their solution relies heavily on this property.
@@
 \section{Conclusion}
+\label{sec:conclusion}
 The BMDP framework is a principled framework for safe decision making under uncertainty, which could be beneficial to the diffusion of Reinforcement Learning in industrial applications.
@@
 In order to scale to large problems, we provide an efficient implementation that exploits the structure of the value function and leverages tools from Deep Distributed Reinforcement Learning.
-We show that on two practical tasks, our solution performs similarly to a baseline Lagrangian relaxation method while only requiring a single model to train.
+We show that on two practical tasks our solution performs similarly to a baseline Lagrangian relaxation method while only requiring a single model to train, and relying on an interpretable $\beta$ instead of the tedious tuning of the penalty $\lambda$.
 

### 2019-05-24 supplementary.tex
Updates from Overleaf

@@
-% \documentclass{article}
-
-% % if you need to pass options to natbib, use, e.g.:
-% %     \PassOptionsToPackage{numbers, compress}{natbib}
-% % before loading neurips_2019
-
-% % ready for submission
-% % \usepackage{neurips_2019}
-
-% % to compile a preprint version, e.g., for submission to arXiv, add add the
-% % [preprint] option:
-% %     \usepackage[preprint]{neurips_2019}
-
-% % to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage{neurips_2019}
-
-% % to avoid loading the natbib package, add option nonatbib:
-% %     \usepackage[nonatbib]{neurips_2019}
-
-% \usepackage[utf8]{inputenc} % allow utf-8 input
-% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-% \usepackage{url}            % simple URL typesetting
-% \usepackage{booktabs}       % professional-quality tables
-% \usepackage{amsfonts}       % blackboard math symbols
-% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
-% \usepackage{microtype}      % microtypography
-
-% % Paper specific includes and commands
-% \usepackage{times}
-% \usepackage{soul}
-% \usepackage[small]{caption}
-% \usepackage{graphicx}
-% \usepackage{amsmath}
-% \usepackage{amssymb}
-% \usepackage{amsfonts}
-% \usepackage{amsthm}
-% \usepackage[shortlabels]{enumitem}
-% \usepackage{algorithm}
-% \usepackage{algorithmic}
-% \usepackage{tabularx}
-% \usepackage{makecell}
-% \urlstyle{same}
-% \usepackage{xargs}
-% \usepackage{mathtools}
-% \usepackage{dsfont}
-% \usepackage{import}
-% \usepackage{svg}
-% \usepackage{pdfpages}
-
-% \input{mathdef}
-
-% \makeatletter
-% \providecommand*{\input@path}{}
-% \g@addto@macro\input@path{{./source/}{./source/ressources/}{./ressources/}}
-% \makeatother
-
-% \urlstyle{same}
-
-% \usepackage{wrapfig}
-% \usepackage{xr-hyper}
-% \usepackage{hyperref}
-% \usepackage[capitalise,noabbrev]{cleveref}
-
-% % the following package is optional:
-% %\usepackage{latexsym}
-
-% \title{Supplementary material}
-
-% \author{
-% Nicolas Carrara$^{1,}$\thanks{equal contribution}\and
-% Edouard Leurent$^{1,3,}$\footnotemark[1]\and
-% Romain Laroche$^{4}$\and
-% Tanguy Urvoy$^{2}$\and
-% Odalric Maillard$^{1}$\and
-% Olivier Pietquin$^{1,2,}$\thanks{now with Google Brain, Paris.}\\
-% \affiliations $^1$SequeL team, INRIA Lille -- Nord Europe, France\\
-% $^2$Orange Labs, Lannion, France\\
-% $^3$Renault Group, France\\
-% $^4$Microsoft Research , Montreal, Canada
-% \emails
-% \{nicolas.carrara, edouard.leurent, odalric.maillard, olivier.pietquin\}@inria.fr,
-% romain.laroche@microsoft.com, tanguy.urvoy@orange.com
-% }
-
-% \begin{document}
-% \maketitle
-
 \begin{center}
-\LARGE Supplementary material
+\LARGE Appendix
 \end{center}
@@
 
+\paragraph{Outline}
+
+This paper gathers all the supplementary material and goes as follows: \Cref{sec:proofs} details all the proofs of the main results.
+\Cref{sec:risk-sensitive-supp} and \Cref{sec:bftq-full} recall respectively the scalable \BFTQ algorithm and the risk-sensitive exploration procedure.
+\Cref{sec:lagragian} describes a naive alternative to \BFTQ based on Lagrangian Relaxation.
+The \Cref{sec:exp-supp} assembles all the assets for visualising and reproducing the experiments, including visualisations of policy executions, algorithms and environment parameters, and instructions for executing the attached source code.
+Finally we fill the Machine Learning Reproducibility Checklist and we justify each statement in \Cref{sec:ml-checklist}.
+
 \section{Proofs of Main Results}
 \label{sec:proofs}
-\subsection{Proposition \ref{prop:bellman-expectation}}
+\subsection{\Cref{prop:bellman-expectation}}
 
 \begin{proof}
+This proof is the same as that in classical multi-objective MDPs.
+
 \begin{align*}
-    V^\pi(\os) &= \expectedvalue\left[ J^\pi \condbar \ov{s_0} = \os\right] \\
-    &=\sum_{\oa\in\ocA} \probability{\oa_0 = \oa \condbar\ov{s_0} = \os} \expectedvalue\left[ J^\pi \condbar \ov{s_0} = \os, \oa_0 = \oa\right]\\
+    V^\pi(\os) &\eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \\
+    &=\sum_{\oa\in\ocA} \probability{\oa_0 = \oa \condbar\ov{s_0} = \os} \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \oa_0 = \oa\right]\\
     &= \sum_{\oa\in\ocA} \pi(\oa | \os) Q^\pi(\os,\oa)
@@
 \begin{align*}
-    Q^\pi(\os, \oa) &= \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t)\condbar \ov{s_0} = \os, \ov{a_0} = \oa\right] \\
-    &= R(\os, \oa) + \sum_{\os'\in\ocS}\probability{\os_1 = \os' \condbar\ov{s_0} = \os, \ov{a_0} = \oa}\expectedvalue\left[\sum_{t=1}^\infty \gamma^t R(\os_t, \oa_t)\condbar \ov{s_1} = \os'\right] \\
+    Q^\pi(\os, \oa) &\eqdef \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t)\condbar \ov{s_0} = \os, \ov{a_0} = \oa\right] \\
+    &= R(\os, \oa) + \sum_{\os'\in\ocS}\probability{\os_1 = \os' \condbar\ov{s_0} = \os, \ov{a_0} = \oa}\cdot \expectedvalue\left[\sum_{t=1}^\infty \gamma^t R(\os_t, \oa_t)\condbar \ov{s_1} = \os'\right] \\
     &= R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar\os, \oa\right) \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t) \condbar \ov{s_0} = \os'\right] \\
@@
 
-\subsection{Theorem \ref{thm:bellman-optimality}}
+\subsection{\Cref{thm:bellman-optimality}}
 
@@
 \item[\eqref{eq:pthm_def}.] Definition of $Q^*$. 
-\item[\eqref{eq:pthm_exp}.] Bellman Expectation expansion from Proposition \ref{prop:bellman-expectation}.
+\item[\eqref{eq:pthm_exp}.] Bellman Expectation expansion from \Cref{prop:bellman-expectation}.
 \item[\eqref{eq:pthm_marg}.] Marginalisation on $\oa'$.
@@
 
-\subsection{Proposition \ref{prop:greedy_optimal}}
+\subsection{\Cref{prop:greedy_optimal}}
 \begin{proof}
 Notice from the definitions of $\cT$ and $\cT^\pi$ in \eqref{eq:bellman-optimality} and \eqref{eq:bellman_expectation_operator} that $\cT$ and $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$ coincide on $Q^*$.
-Moreover, since $Q^* = \cT Q^*$ by \cref{thm:bellman-optimality}, we have: $    \cT^{\pi_\text{greedy}(\cdot;Q^*)} Q^* = \cT Q^* = Q^*
+Moreover, since $Q^* = \cT Q^*$ by \Cref{thm:bellman-optimality}, we have: $    \cT^{\pi_\text{greedy}(\cdot;Q^*)} Q^* = \cT Q^* = Q^*
 $.
-Hence, $Q^*$ is a fixed point of $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$, and by \cref{prop:bellman-expectation} it must be equal to $Q^{\pi_\text{greedy}(\cdot;Q^*)}$
+Hence, $Q^*$ is a fixed point of $\cT^{\pi_\text{greedy}(\cdot;Q^*)}$, and by \Cref{prop:bellman-expectation} it must be equal to $Q^{\pi_\text{greedy}(\cdot;Q^*)}$
 
@@
 
-\subsection{Theorem \ref{thm:contraction}}
+\subsection{\Cref{thm:contraction}}
 \label{sec:proof_contraction}
-In the trivial case $|\cA| = 1$, there exits only one policy $\pi$ and $\cT = \cT^\pi$, which is a contraction by \cref{prop:bellman-expectation}.
+\begin{proof}
+In the trivial case $|\cA| = 1$, there exits only one policy $\pi$ and $\cT = \cT^\pi$, which is a contraction by \Cref{prop:bellman-expectation}.
 
@@
 Then, $\|Q_1-Q_2\|_\infty = 1$.
- $Q_\epsilon^1$ and $Q_\epsilon^2$ are represented in \cref{fig:concavity_example}.
+ $Q_\epsilon^1$ and $Q_\epsilon^2$ are represented in \Cref{fig:concavity_example}.
 
@@
 In other words, $\cT$ is not a contraction for $\|\cdot\|_\infty$.
+\end{proof}
 
-\textbf{Restriction to smooth $Q$-functions.} We now study the contractivity of $\cT$ when restricted to the functions of $\cL_\gamma$ defined as follows:
+\subsection{\Cref{rmk:contractivity-smooth}}
+\label{proof:contraction-with-smooth}
+\begin{proof}
+We now study the contractivity of $\cT$ when restricted to the functions of $\cL_\gamma$ defined as follows:
 \begin{equation}
@@
 
-This sketch of proof makes use of insights detailed in the proof of \cref{prop:bftq_pi_hull}, which we recommend the reader to consult first.
+This sketch of proof makes use of insights detailed in the proof of \Cref{prop:bftq_pi_hull}, which we recommend the reader to consult first.
 
@@
 \begin{enumerate}
-    \item $Q^2\in\cB(Q^1,R)\implies\cF^2\in\cB(\cF^1, R), \forall\os\in\ocS$, where $\cF$ is the top frontier of the convex hull of undominated points, as defined in \cref{sec:proof_pi_hull}.
+    \item $Q^2\in\cB(Q^1,R)\implies\cF^2\in\cB(\cF^1, R), \forall\os\in\ocS$, where $\cF$ is the top frontier of the convex hull of undominated points, as defined in \Cref{sec:proof_pi_hull}.
     \item $Q\in\cL_\gamma \implies \cF$ is the graph of a $L$-Lipschitz function, $\forall\os\in\ocS$.
@@
 Assume $Q^2\in\cB(Q^1,R)$.
-We start by showing this result for $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$ as defined in \cref{sec:proof_pi_hull}:
+We start by showing this result for $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$ as defined in \Cref{sec:proof_pi_hull}:
 
@@
 
-This is illustrated in \cref{fig:contraction_lips_hull}: given a function $Q^1$, we show the locus $\cB(Q_1,R)$ of $Q^2$.
+This is illustrated in \Cref{fig:contraction_lips_hull}: given a function $Q^1$, we show the locus $\cB(Q_1,R)$ of $Q^2$.
 We then draw $\cF^1$ the top frontier of the convex hull of $Q^1$ and alongside the locus of all possible $\cF^2$, which belong to a ball $\cB(\cF^1, R)$.
@@
 
-This can also be seen in \cref{fig:contraction_lips_hull}: the maximum slope of the $\cF^1$ is lower than the maximum slope between two points of $Q^1$.
+This can also be seen in \Cref{fig:contraction_lips_hull}: the maximum slope of the $\cF^1$ is lower than the maximum slope between two points of $Q^1$.
 
@@
 
-We want to bound the optimal reward value $Q_r^{2*}$ under constraint $Q_c^{2*} = \beta$ (regular case in \cref{sec:proof_pi_hull} where the constraint is saturated), for any $\cF^2\in\cB(\cF_1,R)$.
-This quantity is represented as a red double-ended arrow in \cref{fig:contraction_lips_hull}.
+We want to bound the optimal reward value $Q_r^{2*}$ under constraint $Q_c^{2*} = \beta$ (regular case in \Cref{sec:proof_pi_hull} where the constraint is saturated), for any $\cF^2\in\cB(\cF_1,R)$.
+This quantity is represented as a red double-ended arrow in \Cref{fig:contraction_lips_hull}.
 
-Because we are only interested in what happens locally at $Q_c=\beta$, we can zoom in on \cref{fig:contraction_lips_hull} and only consider a thin $\epsilon$-section around $\beta$.
+Because we are only interested in what happens locally at $Q_c=\beta$, we can zoom in on \Cref{fig:contraction_lips_hull} and only consider a thin $\epsilon$-section around $\beta$.
 In the limit $\epsilon\rightarrow 0$, this section becomes the tangent to $\cF^1$ at $Q_c^1=\beta$.
-It is represented in \cref{fig:contraction_lips_hull_slope}, from which we derive a geometrical proof:
+It is represented in \Cref{fig:contraction_lips_hull_slope}, from which we derive a geometrical proof:
 \begin{figure}[ht]
@@
 
-For completeness, the edge case in \cref{sec:proof_pi_hull} should be considered as well.
+For completeness, the edge case in \Cref{sec:proof_pi_hull} should be considered as well.
 
@@
 As a conclusion, $\cT$ is a $\gamma(L+1)$-contraction on $\cL_\gamma$.
+\end{proof}
 
@@
 
-\subsection{Proposition \ref{prop:bftq_pi_hull}}
+\subsection{\Cref{prop:bftq_pi_hull}}
 \label{sec:proof_pi_hull}
@@
 \section{Risk-Sensitive Exploration}
-
-We recall the Risk-Sensitive Exploration algorithm in \cref{algo:risk-sensitive-exploration}
+\label{sec:risk-sensitive-supp}
+We recall the Risk-Sensitive Exploration algorithm in \Cref{algo:risk-sensitive-exploration}
 
@@
 
-We recall the scalable version of BFTQ in \cref{algo:bftq_full} and the architecture of the neural network \cref{fig:architecture}.
+We recall the scalable version of BFTQ in \Cref{algo:bftq_full} and the architecture of the neural network \Cref{fig:architecture}.
 
@@
 \label{sec:lagragian}
-As explained on \cref{fig:Lagrangian}, the optimal deterministic policy can be obtained by a line-search on the Lagrange multiplier values $\lambda$.
+As explained on \Cref{fig:Lagrangian}, the optimal deterministic policy can be obtained by a line-search on the Lagrange multiplier values $\lambda$.
 
@@
 \section{Experiments}
+\label{sec:exp-supp}
 
@@
 We compare two approaches for constructing a batch of samples.
-The animations from \url{exploration.html} display the trajectories collected in each intermediate sub-batch.
+The animations from the html page \url{exploration.html} display the trajectories collected in each intermediate sub-batch.
 The first row corresponds to a classical  risk neutral epsilon-greedy exploration policy while the second row showcases a risk-sensitive exploration strategy introduced in the paper.
@@
 
-We display the evolution in the budgeted policy behavior with respect to the budget on different environments. The policies have been learnt with a risk-sensitive exploration.
+We display the evolution in the budgeted policy behaviour with respect to the budget on different environments. The policies have been learnt with a risk-sensitive exploration.
 
@@
 
-In \cref{table:dialogues}, we display two dialogues done with the same BFTQ policy on \texttt{Slot-Filling}.
+\paragraph{Remark on the \texttt{slot-filling} environment} When receiving an utterance, the system can either understand it $(\mu=\mu_u)$ or misunderstand it $(\mu=\mu_m)$ with a fixed probability called the sentence error rate $ser$.
+Then, the speech recognition score is simulated \citep{Khouzaimi2015}: $srs = (1+\exp(-x))^{-1}$ with $x\sim N(\mu, \sigma)$.
+It's the confidence score of the natural language understanding module about the last utterance.
+Note that here are no recognition errors ($ser=0$ and $srs=1$) when the user provides information using the numeric pad.
+
+In \Cref{table:dialogues}, we display two dialogues done with the same BFTQ policy on \texttt{slot-filling}.
 The policy is given two budgets to respect in expectation, $\beta=0$ and $\beta=0.5$.
-For budget 0, one can see that the system never uses the \texttt{ask\_num\_pad} action.
+For $\beta=0$, one can see that the system never uses the \texttt{ask\_num\_pad} action.
 Instead, it uses \texttt{ask\_oral} , an action subject to recognition errors.
 The system keeps asking for the same slot 2, because it has the lowest speech recognition score.
-It eventually summarizes the form to the user, but then reaches the maximum dialogue length and thus faces a dialogue failure.
-For budget 0.5, the system first asks in a safe way, with \texttt{ask\_oral}.
+It eventually summarises the form to the user, but then reaches the maximum dialogue length and thus faces a dialogue failure.
+For $\beta=0.5$, the system first asks in a safe way, with \texttt{ask\_oral}.
 It may want to \texttt{ask\_num\_pad} if one of the speech recognition score is low.
@@
 
-Animations are displayed on the html page \url{corridors.html} for the \texttt{Corridors} environment.
+Animations are displayed on the html page \url{corridors.html} for the \texttt{corridors} environment.
 When the budget is low, the agent takes the safest path on the left.
@@
 \subsection{Reproducibility}
+\label{subsec:reproducibility-supp}
 
-The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \cref{sec:experiements}.
+The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \Cref{sec:experiements}.
 
@@
 
-All environments parameters are displayed in \cref{tab:param-corridors},\cref{tab:param-slot-filling} and \cref{tab:param-highway-env}.
+All environments parameters are displayed in \Cref{tab:param-corridors}, \Cref{tab:param-slot-filling} and \Cref{tab:param-highway-env}.
 
-\paragraph{Remark on the slot-filling environment} When receiving an utterance, the system can either understand it $(\mu=\mu_u)$ or misunderstand it $(\mu=\mu_m)$ with a fixed probability called the sentence error rate $ser$.
-Then, the speech recognition score is simulated \citep{Khouzaimi2015}: $srs = (1+\exp(-x))^{-1}$ with $x\sim N(\mu, \sigma)$.
-It's the confidence score of the natural language understanding module about the last utterance.
-Note that here are no recognition errors ($ser=0$ and $srs=1$) when the user provides information using the numeric pad.
+\paragraph{State-Space}
+
+The states $s$ (from $\os=(s,\beta)$) of the agent are described in the following:
+
+\begin{itemize}
+    \item \texttt{Corridors}: $s = (x,y)$ where $x$ and $y$ are the 2D coordinates of the agent.
+    \item \texttt{Slot-Filling}: $s = (\text{srs},\text{min},a_u,a_s,t)$ where $\text{srs}$ is a vector of the speech recognition score for each slot, $\text{min}$ is a one hot vector describing the minimum of the $\text{srs}$ vector, $a_u$ is a one hot vector of the last user dialogue act and $a_s$ is the one hot vector of the last system dialogue act.
+Finally $t\in[0,1]$ is the fraction of the current turn with the maximum number of turns authorised.
+    \item \texttt{Highway-Env}: the positions $(x, y)$ and velocities $(\dot{x}, \dot{y})$ of every vehicle on the road.
+\end{itemize}
 
@@
 
-All algorithm parameters are displayed in \cref{tab:param-algo-corridors},\cref{tab:param-algo-slot-filling} and \cref{tab:param-algo-highway-env}.
+All algorithm parameters are displayed in \Cref{tab:param-algo-corridors},\Cref{tab:param-algo-slot-filling} and \Cref{tab:param-algo-highway-env}.
+
+\paragraph{A note on the parameters search}
+
+We performed a shallow grid-search for the classic Neural-Network parameters.
+Most of the parameters don't have a strong influence on the results, however in the \texttt{slot-filling} environment, the choice of the regulation weight is decisive.
 
@@
         epoch (FTQ) & 12 & 12\tabularnewline
-        & 0:0.01:1 & -\tabularnewline
-        beta for duplication & 0:0.1:1 & 0:0.1:1\tabularnewline
-        & (1,1) & (1,1)\tabularnewline
-        & 5000 & 5000\tabularnewline
-        & 10 & 10\tabularnewline
-        & 4 & 4\tabularnewline
-        & 1000 & 1000\tabularnewline
+        $\tilde{\cB}$ & 0:0.01:1 & -\tabularnewline
+        $\gamma$ & 1 & 1\tabularnewline
+        $N=|\cD|$ & 5000 & 5000\tabularnewline
+        $N_\text{minibatch}$ & 10 & 10\tabularnewline
+        $N_\text{seeds}$ & 4 & 4\tabularnewline
+        $N_\text{test}$ & 1000 & 1000\tabularnewline
         decay epsilon scheduling & 0.001 & 0.001\tabularnewline
@@
         epoch (FTQ) & 11 & 11\tabularnewline
-        & 0:0.01:1 & -\tabularnewline
-        beta for duplication & - & -\tabularnewline
-        & (1,1) & (1,1)\tabularnewline
-        & 5000 & 5000\tabularnewline
-        & 10 & 10\tabularnewline
-        & 6 & 6\tabularnewline
-        & 1000 & 1000\tabularnewline
+        $\tilde{\cB}$ & 0:0.01:1 & -\tabularnewline
+        $\gamma$ & 1 & 1\tabularnewline
+        $N=|\cD|$ & 5000 & 5000\tabularnewline
+        $N_\text{minibatch}$ & 10 & 10\tabularnewline
+        $N_\text{seeds}$ & 6 & 6\tabularnewline
+        $N_\text{test}$ & 1000 & 1000\tabularnewline
         decay epsilon scheduling & 0.001 & 0.001\tabularnewline
@@
 \end{table}
-%
 \begin{table}[tp]
@@
         epoch (FTQ) & 15 & 15\tabularnewline
-        & 0:0.01:1 & -\tabularnewline
-        beta for duplication & - & -\tabularnewline
-        & (0.9, 0.9) & (0.9, 0.9)\tabularnewline
-        & 10000 & 10000\tabularnewline
-        & 10 & 10\tabularnewline
-        & 10 & 10\tabularnewline
-        & 150 & 150\tabularnewline
+        $\tilde{\cB}$ & 0:0.01:1 & -\tabularnewline
+        $\gamma$ & 0.9 & 0.9\tabularnewline
+        $N=|\cD|$ & 10000 & 10000\tabularnewline
+        $N_\text{minibatch}$ & 10 & 10\tabularnewline
+        $N_\text{seeds}$ & 10 & 10\tabularnewline
+        $N_\text{test}$ & 150 & 150\tabularnewline
         decay epsilon scheduling & 0.0003 & 0.0003\tabularnewline
@@
 \subsubsection{Instructions for reproducibility}
-
-To reproduce the result displayed in \cref{sec:experiements}, first install the following conventional libraries for python3: pycairo, numpy, scipy and pytorch.
-Then execute the commands recall in \cref{fig:instructions} on a Linux Operating System.
-The Graphic Processing Unit used for experiments is an NVDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}.
+\label{subsubsec:instruction-reproducibility}
+To reproduce the result displayed in \Cref{sec:experiements}, first install the following conventional libraries for python3: pycairo, numpy, scipy and pytorch.
+Then, execute the commands in \Cref{fig:instructions} on a Linux Operating System.
+The Graphic Processing Unit used for experiments is an NVIDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}.
 
@@
 \end{minted}
-\caption{Insctructions to reproduce experiments}
+\caption{Instructions to reproduce experiments}
     \label{fig:instructions}
@@
 
-% \clearpage
-% \bibliographystyle{named}
-% \bibliography{budgeted_rl}
-% \end{document}
+\section{The machine learning reproducibility checklist}
+\label{sec:ml-checklist}
+For all models and algorithms presented, indicate if you include:
+
+\begin{itemize}
+    \item  A clear description of the mathematical setting, algorithm, and/or model:
+
+    \begin{itemize}\item \textbf{yes}, see \Cref{sec:intro}, \Cref{sec:bdp}, \Cref{sec:brl},
+    \Cref{sec:risk-sensitive-supp} and \Cref{sec:bftq-full}.
+    \end{itemize}
+    \item An analysis of the complexity (time, space, sample size) of any algorithm:
+    \begin{itemize}\item \textbf{yes}, see \Cref{subsec:parallel-computing}.
+    \end{itemize}
+    \item A link to a downloadable source code, with specification of all dependencies, including external libraries:
+    \begin{itemize}\item \textbf{yes}, see \Cref{subsubsec:instruction-reproducibility} and the folder \texttt{code} in the supplementary material zip file.
+    \end{itemize}
+\end{itemize}
+
+For any theoretical claim, indicate if you include:
+
+\begin{itemize}
+    \item  A statement of the result:
+     \begin{itemize}
+        \item  \textbf{yes}, see \Cref{sec:bdp} and \Cref{sec:brl}.\end{itemize}
+    \item A clear explanation of any assumptions:
+    \begin{itemize}
+        \item  we make one assumption in \Cref{sec:bdp}. We assume the program is feasible for any state. If not, no algorithm would be able to solve it anyway.\end{itemize}
+    \item A complete proof of the claim:
+    \begin{itemize}
+        \item  \textbf{yes}, see \Cref{sec:proofs}.
+We formulate a conjecture in \Cref{rmk:contractivity-smooth} but we provide a sketch of the proof in \Cref{proof:contraction-with-smooth}.\end{itemize}
+\end{itemize}
+
+For all figures and tables that present empirical results, indicate if you include:
+
+\begin{itemize}
+    \item A complete description of the data collection process, including sample size:
+
+        \begin{itemize}
+            \item\textbf{yes}, see \Cref{sec:experiements} and \Cref{sec:algorithms-parameters}.
+        \end{itemize}
+    \item A link to a downloadable version of the dataset or simulation environment:
+
+        \begin{itemize}
+            \item  \textbf{yes}, two environments   are shipped with the supplementary material (in the \texttt{code} folder) and the third one is fetch from a public repository, see \Cref{subsubsec:instruction-reproducibility} for details.
+        \end{itemize}
+    \item An explanation of any data that were excluded, description of any pre-processing step: 
+        \begin{itemize}
+            \item it's \textbf{not applicable} as data comes from simulated environments, so pre-processing steps are not needed.
+        \end{itemize}
+
+    \item An explanation of how samples were allocated for training / validation / testing: 
+        \begin{itemize}
+            \item it's \textbf{not applicable}.
+The complete dataset is used for training.
+There is no need for validation set.
+Testing is performed in the true environment as in classical online learning approaches.
+        \end{itemize}
+
+    \item The range of hyper-parameters considered, method to select the best hyper-parameter configuration, and specification of all hyper-parameters used to generate results:
+
+        \begin{itemize}
+            \item \textbf{yes}, see \Cref{sec:algorithms-parameters}.
+        \end{itemize}
+
+    \item The exact number of evaluation runs:
+
+        \begin{itemize}
+            \item \textbf{yes}, see $N_{seeds}$ in the tables from \Cref{sec:algorithms-parameters}.
+        \end{itemize}
+
+    \item A description of how experiments were run:              \begin{itemize}
+            \item \textbf{yes}, see the two first paragraphs of \Cref{par:ex-explo}.
+        \end{itemize}
+
+    \item A clear definition of the specific measure or statistics used to report results:
+        \begin{itemize}
+            \item  \textbf{yes}, see \Cref{subsec:results}.
+        \end{itemize}
+
+    \item Clearly defined error bars:
+
+        \begin{itemize}
+            \item  \textbf{yes}, we plot 95\% confidence intervals in all figures, see \Cref{subsec:results}.
+        \end{itemize}
+
+    \item A description of results with central tendency (e.g. mean)  variation (e.g. stddev):
+
+        \begin{itemize}
+            \item  \textbf{yes}, we even observe less variability with our novel approach, see \Cref{subsec:results}.
+        \end{itemize}
+    \item A description of the computing infrastructure used:     \begin{itemize}
+            \item The Graphic Processing Unit used for experiments is an NVIDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}.
+        \end{itemize}
+\end{itemize}

### 2019-05-27 main.tex
Updates from Overleaf

@@
 Because we work with  continuous state space $\cS$ and budget space $\cB$, we need to employ function-approximation in order to generalise to nearby states and budgets.
-Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, R, \os')||_2^2$.
+Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, r, \os')||_2^2$.
 Any model can be used, such as linear models, regression trees, or neural networks.

### 2019-06-11 main.tex
Updates from Overleaf

@@
 % [preprint] option:
-%     \usepackage[preprint]{neurips_2019}
+\usepackage[preprint,nonatbib]{neurips_2019}
 
 % to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage{neurips_2019}
+%\usepackage{neurips_2019}
 
@@
 % Paper specific includes and commands
+\usepackage[round]{natbib}
+\usepackage{xcolor}
 \usepackage{times}
@@
 \usepackage{hyperref}
+
+\definecolor{mylinkcolor}{HTML}{e83f6f}
+\definecolor{mycitecolor}{HTML}{2980B9}
+\definecolor{myurlcolor}{HTML}{d63230}
+% \definecolor{myalgocolor}{HTML}{32936f}
+\definecolor{myalgocolor}{HTML}{000000}
+
+\hypersetup{
+  linkcolor  = mylinkcolor,
+  citecolor  = mycitecolor,
+  urlcolor   = myurlcolor,
+  colorlinks = true,
+}
+
 \usepackage{cleveref}
@@
 \usepackage{makecell}
-\usepackage[lofdepth,lotdepth]{subfig}
+%\usepackage[lofdepth,lotdepth]{subfig}
 \usepackage{wrapfig}
 \usepackage{enumerate}
-\usepackage{minted}
+%\usepackage{minted}
+\usepackage{listings}
+\lstset{basicstyle=\ttfamily,
+  showstringspaces=false,
+  commentstyle=\color{gray},
+  keywordstyle=\color{blue}
+}
 
@@
 \author{
-Nicolas Carrara$^{1,}$\thanks{equal contribution}\and
-Edouard Leurent$^{1,3,}$\footnotemark[1]\and
-Romain Laroche$^{4}$\and
-Tanguy Urvoy$^{2}$\and
-Odalric Maillard$^{1}$\and
-Olivier Pietquin$^{1,2,}$\thanks{now with Google Research - Brain Team, Paris, France.}\\
-\affiliations $^1$SequeL team, INRIA Lille -- Nord Europe, France\\
-$^2$Orange Labs, Lannion, France\\
-$^3$Renault Group, France\\
-$^4$Microsoft Research , Montreal, Canada
-\emails
-\{nicolas.carrara, edouard.leurent, odalric.maillard, olivier.pietquin\}@inria.fr,
-romain.laroche@microsoft.com, tanguy.urvoy@orange.com
+Nicolas Carrara\thanks{Both authors contributed equally.}\\
+SequeL team, INRIA Lille -- Nord Europe\thanks{Univ. Lille, CNRS, Centrale Lille, INRIA UMR 9189 - CRIStAL, Lille, France}\\
+\texttt{nicolas.carrara@inria.fr}
+\And
+Edouard Leurent\footnotemark[1]\\
+SequeL team, INRIA Lille -- Nord Europe\footnotemark[2]\\
+Renault Group, France\\
+\texttt{edouard.leurent@inria.fr}
+\And
+Romain Laroche\\
+Microsoft Research, Montreal, Canada\\
+\texttt{romain.laroche@microsoft.com}
+\And
+Tanguy Urvoy\\
+Orange Labs, Lannion, France\\
+\texttt{tanguy.urvoy@orange.com}
+\And
+Odalric-Ambrym Maillard\\
+SequeL team, INRIA Lille -- Nord Europe\\
+\texttt{odalric.maillard@inria.fr}
+\And
+Olivier Pietquin\\
+Google Research - Brain Team\\
+SequeL team, INRIA Lille -- Nord Europe\footnotemark[2]\\
+\texttt{pietquin@google.com}
 }
@@
 It relies on a notion of risk implemented in the shape of a cost signal constrained  to lie below an -- adjustable -- threshold.
-    %A key difference between Constrained Markov Decision Processes (CMDP) and
-    %BMDP lies in the ability to change this threshold in real-time.
-    %TU: on ne comprend pas cette subtilité à ce stade
-    % il vaut mieux développer plus loin.
-    % "that -- importantly -- can be modified in real-time."
     So far, BMDPs could only be solved in the case of finite state spaces with known dynamics.
@@
 It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R_r, \gamma)$ with measurable state space $\cS$, discrete actions $\cA$, unknown rewards $R_r\in\Real^{\cS \times \cA}$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ , where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$.
-Formally, we seek a policy $\pi\in\cM(A)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$.
+Formally, we seek a policy $\pi\in\cM(\cA)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$.
 
@@
 Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.
-The proofs of our main results are provided \Cref{sec:proofs}.
+The proofs of our main results are provided in \Cref{sec:proofs}.
 
@@
 
-\begin{theorem}[Contractivity of $\cT$]
-\label{thm:contraction}
-If $|\cA| \geq 2$, then $\cT$ is not a contraction.
-\end{theorem}
-
 \paragraph{Budgeted Value Iteration}
@@
 We introduce \Cref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.
+If it were to converge to a unique fixed point, this algorithm would provide a way to compute $Q^*$ and recover the associated optimal budgeted policy $\pi_\text{greedy}(\cdot~; Q^*)$.
 
-Unfortunately, as $\cT$ is not a contraction, we can guarantee neither the convergence of this procedure nor the unicity of its fixed points.
+\begin{theorem}[Non-contractivity of $\cT$]
+\label{thm:contraction}
+For any BMDP ($\cS,\cA,P,R_r,R_c,\gamma$) with $|\cA| \geq 2$, $\cT$ is not a contraction.
+Precisely: $\forall\epsilon>0, \exists Q^1,Q^2\in(\Real^2)^{\ocS\ocA}:\|\cT Q^1-\cT Q^2\|_\infty \geq \frac{1}{\epsilon}\|Q^1-Q^2\|_\infty$.
+\end{theorem}
+
+Unfortunately, as $\cT$ is not a contraction, we can guarantee neither the convergence of \Cref{algo:bvi} nor the unicity of its fixed points.
 Despite those theoretical limitations, we empirically observed the convergence to a fixed point in our experiments (\Cref{sec:experiements}).
@@
 \end{equation*}
-We introduce in \Cref{algo:bftq} the \emph{Budgeted-Fitted-Q} (BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs.
+We introduce in \Cref{algo:bftq} the \emph{Budgeted-Fitted-Q} (\BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs.
 Because we work with  continuous state space $\cS$ and budget space $\cB$, we need to employ function-approximation in order to generalise to nearby states and budgets.
@@
 \label{sec:scalable-bftq}
-In this section, we introduce an implementation of the BFTQ algorithm designed to operate efficiently and handle large batches of experiences $\mathcal{D}$.
+In this section, we introduce an implementation of the \BFTQ algorithm designed to operate efficiently and handle large batches of experiences $\mathcal{D}$.
 
@@
 }}
-\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(\os,\oa))$\tcp*[f]{Budget $\beta$ always respected}}
+\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(\os,\oa))$\tcp*[f]{budget $\beta$ always respected}}
 \caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
@@
 %\vspace{0pt}
-\begin{wrapfigure}{r}{0.5\textwidth}
+%\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{figure}[ht]
     \centering
-    \includegraphics[width=\linewidth]{source/img/pi.pdf}
+    \includegraphics[width=0.7\linewidth]{source/img/pi.pdf}
     \caption{Representation of $\pi_\text{hull}$.
@@
     \label{fig:hull}
-\end{wrapfigure}
+\end{figure}
+%\end{wrapfigure}
 %\end{minipage}
@@
 The resulting total time-complexity is $\cO(\frac{|\mathcal{D}||\cA||\tilde{\cB}|}{1-\gamma} \log |\cA||\tilde{\cB}|)$.
-This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA,\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass.
+This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass.
 By using multiprocessing in the computations of $\pi_\text{hull}$, we enjoy a linear speedup.
@@
 The latter is chosen to be a $\epsilon$-greedy policy slowly transitioning from a random to a greedy policy\footnote{We train this greedy policy using \FTQ.} that aims to maximise $\expectedvalue_{\pi} G_r^\pi$ regardless of $\expectedvalue_{\pi} G_c^\pi$.
-The quality of the resulting batches $\cD$ is assessed by training a BFTQ policy and comparing the resulting performance.
+The quality of the resulting batches $\cD$ is assessed by training a \BFTQ policy and comparing the resulting performance.
 
@@
 We solve each of these CMDP using the standard technique of Lagrangian Relaxation: the cost constraint is converted to a soft penalty weighted by a Lagrangian multiplier $\lambda$ in a surrogate reward function: $\max_{\pi} \expectedvalue_{\pi}[G_r^\pi - \lambda G_c^\pi]$.
-The resulting MDP can be solved by any RL algorithm, and we chose $\FTQ$ for being closest to $\BFTQ$.
-In our experiments, a single training of $\BFTQ$ corresponds to 10 trainings of $\FTQ_l$ policies.
+The resulting MDP can be solved by any RL algorithm, and we chose \FTQ for being closest to \BFTQ.
+In our experiments, a single training of \BFTQ corresponds to 10 trainings of \FTQl policies.
 Each run was repeated $N_{\text{seeds}}$ times.
@@
 
-We first bring to light the role of risk-sensitive exploration in the \texttt{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in \Cref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch.
+We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in \Cref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch.
 The trajectories (orange) in the risk-neutral batch are concentrated along the risky corridor (red) and ignore the safe corridor (green), which results in bad performances in the low-risk regime.
@@
     \includegraphics[width=0.49\linewidth]{source/img/highway}
-    \caption{Performance comparison of \FTQl and BFTQ on \texttt{slot-filling} (left) and \texttt{highway-env}(right) }
+    \caption{Performance comparison of \FTQl and \BFTQ on \text{slot-filling} (left) and \text{highway-env}(right) }
     \label{fig:results}
@@
 For each algorithm, we plot the reward-cost trade-off curve.
-In both cases, BFTQ performs almost as well as \FTQl despite only requiring a single model.
-All budgets are well-respected on \texttt{slot-filling}, but on \texttt{highway-env} we can observe an underestimation of $Q_c$, since e.g.
+In both cases, \BFTQ performs almost as well as \FTQl despite only requiring a single model.
+All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can observe an underestimation of $Q_c$, since e.g.
 $\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
@@
 
-%    \section*{Acknowledgments}
-%    %
-%    This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
+    \section*{Acknowledgments}
     %
-%    Thanks to Guillaume Gautier, Fabrice Clerot, Xuedong Shang
-\clearpage
+    This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
+We thank Guillaume Gautier, Fabrice Clerot, Xuedong Shang for the helpful discussions and valuable insights.
+%\clearpage
 \bibliographystyle{named}

### 2019-06-11 supplementary.tex
Updates from Overleaf

@@
 \begin{center}
-\LARGE Appendix
+\LARGE Appendices
 \end{center}
@@
 Let $(\cS, \cA, P, R_r, R_c)$ be a BMDP.
-For any $0 < \epsilon < 1$, we define $Q_\epsilon^1$ and $Q_\epsilon^2$ as:
+For any $\epsilon > 0$, we define $Q_\epsilon^1$ and $Q_\epsilon^2$ as:
 
@@
       \begin{cases}
-    (0, 1), & \text{if } a = a_0 \\
-    (\frac{1}{\epsilon}, 1+\epsilon), & \text{if } a \neq a_0
+    (0, 0), & \text{if } a = a_0 \\
+    \left(\frac{1}{\gamma}, \epsilon\right), & \text{if } a \neq a_0
   \end{cases}\\
@@
       \begin{cases}
-    (1, 0), & \text{if } a = a_0 \\
-    (1+\frac{1}{\epsilon}, \epsilon), & \text{if } a \neq a_0
+    (0, \epsilon), & \text{if } a = a_0 \\
+    \left(\frac{1}{\gamma}, 2\epsilon\right), & \text{if } a \neq a_0
   \end{cases}
 \end{align*}
-Then, $\|Q_1-Q_2\|_\infty = 1$.
+Then, $\|Q_1-Q_2\|_\infty = \epsilon$.
  $Q_\epsilon^1$ and $Q_\epsilon^2$ are represented in \Cref{fig:concavity_example}.
@@
 
-But for $\oa=(a,\beta_a)$ with $\beta_a = 1+\epsilon/2$, we have:
+But for $\oa=(a,\beta_a)$ with $\beta_a = \epsilon$, we have:
 \begin{align*}
     \|\cT Q_\epsilon^1(\os, \oa) - \cT Q_\epsilon^2(\os, \oa)\|_\infty &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)} \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^1_\epsilon)}Q^1_\epsilon(\os',\oa') - \expectedvalueover{\oa'\sim\pi_\text{greedy}(Q^2_\epsilon)}Q^2_\epsilon(\os',\oa')\right\|_\infty \\
-    &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)}(\frac{1}{2\epsilon}, 1+\frac{\epsilon}{2}) - (1+\frac{1}{\epsilon},\epsilon )\right\|_\infty \\
-    &= \gamma(1+\frac{1}{2\epsilon})
+    &= \gamma\left\|\expectedvalueover{\os'\sim\ov{P}(\os'|\os,\oa)}\left(\frac{1}{\gamma}, \epsilon\right) - (0, \epsilon)\right\|_\infty \\
+    &= \gamma\frac{1}{\gamma} = 1
 \end{align*}
-Hence, $\|\cT Q_\epsilon^1 - \cT Q_\epsilon^2\|_\infty \geq \gamma(1+\frac{1}{2\epsilon})$
+Hence, 
+\begin{align*}
+    \|\cT Q_\epsilon^1 - \cT Q_\epsilon^2\|_\infty &\geq 1 = \frac{1}{\epsilon} \|Q_1-Q_2\|_\infty
+\end{align*}
 
-As a conclusion, there does not exist $L>0$ such that:
+In particular, there does not exist $L>0$ such that:
 $$\forall Q_1,Q_2\in(\Real^2)^{\ocS\ocA}, \|\cT Q^1 - \cT Q^2\|_\infty \leq L \|Q^1 - Q^2\|_\infty$$
@@
 
-\begin{minted}{sh}
+%\begin{minted}{sh}
+\begin{lstlisting}[language=bash,caption={bash version}]
 # Install highway-env
@@
 python3 main/egreedy/main-egreedy.py config/highway-easy.json 0 10
-\end{minted}
+\end{lstlisting}
+%\end{minted}
+
 \caption{Instructions to reproduce experiments}

### 2019-06-11 source/counter_example.py

@@ -6,42 +6,32 @@ import seaborn as sns
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif', size=15)
 
-eps = 0.7
-beta = 1+eps/2
+eps = 0.5
+beta = eps
 
-Q1 = np.array([[1, 1+eps], [0, 1/eps]])
-Q2 = np.array([[0, eps], [1, 1+1/eps]])
-sq1 = np.array([[0, 0], [1, 0], [1, 1], [0, 1], [0, 0]], dtype=np.float32)
-sq2 = sq1.copy()
-sq1[:, 0] += eps
-sq1[:, 1] += 1/eps
-beta_v = np.array([[beta, 0], [beta, 1+1/eps]])
-beta_p = np.array([[beta, 1/(2*eps)]])
+Q1 = np.array([[0, eps], [0, 1]])
+Q2 = np.array([[eps, 2*eps], [0, 1]])
 
-fig = plt.figure(figsize=(3, 3.5))
+fig = plt.figure(figsize=(3, 3))
 ax = fig.add_subplot(1, 1, 1)
-
 plt.plot(Q1[0], Q1[1], marker='.', markersize=12)
 plt.plot(Q2[0], Q2[1], marker='.', markersize=12)
-plt.plot(sq1[:, 0], sq1[:, 1], color="grey", linestyle="dashed")
-plt.plot(sq2[:, 0], sq2[:, 1], color="grey", linestyle="dashed")
-# plt.plot(beta_v[:, 0], beta_v[:, 1], color="grey", linestyle="dashed")
-plt.plot(beta_p[:, 0], beta_p[:, 1], color="grey", marker='.', markersize=12)
 # plt.legend(["Q1", "Q2"])
+plt.text(eps+0.01, 1/2, r'$\beta$', rotation=0, horizontalalignment="left", verticalalignment="center")
 
 plt.axis('equal')
 
-xticks = [0, 1, beta, 1+eps]
-plt.xticks(xticks, ('0', '1', r'$\beta$', r'$1+\epsilon$'))
-yticks = [1, 1/eps, 1+1/eps]
-plt.yticks(yticks, ('1', r'$\frac{1}{\epsilon}$', r'$1+\frac{1}{\epsilon}$'))
+xticks = [0, eps, 2*eps]
+plt.xticks(xticks, ('0', r'$\epsilon$', r'$2\epsilon$'))
[…]

### 2019-06-11 poster/poster.tex
Add structure for poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (15/5/13)
%
% Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paperwidth=36in, paperheight=48in,portrait,fontscale=0.35, margin=2cm]{baposter}

\usepackage[utf8]{inputenc}
\usepackage[font=normalsize,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\captionsetup{justification=centering}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lmodern}
\usepackage{relsize} % Used for making text smaller in some places

% Other packages
\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}}
\def\imagetop#1{\vtop{\null\hbox{#1}}}
\usepackage{adjustbox}
\usepackage{subfig}
\usepackage{floatrow} % sidecapfloat
\usepackage{setspace}
\usepackage{xspace}
\usepackage{noReferences}
\usepackage{caption}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
%\usepackage{subcaption}
\usepackage{epsf}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{float}
\usepackage{cases}
\usepackage{verbatim}
\usepackage{array}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{listings}
%\usepackage[authoryear]{natbib}
\usepackage{authblk}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{eqparbox}
\usepackage{xcolor}
\usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}
\usepackage{lipsum}
\usepackage{svg}

\bibliography{../references.bib}

\input{style}
\input{../mathdef}

\begin{document}

%\background{ % Set the background to an image (background.pdf)
%\begin{tikzpicture}[remember picture,overlay]
%\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%{\includegraphics[height=1.1\textheight]{background.png}};
%\end{tikzpicture}
%}
\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=boxcolor, % Background color for the content in the content boxes
headershape=roundedright, % Specify the rounded corner in the content box headers
%headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
headerfont=\Large\bf\textsc, %Sans Serif
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain,
textfont={\setlength{\parindent}{0.0em}\sffamily},
headerheight={0.05\textheight},
eyecatcher=true
%columns=5
}
%
%----------------------------------------------------------------------------------------
%	Title and authors
%----------------------------------------------------------------------------------------
%
{
\includegraphics[width=14em]{./img/renault_group}
}
{
Practical Open-Loop Optimistic Planning
}
{
Edouard Leurent, Odalric-Ambrym Maillard
\vspace{-3\baselineskip}
}
{
\includegraphics[width=14em]{./img/inria_sc}
}

\setlength{\colheight}{0.92\textheight}

%----------------------------------------------------------------------------------------
%	Motivation
%----------------------------------------------------------------------------------------

\headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{

}

%----------------------------------------------------------------------------------------
%	Setting
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Open-Loop Optimistic Planning}}{name=setting,span=1,column=0,row=0,below=motivation}{

}

%----------------------------------------------------------------------------------------
% Tools
%----------------------------------------------------------------------------------------
\headerbox{\textsc{What's wrong with OLOP?}}
{name=tools,column=0,below=setting,span =1}
{

}

%----------------------------------------------------------------------------------------
%	KL-OLOP
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Kullback-Leibler OLOP}}{name=klolop,column=1,span=2}{

}

%----------------------------------------------------------------------------------------
%	Sample Complexity
%----------------------------------------------------------------------------------------
\headerbox{Sample complexity}
{name=sample,span=2,column=1,below=klolop}{ 
}

%----------------------------------------------------------------------------------------
%	Time and Memory Complexity
%----------------------------------------------------------------------------------------
\headerbox{Time and memory complexity}{name=time,span=2,column=1,below=sample}{

}

%
%%----------------------------------------------------------------------------------------
%% Experiments
%%----------------------------------------------------------------------------------------
%
\headerbox{Experiments}{name=experiments,span=2,column=1,below=time, above=bottom}{

}

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
% \headerbox{Acknowledgements}{name=ack,column=0,span=1,below=tools}{
% This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
% }

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------
\headerbox{References}{name=refs,column=0,span=1,below=tools, above=bottom}{
    {
    \AtNextBibliography{\footnotesize}
    \setlength{\bibitemsep}{3pt}
    \printbibliography[heading=none]
    }
}

\end{poster}

\end{document}

### 2019-08-19 author_response.tex
Updates from Overleaf

\documentclass{article}

\usepackage{neurips_2019_author_response}
%\input{mathdef.tex}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{todonotes}
\long\def\td#1{\todo[inline]{#1}}

\usepackage{lipsum}

\begin{document}

%\section{Common}

First of all, we would like to thank the reviewers for their valuable comments, which will help us improve the revised manuscript. We endeavour to address all the major remarks below.

Second, we would like to notify all the reviewers about a major theoretical improvement allowing us to transform the conjecture at line 107 into a theorem. At submission time, the proof was missing two steps (lines 443--444 and 465), and we decided to present it as a conjecture for the sake of caution. In the meantime, we have proved them and we propose to make it a theorem.
%The \textbf{main concern} expressed was perhaps that of Reviewer \#3 regarding the presentation of the result on the optimality operator. The reason for this is that the proof was still incomplete at the time of the submission deadline. Two small parts were missing: l443-444 and l465. Hence, by way of precaution, we preferred to present this property as a conjecture with a sketch of proof rather than a result. Our intention was to convey to the reader an intuition of why we do observe empirical convergence despite our negative result regarding contraction in the general case, though details were missing. The proof has been completed since, and the contractivity result on smooth functions is now presented as a theorem.

Third, a remark common to all reviewers is that Algorithm 3 (convex hull policy) lacks a detailed explanation, which we acknowledge: it was a questionable decision due to space constraints. We will include a descriptive paragraph.

\paragraph{\large Reviewer \#1}

%We will add the missing appendices.

%The framework searches for a minimum cost policy given a maximum reward policy satisfying budget restrictions. In principle it could be framed the other way around searching for a maximum reward policy given minimum cost. Add a note of why this is not a good idea.

\textbf{In principle it could be framed [...] why this is not a good idea.}
Swapping the min and max would lead to a different set of policies. Indeed,  the BMDP solution is generally not a saddle point: max-min $<$ min-max. % and make the constraint pointless (a minimum-cost policy satisfies any feasible budget). The true objective in a BMDP is really to maximise the rewards under cost constraint, and the cost minimisation is rather an artefact only introduced to break the ties in the definition of $V_c^*$.

\textbf{When creating batch samples what happens when the budget is zero? The episodes ends with a negative reward?} If the agent explores/exploits with zero budget, it will always select the estimated safest action, with minimal estimated expected cost $Q_c$. If costs are still incurred, the model $Q_\theta$ will be updated accordingly.

\textbf{As it is not a contraction in general, in which cases
should we not use the proposed approach?} The Remark 1 and the counter-example in Theorem 2 both indicate that the proposed approach is more likely to diverge when $Q^*$ is steep, i.e. in problems where a small increase in budget $\beta$ leads to substantial gains in rewards. As in most theoretical works, the required assumptions may be transgressed in real world. But, it does not necessarily imply that the algorithm would not work in practice. For a sensitive real world application, we would not refrain from using the algorithm, but would strongly advise to supervise it.% For a sensitive real world application, we would not refrain from using the algorithm, but recommend setting strong safety margins on the budget $\beta$.

\textbf{How can the budget lie between two $Q_c$, is it not the case that all of
them should respect the budget constraint?}
$Q^*_c(\overline{s},\overline{a})$ is the cost induced by \textit{first} executing action $\overline{a}$ and only \textit{then} following the optimal budgeted policy $\pi^*$. Among all these actions, some of them may not be feasible and exceed the budget $\beta$, which is why an additional optimisation step is required in the definition of the greedy policy.
%TU: To dodge the "expectation weakness"
%can we tune the mixture to obtain safer policies (how "safer" being left as future work) ?

\paragraph{\large Reviewer \#2}

%\td{I am confused as to why/how the $\beta$ evolves during the learning process and how this effects the goals one has when using a BMDP. Do we not care about how much risk is induced (or negative effects experienced) during training? Moreover, if the policy selects a new $\beta$ at each new time step, how do we enforce a *specific* $\beta$ when we wish to *use* the policy?}
\textbf{(A.1)} 
The cost constraints only apply to the final trained budgeted policy, we indeed do not care about the costs incurred during training. To enforce a specific $\beta$ at test time, set the initial augmented state to $\overline{s}_0 =(s_0,\beta)$.

%\td{The convex hull procedure seems critical to being able to actually use the algorithm, but the explanation is lacking any intuitive interpretation. For example, it's not clear to me from 4.1 and Algorithm 3 exactly how "Budget $\beta$ [is] always respected" (comment on Algorithm 3, line 9). Can the authors provide more explanation/intuition about what is going on here?}

\textbf{(A.2)} If we reach Line 9 in Algorithm 3, it must be that the condition $\beta < q_c^2$ in Line 5 never holds, which means that every action verifies $q_c \leq \beta$: it is a case where the budget is always respected.

% (3) The authors state that "The pseudo-code of our exploration procedure is shown in Algorithm 4 in Appendix B." Since this component of your algorithm is one of the main hypotheses that is being validated, it should appear in the main paper. Suggest swapping for Algorithms 1/2 since these are basic extensions of existing techniques.

\textbf{(A.3)} We will follow your suggestion: indeed Algorithm 4 is more specific to our work than Algorithms 1/2.

%\td{ There is a mismatch between the introduction of the paper, which speaks generally to using BMDPs, and the actual experiments run, which show the benefit of using BMDPs *compared to* computing a set of solutions using existing techniques like FTQ($\lambda$). The introduction needs to mention that approaches like the latter *are* available solutions and frame the contribution of the paper rather as one of providing a "better" solution in whichever way the authors feel this is best described (more-efficient, etc.).}
\textbf{(B)} We will make the introduction clearer and properly introduce the "set of solutions" techniques like FTQ($\lambda$) along with a proper comparison to the BMDP approach. Namely, they do not recover optimal budgeted policies, in addition to being inefficient.   
%\textbf{(B)} FTQ($\lambda$) is a good heuristic but is not optimal for BMDPs (in addition to being inefficient). To be convinced of this: an MDP (obtained after relaxation) always admits a deterministic solution, whereas it is not necessarily the case for a CMDP and hence a BMDP.

%\td{Several times in the paper, it is mentioned that experiments are done in "two environments," but aren't there three?}

\textbf{Minor comment 2} Indeed, but one of them is just a toy example to illustrate the risk-sensitive exploration only.

%\td{The definition in (2) is odd given that you say the "budget evolves as part of the dynamics." That is, if $\beta'=\beta_a$ (as suggested by the Dirac function, then it is merely whatever the action says it was, correct? Why is that part of "the dynamics?"}

\textbf{Minor comment 3} This is absolutely correct. The problem is that since we framed the problem in an augmented space where the budget $\beta$ is part of the state $\overline{s}$, its evolution can only be described within the dynamics of $\overline{s}$ (which we specify). We agree that this makes the notations quite obscure for a very simple idea, but on the other hand casting the problem in such a way drastically simplifies the next definitions and proofs, e.g. that of Proposition 1.

\paragraph{\large Reviewer \#3}

\textbf{Presentation of the result on the optimality operator.} The main criticism concerned the presentation of the contractivity of $\mathcal{T}$ over smooth Q-functions as a conjecture rather than a proven result. We hope that these reservations have been lifted by the upgrade of the conjecture to a theorem.

\textbf{The exploration strategy from Section 3.2 is merely described}: This strategy is motivated by the observation that a wide variety of risk levels needs to be experienced during training, which can be achieved by enforcing the risk constraints during data collection. This intuition was meant to be conveyed by the corridor example where a conventional greedy exploration procedure fails to visit the safe region. We will add an additional discussion to clarify this point.

\textbf{The implementation proposed in Section 4 seems to rely heavily on results from Boutilier and Lu, 2016, but a lot of the necessary context is missing}. Several differences exist between the approach from Boutilier and Lu (2016)  - the greedy budget allocation (GBA) algorithm - and ours: they rely on a finite set of non-dominated budget points $B$ which grows exponentially with the horizon and becomes uncountably infinite in continuous state spaces. They also compute and sort a matrix of bang-per-buck ratios of size $|S|\times B$, which is again infeasible when $S$ is continuous. In contrast, we instead rely on estimating the optimal cost-to-go $Q_c^*$, which requires an additional min constraint in its definition that does not appear in Boutilier and Lu (2016) (they only estimate $V_r^*$).

\end{document}

### 2019-08-19 poster/poster.tex
Updates from Overleaf

@@
 
-\documentclass[paperwidth=36in, paperheight=48in,portrait,fontscale=0.35, margin=2cm]{baposter}
+\documentclass[a0paper,portrait,fontscale=0.355, margin=2cm]{baposter}
 
@@
 \usepackage{svg}
+\usepackage{cleveref}
 
-\bibliography{../references.bib}
+\bibliography{../budgeted_rl.bib}
 
@@
 {
-\includegraphics[width=14em]{./img/renault_group}
+\includegraphics[width=18em]{./img/companies}
 }
 {
-Practical Open-Loop Optimistic Planning
+Budgeted Reinforcement Learning in Continuous State Space
 }
 {
-Edouard Leurent, Odalric-Ambrym Maillard
+Nicolas Carrara, Edouard Leurent, Romain Laroche, \\Tanguy Urvoy, Odalric-Ambrym Maillard, Olivier Pietquin
 \vspace{-3\baselineskip}
@@
 
+\textbf{Markov Decision Process} $(\cS, \cA, P, R_r, \gamma)$:
+\begin{equation*}
+    \max_\pi \expectedvalue_\pi \underbrace{\sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)}_{G_r^\pi}
+\end{equation*}
+
+\begin{itemize}[nolistsep]
+    \item[\incarrow] \hlr{Single scalar reward} to represent \hlr{multiple} \hlr{contradictory} aspects (e.g. complete task and avoid collisions)
+    \item[\incarrow] \hlr{No control over the spread} of the performance distribution\\
+
+\end{itemize}
+
+\textbf{Constrained MDP} $(\cS, \cA, P, R_r, R_c, \gamma)$
+\begin{itemize}[nolistsep]
+    \item $[$\cite{BEUTLER1985236,Altman95constrainedmarkov}$]$
+    \item Introduce a \hlb{cost signal} $R_c$
+    \item Constrained objective
+\begin{equation*}
+\begin{array}{lcr}
+ \displaystyle \max_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s] & \text{\hlb{ s.t. }} & \expectedvalue[G_c^\pi | s_0=s] \leq \beta
+\end{array}
+\end{equation*}
+\item[\incarrow] The cost budget $\beta$ \hlr{cannot be changed} after training\\
+\end{itemize}
+
+\textbf{Budgeted MDP}
+
+\begin{itemize}[nolistsep]
+    \item $[$\cite{Boutilier_Lu:uai16}$]$
+    \item We seek \hlg{one general policy} $\pi(s,\beta)$ that solves every CMDP for any $\beta$
+    \item[\incarrow] Can only be solved for \hlr{finite} $\cS$ and \hlr{known} $P, R_r, R_c$.
+\end{itemize}
 }
@@
 %----------------------------------------------------------------------------------------
-\headerbox{\textsc{Open-Loop Optimistic Planning}}{name=setting,span=1,column=0,row=0,below=motivation}{
+\headerbox{\textsc{Setting}}{name=setting,span=1,column=0,row=0,below=motivation}{
+
+\textbf{Budgeted policies} $\pi$
+\begin{itemize}
+    \item Take a budget $\beta$ as an additional input
+    \item Output a next budget $\beta'$ 
+\end{itemize}
+\begin{equation*}
+    \pi:\underbrace{(s,\beta)}_{\os} \rightarrow \underbrace{(a,\beta')}_{\oa}
+\end{equation*}
+
+\textbf{2D  signals}
+\begin{enumerate}
+    \item Rewards $R = (R_r, R_c)$
+    \item Returns $G^\pi = (G_r^\pi, G_c^\pi)$
+    \item Values $V^\pi = (V_r^\pi, V_c^\pi)$ and $Q^\pi = (Q_r^\pi, Q_c^\pi)$\\
+\end{enumerate}
+
+\textbf{Policy Evaluation}\\
+
+The Bellman Expectation equations% [\cite{Bellman}]
+are preserved, and the Bellman Expectation Operator $\cT^\pi$ is a \hlg{$\gamma$-contraction}.
+}
+
+%----------------------------------------------------------------------------------------
+%	Budgeted Optimality
+%----------------------------------------------------------------------------------------
+\headerbox{Budgeted Optimality}{name=bopt,column=0,span=1,below=setting}{
+\begin{definition*}In that order, we want to:
+\begin{enumerate}
+    \item[(i)] \hlb{Respect the budget $\beta$}: 
+    \begin{equation*}
+    \Pi_a(\os) \eqdef \{\pi\in\Pi: V_c^\pi(s, \beta) \leq \beta\}
+    \end{equation*}
+    \item[(ii)] \hlg{Maximise the rewards}:
+    \begin{equation*}
+        V_r^*(\os) \eqdef {\max}_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\quad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
+    \end{equation*}
+    \item[(iii)] \hlr{Minimise the costs}: 
+    \begin{equation*}
+        V_c^*(\os) \eqdef {\min}_{\pi\in\Pi_r(\os)}  V_c^\pi(\os), \qquad\quad \Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
+    \end{equation*}
+\end{enumerate}
+
+We define the budgeted action-value function $Q^*$ similarly
+\end{definition*}
 
@@
 %----------------------------------------------------------------------------------------
-% Tools
+%	Budgeted Dynamic Programming
 %----------------------------------------------------------------------------------------
-\headerbox{\textsc{What's wrong with OLOP?}}
-{name=tools,column=0,below=setting,span =1}
-{
+\headerbox{Budgeted Dynamic Programming}{name=bdp,column=1,span=2}{
 
+\begin{minipage}{0.7\textwidth}
+\bookboxx{
+\begin{theorem*}[Budgeted Bellman Optimality]
+\label{thm:bellman-optimality}
+$Q^*$ verifies:
+\begin{equation}
+\label{eq:bellman-optimality}
+    \hleb{\vphantom{\frac{A}{A}} Q^{*}(\os, \oa) = \cT Q^{*}(\os, \oa)}  \eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in \ocA} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'}),
+\end{equation}
+where the greedy policy $\pi_\text{greedy}$ is defined by:
+\begin{subequations}
+\label{eq:pi_greedy}
+\begin{align}
+    \pi_\text{greedy}(\oa|\os; Q) \in &\hler{\argmin}_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa), \label{eq:pi_greedy_cost}\\
+    \text{where }\quad\Pi_r^Q \eqdef &\hleg{\argmax}_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \label{eq:pi_greedy_reward}\\
+    & \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \hleb{\leq \beta} \label{eq:pi_greedy_constraint}
+\end{align}
+\end{subequations}
+\end{theorem*}
+}
+\bookboxx{
+\begin{proposition*}
+\label{prop:greedy_optimal}
+$\pi_\text{greedy}(\cdot~; Q^*)$ is \hlg{simultaneously optimal} in all states $\os\in\ocS$: $$\pi_\text{greedy}(\cdot~; Q^*)\in\Pi^*(\os)$$
+
+In particular, $V^{\pi_\text{greedy}(\cdot; Q^*)} = V^*$ and $Q^{\pi_\text{greedy}(\cdot; Q^*)}= Q^*$.
+\end{proposition*}
+}
+
+\end{minipage}
+\hfill%
+\begin{minipage}{0.25\textwidth}
+\begin{algorithm}[H]
+\DontPrintSemicolon
+\KwData{$P, R_r, R_c$}
+\KwResult{$Q^*$}
+$Q_{0} \leftarrow 0$\;
+\Repeat{convergence}{
+    $Q_{k+1} \leftarrow \cT Q_k$\;
+}
+\caption{Budgeted Value Iteration}
+\label{algo:bvi}
+\end{algorithm}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{source/img/concavity_example.pdf}    
+\end{center}
+\end{minipage}
+
+\bookboxx{
+\begin{theorem*}[Contractivity]
+\label{thm:contraction}
+For any BMDP ($\cS,\cA,P,R_r,R_c,\gamma$) with $|\cA| \geq 2$, \hlr{$\cT$ is not a contraction}. 
+$$\forall\epsilon>0, \exists Q^1,Q^2\in(\Real^2)^{\ocS\ocA}:\|\cT Q^1-\cT Q^2\|_\infty \geq \frac{1}{\epsilon}\|Q^1-Q^2\|_\infty$$
+\end{theorem*}
+}
+
+Despite those theoretical limitations, we observed \hlg{empirical convergence} in our experiments
+
+\bookboxx{
+\begin{remark*}[Contractivity on smooth $Q$-functions]
+\label{rmk:contractivity-smooth}
+We conjecture that \hlg{$\cT$ is a contraction when restricted} to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is \hlg{$L$-Lipschitz} with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
+\end{remark*}
+}
 }
@@
 %----------------------------------------------------------------------------------------
-%	KL-OLOP
+%	Budgeted Reinforcement Learning
 %----------------------------------------------------------------------------------------
-\headerbox{\textsc{Kullback-Leibler OLOP}}{name=klolop,column=1,span=2}{
+\headerbox{Budgeted Reinforcement Learning}
+{name=brl,span=2,column=1,below=bdp}{
 
+We address several limitations of \Cref{algo:bvi}.\\
+
+\begin{minipage}{0.65\textwidth}
+\begin{enumerate}
+    \item The BMDP is \hlr{unknown}
+\begin{itemize}[nolistsep]
+    \item[\incarrow] Work with a \hlg{batch} of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$
+\end{itemize}
+\item $\cT$ contains an \hlr{expectation} $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$
+\begin{itemize}[nolistsep]
+    \item[\incarrow] Replace it with a \hlg{sampling} operator $\hat{\cT}$:
+\begin{equation*}
+    \hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
+\end{equation*}
+\end{itemize}
+\item $\cS$ is \hlr{continuous}
+\begin{itemize}[nolistsep]
+    \item[\incarrow] Employ \hlg{function approximation} $Q_\theta$, and minimise a regression loss 
+    $$\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, r, \os')||_2^2$$
+\end{itemize}
+\end{enumerate}
+
+\end{minipage}
+\begin{minipage}{0.34\textwidth}
+\begin{algorithm}[H]
+\DontPrintSemicolon
+\KwData{$\cD$}
+\KwResult{$Q^*$}
+$Q_{\theta_0} \leftarrow 0$\;
+\Repeat{convergence}{
+    $\theta_{k+1} \leftarrow \argmin_\theta \cL(Q_\theta, \hat{\cT} Q_{\theta_{k}}; \cD)$\;
+}
+\caption{Budgeted Fitted-Q Iteration}
+\label{algo:bftq}
+\end{algorithm}
+\vspace{1em}
+\begin{enumerate}[nolistsep]
+    \item[4.] How to \hlr{collect the batch} $\cD$?
+    \begin{itemize}[nolistsep]
+        \item[\incarrow] We propose a \hlg{risk-sensitive} exploration procedure
+    \end{itemize}
+\end{enumerate}
+\end{minipage}
 }
@@
 %----------------------------------------------------------------------------------------
-%	Sample Complexity
+%	Scalable Implementation
 %----------------------------------------------------------------------------------------
-\headerbox{Sample complexity}
-{name=sample,span=2,column=1,below=klolop}{ 
+\headerbox{Scalable Implementation}{name=scale,span=2,column=1,below=brl}{
+
+\begin{minipage}[t]{0.48\textwidth}
+\textbf{How to compute the greedy policy?}
+
+\begin{center}
+\includegraphics[width=0.8\linewidth]{../source/img/pi.pdf}    
+\end{center}
+\vspace*{-1em}
+\bookboxx{
+\begin{proposition*}[Hull policy]
+\label{prop:bftq_pi_hull}
+$\pi_\text{greedy}$ in \eqref{eq:pi_greedy} can be \hlg{computed explicitly}, as a mixture of two points that lie on the convex hull of $Q$.
+\end{proposition*}
 }
 
-%----------------------------------------------------------------------------------------
-%	Time and Memory Complexity
-%----------------------------------------------------------------------------------------
-\headerbox{Time and memory complexity}{name=time,span=2,column=1,below=sample}{
+\end{minipage}
+\hfill%
+\begin{minipage}[t]{0.48\textwidth}
 
+\textbf{Function approximation}
+
+\begin{center}
+\resizebox{.7\textwidth}{!}{%
+\input{../source/architecture.tex}
+}    
+\end{center}
+
+\textbf{Parallel computing}
+
+Experience collection and computation of $\pi_\text{greedy}$ can be distributed over several cores.
+
+\end{minipage}
 }
@@
 %
-\headerbox{Experiments}{name=experiments,span=2,column=1,below=time, above=bottom}{
+\headerbox{Experiments}{name=experiments,span=2,column=1,below=scale, above=bottom}{
 
+\begin{minipage}{0.49\textwidth}
+\centering
+Risk-sensitive exploration
+
+\includegraphics[width=0.3\textwidth]{source/img/test.pdf}
+\includegraphics[page=1, width=0.49\textwidth]{source/img/corridors}
+\end{minipage}
+\begin{minipage}{0.5\textwidth}
+\centering
+Pareto frontier
+
+\includegraphics[width=0.49\linewidth]{../source/img/slot-filling}
+\includegraphics[width=0.49\linewidth]{../source/img/highway}
+\end{minipage}
 }
@@
 %----------------------------------------------------------------------------------------
-% \headerbox{Acknowledgements}{name=ack,column=0,span=1,below=tools}{
-% This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
-% }
+\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=bopt}{
+This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
+}
 
@@
 %----------------------------------------------------------------------------------------
-\headerbox{References}{name=refs,column=0,span=1,below=tools, above=bottom}{
+\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
     {

### 2019-08-19 supplementary.tex
Updates from Overleaf

@@
     Q_r^*(\os, \oa) &=  \max_{\rho, \pi'} Q_r^{\rho, \pi'}(\os', \oa') \label{eq:pthm_def}\\
-    &= \max_{\rho, \pi'} R_r(\os, \oa) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa) V_r^{\rho, \pi'}(\os') \label{eq:pthm_exp}\\
-    &= R_r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
-    &= R_r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
-    &= R_r(\os, \oa) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
+    &= \max_{\rho, \pi'} R_r(s, a) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa) V_r^{\rho, \pi'}(\os') \label{eq:pthm_exp}\\
+    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
+    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
+    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
 \end{align}
@@
 \label{sec:risk-sensitive-supp}
-We recall the Risk-Sensitive Exploration algorithm in \Cref{algo:risk-sensitive-exploration}
+We recall the Risk-Sensitive Exploration in \Cref{algo:risk-sensitive-exploration}:
 

### 2019-08-19 talk/talk.tex
Updates from Overleaf

\documentclass{beamer}
\usepackage{comment}
% wow this is a hack that lets you have more definitions in latex.
% see http://www.tex.ac.uk/cgi-bin/texfaq2html?label=noroom
\usepackage{natbib}
\usepackage{etex}
\usepackage{pgffor}
\usepackage[utf8]{inputenc}
%\usepackage[cyr]{aeguill}
\reserveinserts{28}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
%\usepackage[utf8]{inputenc}
%\mode<presentation>{\usetheme{Caltech}}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{multimedia}
\usepackage{color}
\usepackage{esint}
\usepackage{stmaryrd}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage[squaren]{SIunits}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{diagbox}
\usepackage{pdfpages}
\usepackage{dsfont}
\usepackage{xcolor}
\usepackage{soul}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}

\newcommand{\mathcolorbox}[2]{\colorbox{#1}{$\displaystyle #2$}}

\input{talk/mathdef}

\setbeamertemplate{footline}{%
\hfill%
\usebeamercolor[fg]{page number in head/foot}%
\usebeamerfont{page number in head/foot}%
\insertframenumber%
%\,/\,\inserttotalframenumber
\kern1em\vskip2pt%
}

\beamertemplatenavigationsymbolsempty

\author[shortname]{
Nicolas Carrara \inst{3} \and
Edouard Leurent\inst{3,5} \and
Tanguy Urvoy \inst{1} \and
Romain Laroche \inst{2} \and
Odalric-Ambrym Maillard \inst{3} \and
Olivier Pietquin \inst{3,4}}
\institute[shortinst]{\inst{1} Orange Labs\and %
\inst{2} Microsoft Montr\'eal. \and
\inst{3} Univ. Lille, CNRS, Centrale Lille, INRIA UMR 9189 - CRIStAL\and
\inst{4} Google Research, Brain Team, Paris\and
\inst{5} Renault}

\title[]{Budgeted Reinforcement Learning in Continuous State Space}
\date{10 august 2018.}
\begin{document}
    \begin{frame}
        \maketitle
        \centering
    \end{frame}

    \foreach \n in {0,1,2,3,4,5}{
    \begin{frame}{\textbf{Introduction} | Use-case}
        \begin{figure}
            \centering
            \includegraphics[scale=0.5]{talk/img/\n}
        \end{figure}
    \end{frame}
    }

    \begin{frame}{\textbf{Introduction} | Use-case}
        \begin{block}{Problem}
            Given past trajectories, find a way to:
            \begin{itemize}
                \pause\item gather gold as much as possible;
                \pause\item limit the villager casualties under some budget;
                \pause\item being able to change the budget in real time.
            \end{itemize}

        \end{block}

        \pause
        \begin{block}{Solution}
            \begin{itemize}
                \item This problem can be cast as a Budgeted Markov Decision Process.
            \end{itemize}
        \end{block}
    \end{frame}

    \begin{frame}{Setting}
        \begin{block}{Markov Decision Process}
            We define a MDP as a tuple $(\cS, \cA, P, R_r, \gamma)$ where:
            \begin{itemize}
                \pause\item  $\cS$ is the state space, $\cA$ the action space,
                \pause\item $R_r\in\Real^{\cS \times \cA}$ the rewards,
                \pause\item $P\in \cM(\cS)^{\cS \times \cA}$ the dynamics, %  where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$
                \pause\item and $\gamma$ the discounted factor.
            \end{itemize}
        \end{block}
        \pause
        \begin{block}{Objective}
            \begin{itemize}
                \pause\item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of rewards.
                \pause\item Find $\pi^*$ s.t $\forall s\in\cS$:
                \begin{equation}
                    \label{eq:mdp}
                    \begin{array}{lcr}
                        \displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s]
                    \end{array}
                \end{equation}

            \end{itemize}
        \end{block}
    \end{frame}

    \begin{frame}{Setting}
        \begin{block}{\textcolor{red}{Constrained} Markov Decision Process}
            We define a \textcolor{red}{CMDP} as a tuple $(\cS, \cA, P, R_r,\textcolor{red}{R_c}, \gamma,\textcolor{red}{\beta})$ where:
            \begin{itemize}
                \item  $\cS$ is the state space, $\cA$ the action space,
                \item $R_r\in\Real^{\cS \times \cA}$ the rewards, and $\textcolor{red}{R_c\in\Real^{\cS \times \cA}}$ the costs
                \item $P\in \cM(\cS)^{\cS \times \cA}$ the dynamics, %  where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$
                \item $\gamma$ the discounted factor, and $\textcolor{red}{\beta}$ the budget.
            \end{itemize}
        \end{block}

        \begin{block}{Objective}
            \begin{itemize}
                \item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of rewards.
                \item \textcolor{red}{ $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ the $\gamma$-discounted return of costs.}
                \item Find $\pi^*$ s.t $\forall s\in\cS$:
                \begin{equation}
                    \label{eq:cmdp}
                    \begin{array}{lcr}
                        \displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s] \\
                        \text{ s.t. }  \textcolor{red}{\expectedvalue[G_c^\pi | s_0=s] \leq \beta}
                    \end{array}
                \end{equation}
            \end{itemize}
        \end{block}
    \end{frame}

    \begin{frame}{Setting}
        \begin{block}{\textcolor{red}{Budgeted} Markov Decision Process}
            We define a \textcolor{red}{BMDP} as a tuple $(\cS, \cA, P, R_r,{R_c}, \gamma,\textcolor{red}{\cB})$ where:
            \begin{itemize}
                \item  $\cS$ is the state space, $\cA$ the action space,
                \item $R_r\in\Real^{\cS \times \cA}$ the rewards, and $R_c\in\Real^{\cS \times \cA}$ the costs
                \item $P\in \cM(\cS)^{\cS \times \cA}$ the dynamics, %  where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$
                \item $\gamma$ the discounted factor, and $\textcolor{red}{\cB}$ the budget space.
            \end{itemize}
        \end{block}

        \begin{block}{Objective}
            \begin{itemize}
                \item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of rewards.
                \item  $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ the $\gamma$-discounted return of costs.
                \item Find $\pi^*$ s.t $\forall (s,\textcolor{red}{\beta})\in\cS\times\textcolor{red}{\cB}$:
                \begin{equation}
                    \label{eq:bmdp}
                    \begin{array}{lcr}
                        \displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA\times\textcolor{red}{\cB})^{\cS\times\textcolor{red}{\cB}}} \expectedvalue[G_r^\pi | s_0=s,\textcolor{red}{\beta_0=\beta}] \\
                        \text{ s.t. }  \expectedvalue[G_c^\pi | s_0=s,\textcolor{red}{\beta_0=\beta}] \leq \beta
                    \end{array}
                \end{equation}
            \end{itemize}
        \end{block}
    \end{frame}

    \foreach \n in {0,1,2,3,4}{
    \begin{frame}{The Dynamic Programming solution: intuition}
        \begin{figure}
            \centering
            %%\vspace{-1.5em}
            \includegraphics[scale=0.6,page=1]{talk/img/a\n}
        \end{figure}
    \end{frame}
    }

    \begin{frame}{Augmented Settings}

        \textbf{Budgeted policies} $\pi$
        \begin{itemize}
            \pause\item Take a budget $\beta$ as an additional input
            \pause\item Output a next budget $\beta'$ 
            \pause\item $ \pi:\underbrace{(s,\beta)}_{\os} \rightarrow \underbrace{(a,\beta')}_{\oa}$
        \end{itemize}

        \textbf{Domain}
        \begin{itemize}
                \pause\item States $\ocS = \cS\times\cB$.
                \pause\item Actions $\ocA = \cA\times\cB$.
                \pause\item Dynamics $\ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) \eqdef P(s'|s, a)\delta(\beta' - \beta_a)$.
                \end{itemize}
        \textbf{2D  signals}
        \begin{enumerate}
            \pause\item Rewards $R = (R_r, R_c)$
            \pause\item Returns $G^\pi = (G_r^\pi, G_c^\pi)$
            \pause\item $V^\pi(\os) = (V_r^\pi, V_c^\pi) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right]$
                \pause\item $Q^\pi(\os, \oa)= (Q_r^\pi, Q_c^\pi) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right]$
        \end{enumerate}

        \pause\textbf{Policy Evaluation}\\

        The Bellman Expectation equations% [\cite{Bellman}]
        are preserved, and the Bellman Expectation Operator $\cT^\pi$ is a $\gamma$-contraction.
    \end{frame}

    \begin{frame}{Augmented Optimality}
        \begin{definition}In that order, we want to:
\begin{enumerate}
    \item[(i)] \pause\colorbox{red}{Respect the budget $\beta$}: 
    \begin{equation*}
    \Pi_a(\os) \eqdef \{\pi\in\Pi: V_c^\pi(s, \beta) \mathcolorbox{red}{\leq \beta}\}
    \end{equation*}
    \item[(ii)] \pause\colorbox{green}{Maximise the rewards}:
    \begin{equation*}
        V_r^*(\os) \eqdef \mathcolorbox{green}{\max}_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\quad \Pi_r(\os) \eqdef \mathcolorbox{green}{\argmax}_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
    \end{equation*}
    \item[(iii)] \pause\colorbox{yellow}{Minimise the costs}: 
    \begin{equation*}
        V_c^*(\os) \eqdef \mathcolorbox{yellow}{\min}_{\pi\in\Pi_r(\os)}  V_c^\pi(\os), \qquad\quad \Pi^*(\os) \eqdef \mathcolorbox{yellow}{\argmin}_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
    \end{equation*}
\end{enumerate}

\pause We define the budgeted action-value function $Q^*$ similarly
\end{definition}
    \end{frame}

    \begin{frame}{Budgeted Bellman Optimality Equation}
        \begin{theorem}[Budgeted Bellman Optimality Equation]
            $Q^*$ verifies the following equation:
            \begin{align*}
                Q^{*}(\os, \oa) &= \cT Q^{*}(\os, \oa)\\
                &\eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in \ocA} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'}),
            \end{align*}
            where the greedy policy $\pi_\text{greedy}$ is defined by:
            \begin{align*}
                \pi_\text{greedy}(\oa|\os; Q) \in &\mathcolorbox{yellow}{\argmin}_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa), \\
                \text{where }\quad \Pi_r^Q \eqdef &\mathcolorbox{green}{\argmax}_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \\
                & \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \mathcolorbox{red}{\leq \beta}
            \end{align*}
        \end{theorem}
    \end{frame}

    \begin{frame}{Optimality of the policy}
       \begin{proposition}[Optimality of the policy]
$\pi_\text{greedy}(\cdot~; Q^*)$ is \colorbox{yellow}{simultaneously optimal} in all states $\os\in\ocS$: $$\pi_\text{greedy}(\cdot~; Q^*)\in \mathcolorbox{yellow}{\Pi^*(\os)}$$

In particular, $V^{\pi_\text{greedy}(\cdot; Q^*)} = V^*$ and $Q^{\pi_\text{greedy}(\cdot; Q^*)}= Q^*$.
\end{proposition}

    \end{frame}

    \begin{frame}{Solving the untractable program}

        \begin{proposition}[$\pi_{\text{greedy}}=\pi_\text{hull}$]
            $\pi_\text{greedy}$ can be computed explicitly, as a mixture of two points that lie on the convex hull of $Q$.
        \end{proposition}

    \end{frame}

    \foreach \n in {0,1,2,3,4}{
    \begin{frame}{Solving the non-linear programming problem: intuition}
        \begin{figure}
            \centering
            %%\vspace{-1.5em}
            \includegraphics[scale=1.0,page=1]{talk/img/b\n}
        \end{figure}
    \end{frame}
    }

    \begin{frame}{Not a contraction}
        \begin{theorem}[Non-Contractivity]
            For any BMDP ($\cS,\cA,P,R_r,R_c,\gamma$) with $|\cA| \geq 2$, $\cT$ is not a contraction.
            $$\forall\epsilon>0, \exists Q^1,Q^2\in(\Real^2)^{\ocS\ocA}:\|\cT Q^1-\cT Q^2\|_\infty \geq \frac{1}{\epsilon}\|Q^1-Q^2\|_\infty$$
        \end{theorem}
    \end{frame}

    \foreach \n in {0,1,2}{
    \begin{frame}{Not a contraction: intuition}
        \begin{figure}
            \centering
            %%\vspace{-1.5em}
            \includegraphics[scale=0.9,page=1]{talk/img/d\n}
        \end{figure}
    \end{frame}
    }

    \begin{frame}{Contractivity on smooth $Q$-functions}
        \begin{conjecture}[Contractivity $\cL_\gamma$]
            $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
        \end{conjecture}

    \end{frame}

    \begin{frame}{Budgeted Dynamic Programming}
        \begin{algorithm}[H]
            \DontPrintSemicolon
            \KwData{$P, R_r, R_c$}
            \KwResult{$Q^*$}
            $Q_{0} \leftarrow 0$\;
            \Repeat{convergence}{
            $Q_{k+1} \leftarrow \cT Q_k$\;
            }
            \caption{Budgeted Value-Iteration}
        \end{algorithm}
    \end{frame}

    \begin{frame}{Budgeted Reinforcement Learning}

        We address several limitations of Budgeted Value-Iteration

        \begin{enumerate}
            \pause\item If the $P$, $R_r$ and $R_c$ are unknown:
            \begin{itemize}
                \pause\item Work with a {batch} of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$
                \pause\item Replace $\cT$ with a sampling operator $\hat{\cT}$:
                \begin{equation*}
                    \hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
                \end{equation*}
            \end{itemize}
            \pause\item If $\cS$ is continuous:
            \begin{itemize}
                \pause\item Employ function approximation $Q_\theta$, and minimise a regression loss
                $$\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, r, \os')||_2^2$$
            \end{itemize}

        \end{enumerate}

    \end{frame}

    \begin{frame}{Budgeted Fitted-Q}
        \begin{algorithm}[H]
            \DontPrintSemicolon
            \KwData{$\cD$}
            \KwResult{$Q^*$}
            $Q_{\theta_0} \leftarrow 0$\;
            \Repeat{convergence}{
            $\theta_{k+1} \leftarrow \argmin_\theta \cL(Q_\theta, \hat{\cT} Q_{\theta_{k}}; \cD)$\;
            }
            \caption{Budgeted Fitted-Q Iteration}
        \end{algorithm}
    \end{frame}

    \begin{frame}{More scaling}

        \begin{itemize}
            \pause\item CPU parallel computing of the targets $\sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i})\ \forall i$
            \pause\item Same for samples generation.
            \pause\item Neural Network as function approximator: 
        \end{itemize}

        \begin{center}
            \resizebox{.7\textwidth}{!}{%
            \input{architecture.tex}
            }
        \end{center}

    \end{frame}

    \begin{frame}{Experiments: performances of BFTQ}
        \begin{itemize}
            \item Baseline: $\lambda$-FTQ, Lagrangian relaxation
            \begin{itemize}
                \item $R_r(s,a) \leftarrow R_r(s,a) - \lambda R_c(s,a) \text{ where } \lambda \geq 0$
            \end{itemize}.
            \item Applications:
            \begin{itemize}
                \item dialogue systems
                \item autonomous driving
            \end{itemize}
        \end{itemize}
    \end{frame}

    \begin{frame}{Experiments: dialogue systems}
        \begin{itemize}
            \item A slot-filling problem: the agent (the dialogue system) fills a form by asking the user each slot.
            \pause\item Two ways to deal with recognition errors:
            \begin{itemize}
                \item ask to repeat with voice (safe/slow);
                \item Ask to repeat with numeric pad. (unsafe/fast).
            \end{itemize}
        \end{itemize}

    \end{frame}
    \begin{frame}{Experiments: dialogue systems}
        \begin{center}
            \includegraphics[width=\textwidth]{talk/img/slot-filling.pdf}
        \end{center}
    \end{frame}

    \begin{frame}{Experiments: autonomous driving}
        \begin{itemize}
            \item the agent (the car) is on a two-way road with a car in front of it, two solutions: 
            \begin{itemize}
                \item it can stay behind (safe/slow);
                \item it can overtake (unsafe/fast).
            \end{itemize}
        \end{itemize}
    \end{frame}
    \begin{frame}{Experiments: autonomous driving}
        \begin{center}
            \includegraphics[width=\textwidth]{talk/img/highway.pdf}
        \end{center}
    \end{frame}

    \begin{frame}{Experiments: autonomous driving}
        \href{scaling-up-brl.github.io-master/highway-env.html}{\beamergotobutton{BFTQ on the highway environment}}

    \end{frame}

    \begin{frame}{Risk-sensitive exploration}

        How to collect the batch $\cD$?

        \begin{itemize}
            \item We propose an $\epsilon$-greedy exploration procedure
            \begin{itemize}
                \pause\item Sample an initial budget $\beta_0$
                \pause\item At each step, where $\os=(s,\beta)$ only explore feasible budgets:
                \pause\begin{align*}
                    &\oa = (a, \beta_a)\sim\mathcal{U}(\Delta_{\cA\cB})\\
                    &\text{ where }  \Delta \text{ is such that }\probability{a, \beta_a|s, \beta} \text{verifies} \expectedvalue[\beta_a]\leq\beta
                \end{align*}
            \end{itemize}
        \end{itemize}

    \end{frame}

    \begin{frame}{Experiments: risk-sensitive exploration}

        \begin{itemize}
            \item Validate the risk-sensitive exploration procedure on the corridor environment
            \pause\item Learn 2 BFTQ policies with respectively:
            \begin{itemize}
                \item A batch generated by a risk-neutral $\epsilon$-greedy procedure
                \item A batch generated by a risk-sensitive $\epsilon$-greedy procedure
            \end{itemize}
        \end{itemize}
    \end{frame}

    \begin{frame}{Experiments: corridors}

        \begin{columns}
            \begin{column}{0.7\textwidth}
                \begin{itemize}
                    \item 2 corridors:
                    \begin{itemize}
                        \item high costs/high rewards around the starting state
                        \item no costs/low rewards around the starting state
                    \end{itemize}
                    \item The outermost cells are the ones yielding the most reward
                \end{itemize}
            \end{column}
            \begin{column}{0.3\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{talk/img/env-corridor.pdf}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

    \begin{frame}{Experiments: corridors}

        \begin{columns}
            \begin{column}{0.7\textwidth}
                \includegraphics[page=1, width=\textwidth]{talk/img/corridors}
            \end{column}
            \begin{column}{0.3\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{talk/img/test.pdf}
                \end{center}
            \end{column}
        \end{columns}
    \end{frame}

    \begin{frame}{Experiments: corridors}
        \href{scaling-up-brl.github.io-master/corridors.html}{\beamergotobutton{Risk-sensitive vs Risk-Neutral on the corridors environment}}

    \end{frame}

    \begin{frame}{Summary}

        \begin{enumerate}[+]
            \pause\item<1-> Budgeted Bellman Optimality Operator.
            \begin{itemize}
                \pause\item Fixed point.
                \pause\item Not a contraction but converging in practice.
            \end{itemize}
            \pause\item<2-> Scalable for RL in continuous state space.
            \begin{itemize}
                \pause\item Function approximation with Neural Network (dedicated architecture).
                \pause\item Solving of the untractable program using convex hull.
                \pause\item CPU parallel computing of the target.
                \pause\item Risk-sensitive exploration procedure.
            \end{itemize}

            \pause\item<3-> Experiments on two applications.
            \begin{itemize}
                \pause\item BFTQ reaches similar performances as Lagrangian relaxation,
                \pause\item with no need for calibration,
                \pause\item and less variance.
            \end{itemize}
        \end{enumerate}

    \end{frame}

\end{document}

### 2019-08-19 clean.py

@@ -1,3 +1,0 @@
-from subprocess import call
-
-call(['sh','compile.sh','clean'])

### 2019-08-19 main.py

@@ -1,12 +1,0 @@
-from subprocess import call
-
-call(['sh','compile.sh','compile','annexes'])
-
-call(['sh','compile.sh','compile','main'])
-call(['git','commit','-a','-m','\"update on pycharm\"'])
-call(['git','push'])
-
-# call(['sh','compile.sh','compile','analysis'])
-# call(['sh','compile.sh','compile','analysis_2'])
-
-

### 2019-09-04 paper/main.tex
Move paper to a separate folder

--- 2019-06-11 main.tex
@@
 where $\delta$ is the Dirac indicator distribution.
-% TU: on a rappelé ce que c'est qu'un MDP, on pourrait peut-être rappeler ce qu'est un Dirac ?
-% TU: si je comprend bien \beta_a = \beta' c'est le budget restant et ça n'es tplus le delta de budget. La relation avec \beta reste très "implicite".
 
@@
 \paragraph{Autonomous driving}
-In our third application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{Leurent2018} for simulated highway driving and behavioural decision-making.
+In our third application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{highway-env} for simulated highway driving and behavioural decision-making.
 We define a task that displays a clear trade-off between safety and efficiency.

### 2019-09-04 paper/supplementary.tex
Move paper to a separate folder

--- 2019-08-19 supplementary.tex
@@
 \end{align}
-Assume $Q^2\in\cB(Q^1,R)$.
-We start by showing this result for $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$ as defined in \Cref{sec:proof_pi_hull}:
+Assume $Q^2\in\cB(Q^1,R)$, we show by contradiction that $\cF^2 \in B(\cF^1, R)$.
+Indeed, assume there exists $q^1\in \cF^1$ such that $\cF^2 \cap B(q^1, R) = \emptyset$.
+Denote $q^2$ the unique point of $\cF^2$ such that $q^2_c = q^1_c$.
+By construction of $q^1$, we know that $\|q^1-q^2\|_\infty > R$.
+There are two possible cases:
+\begin{itemize}
+    \item $q^2_r > q^1_r$: this also directly implies that $q^2_r > q^1_r + R$.
+But $q^2\in\cF^2$, so there exist $q^2_1, q^2_2\in Q^{2},\lambda\in\Real$ such that $q^2 = (1-\lambda)q^2_1 + \lambda q^2_2$.
+But since $Q^2\in B(Q^1, R)$, there also exist $q_1^1, q^1_2\in Q^1$ such that $\|q^1_1-q^2_1\|_\infty \leq R$ and $\|q^1_2-q^2_2\|_\infty \leq R$, and in particular $q^1_{1r}\geq q^2_{1r}-R$ and $q^1_{2r}\geq q^2_{2r}-R$.
+But then, the point $q^{1'}=(1-\mu)q^1_1 + \mu q^1_2$ with $\mu=(q^2_c-q^1_{1c})/(q^2_{2c}-q^1_{1c})$ verifies $q^{1'}_c = q^1_c$ and $q^{1'}_r \geq q^2_r - R > q^1_r$ which contradicts the definition of $q_1\in\cF^1$ as defined in \eqref{eq:top-frontier}.
+    \item $q^2_r < q^1_r$: then the same reasoning can be applied by simply swapping the indexes 1 and 2.
+\end{itemize}
 
-Let $\os\in\ocS$ and $q^2\in\cC^2(Q^{2-})$, $\exists\lambda\in[0,1], \oa_1,\oa_2\in\ocA: q^2 = (1-\lambda)Q^2(\os,\oa_1) + \lambda Q^2(\os,\oa_2)$.
+% We start by showing this result for $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$ as defined in \Cref{sec:proof_pi_hull}:
+% Let $\os\in\ocS$ and $q^2\in\cC^2(Q^{2-})$, $\exists\lambda\in[0,1], \oa_1,\oa_2\in\ocA: q^2 = (1-\lambda)Q^2(\os,\oa_1) + \lambda Q^2(\os,\oa_2)$.
 Define $q^1 = (1-\lambda)Q^1(\os,\oa_1) + \lambda Q^1(\os,\oa_2)$.
 Then
-\begin{align*}
-    \|q^2-q^1\|_\infty &= \|(1-\lambda)(Q^2(\os,\oa_1) - Q^1(\os,\oa_1)) + \lambda (Q^2(\os,\oa_2) - Q^1(\os,\oa_2))\|_\infty\\
-    &\leq  (1-\lambda)\|Q^2(\os,\oa_1) - Q^1(\os,\oa_1)\|_\infty + \lambda \|Q^2(\os,\oa_2) - Q^1(\os,\oa_2)\|_\infty\\
-    &\leq (1-\lambda)R+\lambda R = R
-\end{align*}
+% \begin{align*}
+%     \|q^2-q^1\|_\infty &= \|(1-\lambda)(Q^2(\os,\oa_1) - Q^1(\os,\oa_1)) + \lambda (Q^2(\os,\oa_2) - Q^1(\os,\oa_2))\|_\infty\\
+%     &\leq  (1-\lambda)\|Q^2(\os,\oa_1) - Q^1(\os,\oa_1)\|_\infty + \lambda \|Q^2(\os,\oa_2) - Q^1(\os,\oa_2)\|_\infty\\
+%     &\leq (1-\lambda)R+\lambda R = R
+% \end{align*}
 
-It remains to show that when taking the top frontiers of the convex sets $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$, they remain at a distance of at most $R$.
+% It remains to show that when taking the top frontiers of the convex sets $\cC^2(Q^{1-})$ and $\cC^2(Q^{2-})$, they remain at a distance of at most $R$.
 
+We have shown that $\cF^2 \in B(\cF^1, R)$.
 This is illustrated in \Cref{fig:contraction_lips_hull}: given a function $Q^1$, we show the locus $\cB(Q_1,R)$ of $Q^2$.
@@
 
-For completeness, the edge case in \Cref{sec:proof_pi_hull} should be considered as well.
+Finally, consider the edge case in \Cref{sec:proof_pi_hull}: the constraint is not active, and the optimal value is simply $\argmax_{q\in\cF} q^r$.
+In particular, since we showed that $\cF^2\in B(\cF^1, R)$, and since $Q^{2*}\in \cF^2$, there exist $q^1\in \cF^1: \|Q^{2*}-q^1\|_\infty\leq R$ and in particular $Q^{1*}_r \geq q^1_r \geq Q^{2*}_r - R$.
+Reciprocally, by the same reasoning, $Q^{2*}_r \geq Q^{1*}_r - R$.
+Hence, we have that $| Q_r^{2*} - Q_r^{1*}| \leq R \leq R(L+1).$
 
@@
     \item $q^*$ belongs to the \emph{top frontier} $\cF$:
-    \begin{equation*}
+    \begin{equation}
+    \label{eq:top-frontier}
         \cF_Q = \{q\in \cC^2(Q^-): \not\exists q'\in \cC^2(Q^-): q_c=q_c'\text{ and }q_r<q_r'\}
-    \end{equation*}
+    \end{equation}
     Trivially, otherwise q' would be a better candidate than $q^*$.

### 2019-09-04 source/counter_example.py

@@ -1,40 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-# sns.set()
-# sns.set_style("white")
-plt.rc('text', usetex=True)
-plt.rc('font', family='serif', size=15)
-
-eps = 0.5
-beta = eps
-
-Q1 = np.array([[0, eps], [0, 1]])
-Q2 = np.array([[eps, 2*eps], [0, 1]])
-
-fig = plt.figure(figsize=(3, 3))
-ax = fig.add_subplot(1, 1, 1)
-plt.plot(Q1[0], Q1[1], marker='.', markersize=12)
-plt.plot(Q2[0], Q2[1], marker='.', markersize=12)
-# plt.legend(["Q1", "Q2"])
-plt.text(eps+0.01, 1/2, r'$\beta$', rotation=0, horizontalalignment="left", verticalalignment="center")
-
-plt.axis('equal')
-
-xticks = [0, eps, 2*eps]
-plt.xticks(xticks, ('0', r'$\epsilon$', r'$2\epsilon$'))
-yticks = [0, 1]
-plt.yticks(yticks, ('0', r'$\frac{1}{\gamma}$'))
-
-xticks_grid = [beta]
-ax.set_xticks(xticks_grid, minor=True)
-ax.grid(which='minor')
-plt.ylim([0, 1])
-# plt.tight_layout(pad=0.4, w_pad=0.4, h_pad=0.)
-plt.subplots_adjust(left=0.25, bottom=0.2)
-
-plt.xlabel(r"$Q_c$")
-plt.ylabel(r"$Q_r$")
-
-plt.savefig("concavity_example.pdf")
-plt.show()

### 2019-09-04 supplementary.tex
Updates from Overleaf

@@
         \item  \textbf{yes}, see \Cref{sec:proofs}. 
-We formulate a conjecture in \Cref{rmk:contractivity-smooth} but we provide a sketch of the proof in \Cref{proof:contraction-with-smooth}.\end{itemize}
+    \end{itemize}
 \end{itemize}

### 2019-09-11 paper/talk-15min/main.tex
adding talk 15 min

--- 2019-08-19 talk/talk.tex
@@
 \usepackage{graphicx}
-\usepackage{subfig}
 \usepackage{diagbox}
@@
 
+\usepackage{subfig}
+
 \newcommand{\mathcolorbox}[2]{\colorbox{#1}{$\displaystyle #2$}}
 
-\input{talk/mathdef}
+\input{mathdef}
 
@@
 
-\title[]{Budgeted Reinforcement Learning in Continuous State Space}
-\date{10 august 2018.}
+\title[]{Budgeted Reinforcement Learning in Continuous State Space (NeurIPS 2019)}
+\date{Last update: August the 11th, 2019.}
 \begin{document}
@@
 
-    \foreach \n in {0,1,2,3,4,5}{
-    \begin{frame}{\textbf{Introduction} | Use-case}
-        \begin{figure}
-            \centering
-            \includegraphics[scale=0.5]{talk/img/\n}
-        \end{figure}
+    \begin{frame}{Use-case}
+        \begin{columns}
+            \begin{column}{0.5\textwidth}
+                \begin{itemize}[<+->]
+                    \item<1,2,3,4,5> Let is an environment where:
+                    \item<2,3,4,5> the \colorbox{green}{objectif} is the gold mine,
+                    \item<3,4,5> and the \colorbox{red}{constraints} are milician patrolling randomly.
+                    \item<4,5> Given past \colorbox{blue}{trajectories},
+                    \item<5> find the \colorbox{yellow}{best strategy} to gather gold, given any villager casualty.
+                \end{itemize}
+            \end{column}
+            \begin{column}{0.5\textwidth}  %%<--- here
+                \begin{center}
+                    \foreach \n in {1,2,3,4,5}{
+                    \only<\n>{\includegraphics[width=1\textwidth]{img/aoe\n.png}}
+                    }
+                \end{center}
+            \end{column}
+        \end{columns}
+
     \end{frame}
-    }
 
-    \begin{frame}{\textbf{Introduction} | Use-case}
-        \begin{block}{Problem}
-            Given past trajectories, find a way to:
-            \begin{itemize}
-                \pause\item gather gold as much as possible;
-                \pause\item limit the villager casualties under some budget;
-                \pause\item being able to change the budget in real time.
-            \end{itemize}
+    \begin{frame}{Use-case}
+        \begin{block}{Solution}
+            This problem can be cast as a \textbf{Budgeted Markov Decision Process}.
 
@@
 
-        \pause
-        \begin{block}{Solution}
-            \begin{itemize}
-                \item This problem can be cast as a Budgeted Markov Decision Process.
-            \end{itemize}
-        \end{block}
     \end{frame}
@@
                 \pause\item  $\cS$ is the state space, $\cA$ the action space,
-                \pause\item $R_r\in\Real^{\cS \times \cA}$ the rewards,
+                \pause\item $R_r\in\Real^{\cS \times \cA}$ the \colorbox{green}{rewards},
                 \pause\item $P\in \cM(\cS)^{\cS \times \cA}$ the dynamics, %  where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$
@@
             \begin{itemize}
-                \pause\item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of rewards.
+                \pause\item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of \colorbox{green}{rewards}.
                 \pause\item Find $\pi^*$ s.t $\forall s\in\cS$:
@@
                     \begin{array}{lcr}
-                        \displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s]
+                        \displaystyle \pi^* \in \mathcolorbox{yellow}{\argmax}_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s]
                     \end{array}
@@
     \begin{frame}{Setting}
-        \begin{block}{\textcolor{red}{Constrained} Markov Decision Process}
-            We define a \textcolor{red}{CMDP} as a tuple $(\cS, \cA, P, R_r,\textcolor{red}{R_c}, \gamma,\textcolor{red}{\beta})$ where:
+        \begin{block}{\textcolor{teal}{Constrained} Markov Decision Process}
+            We define a \textcolor{teal}{CMDP} as a tuple $(\cS, \cA, P, R_r,\textcolor{teal}{R_c}, \gamma,\textcolor{teal}{\beta})$ where:
             \begin{itemize}
                 \item  $\cS$ is the state space, $\cA$ the action space,
-                \item $R_r\in\Real^{\cS \times \cA}$ the rewards, and $\textcolor{red}{R_c\in\Real^{\cS \times \cA}}$ the costs
+                \item $R_r\in\Real^{\cS \times \cA}$ the \colorbox{green}{rewards}, and $\textcolor{teal}{R_c\in\Real^{\cS \times \cA}}$ the \colorbox{red}{costs}
                 \item $P\in \cM(\cS)^{\cS \times \cA}$ the dynamics, %  where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$
-                \item $\gamma$ the discounted factor, and $\textcolor{red}{\beta}$ the budget.
+                \item $\gamma$ the discounted factor, and $\textcolor{teal}{\beta}$ the budget.
             \end{itemize}
@@
             \begin{itemize}
-                \item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of rewards.
-                \item \textcolor{red}{ $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ the $\gamma$-discounted return of costs.}
+                \item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of \colorbox{green}{rewards}.
+                \item \textcolor{teal}{ $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ the $\gamma$-discounted return of \colorbox{red}{costs}.}
                 \item Find $\pi^*$ s.t $\forall s\in\cS$:
@@
                     \begin{array}{lcr}
-                        \displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s] \\
-                        \text{ s.t. }  \textcolor{red}{\expectedvalue[G_c^\pi | s_0=s] \leq \beta}
+                        \displaystyle \pi^* \in \mathcolorbox{yellow}{\argmax}_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s] \\
+                        \text{ s.t. }  \textcolor{teal}{\expectedvalue[G_c^\pi | s_0=s] \mathcolorbox{red}{\leq \beta}}
                     \end{array}
@@
     \begin{frame}{Setting}
-        \begin{block}{\textcolor{red}{Budgeted} Markov Decision Process}
-            We define a \textcolor{red}{BMDP} as a tuple $(\cS, \cA, P, R_r,{R_c}, \gamma,\textcolor{red}{\cB})$ where:
+        \begin{block}{\textcolor{teal}{Budgeted} Markov Decision Process}
+            We define a \textcolor{teal}{BMDP} as a tuple $(\cS, \cA, P, R_r,{R_c}, \gamma,\textcolor{teal}{\cB})$ where:
             \begin{itemize}
                 \item  $\cS$ is the state space, $\cA$ the action space,
-                \item $R_r\in\Real^{\cS \times \cA}$ the rewards, and $R_c\in\Real^{\cS \times \cA}$ the costs
+                \item $R_r\in\Real^{\cS \times \cA}$ the \colorbox{green}{rewards}, and $R_c\in\Real^{\cS \times \cA}$ the \colorbox{red}{costs}
                 \item $P\in \cM(\cS)^{\cS \times \cA}$ the dynamics, %  where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$
-                \item $\gamma$ the discounted factor, and $\textcolor{red}{\cB}$ the budget space.
+                \item $\gamma$ the discounted factor, and $\textcolor{teal}{\cB}$ the budget space.
             \end{itemize}
@@
             \begin{itemize}
-                \item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of rewards.
-                \item  $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ the $\gamma$-discounted return of costs.
-                \item Find $\pi^*$ s.t $\forall (s,\textcolor{red}{\beta})\in\cS\times\textcolor{red}{\cB}$:
+                \item $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$ the $\gamma$-discounted return of \colorbox{green}{rewards}.
+                \item  $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ the $\gamma$-discounted return of \colorbox{red}{costs}.
+                \item Find $\pi^*$ s.t $\forall (s,\textcolor{teal}{\beta})\in\cS\times\textcolor{teal}{\cB}$:
                 \begin{equation}
@@
                     \begin{array}{lcr}
-                        \displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA\times\textcolor{red}{\cB})^{\cS\times\textcolor{red}{\cB}}} \expectedvalue[G_r^\pi | s_0=s,\textcolor{red}{\beta_0=\beta}] \\
-                        \text{ s.t. }  \expectedvalue[G_c^\pi | s_0=s,\textcolor{red}{\beta_0=\beta}] \leq \beta
+                        \displaystyle \pi^* \in \mathcolorbox{yellow}{\argmax}_{\pi\in\cM(\cA\times\textcolor{teal}{\cB})^{\cS\times\textcolor{teal}{\cB}}} \expectedvalue[G_r^\pi | s_0=s,\textcolor{teal}{\beta_0=\beta}] \\
+                        \text{ s.t. }  \expectedvalue[G_c^\pi | s_0=s,\textcolor{teal}{\beta_0=\beta}] \mathcolorbox{red}{\leq \beta}
                     \end{array}
@@
     \end{frame}
-
-    \foreach \n in {0,1,2,3,4}{
-    \begin{frame}{The Dynamic Programming solution: intuition}
-        \begin{figure}
-            \centering
-            %%\vspace{-1.5em}
-            \includegraphics[scale=0.6,page=1]{talk/img/a\n}
-        \end{figure}
-    \end{frame}
-    }
 
@@
 
-        The Bellman Expectation equations% [\cite{Bellman}]
-        are preserved, and the Bellman Expectation Operator $\cT^\pi$ is a $\gamma$-contraction.
+        The Bellman Expectation equations are preserved, and the Bellman Expectation Operator $\cT^\pi$ is a $\gamma$-contraction.
     \end{frame}
@@
     \begin{frame}{Augmented Optimality}
-        \begin{definition}In that order, we want to:
+        \begin{definition}
+            In that order, we want to:
             \begin{enumerate}
@@
 
-    \begin{frame}{Optimality of the policy}
-       \begin{proposition}[Optimality of the policy]
-$\pi_\text{greedy}(\cdot~; Q^*)$ is \colorbox{yellow}{simultaneously optimal} in all states $\os\in\ocS$: $$\pi_\text{greedy}(\cdot~; Q^*)\in \mathcolorbox{yellow}{\Pi^*(\os)}$$
-
-In particular, $V^{\pi_\text{greedy}(\cdot; Q^*)} = V^*$ and $Q^{\pi_\text{greedy}(\cdot; Q^*)}= Q^*$.
-\end{proposition}
-
-    \end{frame}
-
     \begin{frame}{Solving the untractable program}
@@
             %%\vspace{-1.5em}
-            \includegraphics[scale=1.0,page=1]{talk/img/b\n}
+            \includegraphics[scale=1.0,page=1]{img/b\n}
         \end{figure}
@@
 
-    \foreach \n in {0,1,2}{
-    \begin{frame}{Not a contraction: intuition}
-        \begin{figure}
-            \centering
-            %%\vspace{-1.5em}
-            \includegraphics[scale=0.9,page=1]{talk/img/d\n}
-        \end{figure}
-    \end{frame}
-    }
-
     \begin{frame}{Contractivity on smooth $Q$-functions}
-        \begin{conjecture}[Contractivity $\cL_\gamma$]
+        \begin{theorem}[Contractivity $\cL_\gamma$]
             $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
-        \end{conjecture}
+        \end{theorem}
 
@@
                 \begin{equation*}
-                    \hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
+                    \hat{\cT} Q(\os, \oa, r, \os') \eqdef r + \gamma \sum_{\ov{a'}\in \cA} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q) Q(\ov{s'}, \ov{a'}).
                 \end{equation*}
@@
 
-    \begin{frame}{More scaling}
-
-        \begin{itemize}
-            \pause\item CPU parallel computing of the targets $\sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i})\ \forall i$
-            \pause\item Same for samples generation.
-            \pause\item Neural Network as function approximator: 
-        \end{itemize}
-
-        \begin{center}
-            \resizebox{.7\textwidth}{!}{%
-            \input{architecture.tex}
-            }
-        \end{center}
-
-    \end{frame}
+    %    \begin{frame}{More scaling}
+    %
+    %        \begin{itemize}
+    %            \pause\item CPU parallel computing of the targets $\sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i})\ \forall i$
+    %            \pause\item Same for samples generation.
+    %            \pause\item Neural Network as function approximator:
+    %        \end{itemize}
+    %
+    %        \begin{center}
+    %            \resizebox{.7\textwidth}{!}{%
+    %            \input{architecture.tex}
+    %            }
+    %        \end{center}
+    %
+    %    \end{frame}
 
@@
     \begin{frame}{Experiments: dialogue systems}
-        \begin{itemize}
-            \item A slot-filling problem: the agent (the dialogue system) fills a form by asking the user each slot.
-            \pause\item Two ways to deal with recognition errors:
-            \begin{itemize}
-                \item ask to repeat with voice (safe/slow);
-                \item Ask to repeat with numeric pad. (unsafe/fast).
-            \end{itemize}
-        \end{itemize}
-
-    \end{frame}
-    \begin{frame}{Experiments: dialogue systems}
         \begin{center}
-            \includegraphics[width=\textwidth]{talk/img/slot-filling.pdf}
+            \includegraphics[width=\textwidth]{img/slot-filling.pdf}
         \end{center}
@@
     \begin{frame}{Experiments: autonomous driving}
-        \begin{itemize}
-            \item the agent (the car) is on a two-way road with a car in front of it, two solutions: 
-            \begin{itemize}
-                \item it can stay behind (safe/slow);
-                \item it can overtake (unsafe/fast).
-            \end{itemize}
-        \end{itemize}
-    \end{frame}
-    \begin{frame}{Experiments: autonomous driving}
         \begin{center}
-            \includegraphics[width=\textwidth]{talk/img/highway.pdf}
+            \includegraphics[width=\textwidth]{img/highway.pdf}
         \end{center}
-    \end{frame}
-
-    \begin{frame}{Experiments: autonomous driving}
-        \href{scaling-up-brl.github.io-master/highway-env.html}{\beamergotobutton{BFTQ on the highway environment}}
-
     \end{frame}
@@
         \begin{itemize}
-            \item Validate the risk-sensitive exploration procedure on the corridor environment
+            \item Validate the risk-sensitive exploration procedure on a 2D world environment
             \pause\item Learn 2 BFTQ policies with respectively:
@@
 
-    \begin{frame}{Experiments: corridors}
-
-        \begin{columns}
-            \begin{column}{0.7\textwidth}
-                \begin{itemize}
-                    \item 2 corridors:
-                    \begin{itemize}
-                        \item high costs/high rewards around the starting state
-                        \item no costs/low rewards around the starting state
-                    \end{itemize}
-                    \item The outermost cells are the ones yielding the most reward
-                \end{itemize}
-            \end{column}
-            \begin{column}{0.3\textwidth}  %%<--- here
-                \begin{center}
-                    \includegraphics[width=\textwidth]{talk/img/env-corridor.pdf}
-                \end{center}
-            \end{column}
-        \end{columns}
-
-    \end{frame}
-
-    \begin{frame}{Experiments: corridors}
-
-        \begin{columns}
-            \begin{column}{0.7\textwidth}
-                \includegraphics[page=1, width=\textwidth]{talk/img/corridors}
-            \end{column}
-            \begin{column}{0.3\textwidth}  %%<--- here
-                \begin{center}
-                    \includegraphics[width=\textwidth]{talk/img/test.pdf}
-                \end{center}
-            \end{column}
-        \end{columns}
-    \end{frame}
-
-    \begin{frame}{Experiments: corridors}
-        \href{scaling-up-brl.github.io-master/corridors.html}{\beamergotobutton{Risk-sensitive vs Risk-Neutral on the corridors environment}}
-
+    \begin{frame}{Experiments: 2D world}
+        \includegraphics[page=1, width=\textwidth]{img/corridors}
     \end{frame}
@@
             \begin{itemize}
-                \pause\item Function approximation with Neural Network (dedicated architecture).
                 \pause\item Solving of the untractable program using convex hull.
+                \pause\item Risk-sensitive exploration procedure.
+                \pause \item Not discussed here :
+                \begin{itemize}
+                    \pause\item Function approximation with dedicated Neural Network architecture.
                     \pause\item CPU parallel computing of the target.
-                \pause\item Risk-sensitive exploration procedure.
+                \end{itemize}
             \end{itemize}

### 2019-09-11 paper/talk-15min/compile.sh

@@ -1,0 +1,81 @@
+#!/bin/bash
+
+clean () {
+    rm *.aux 2> /dev/null
+    rm *.bbl 2> /dev/null
+    rm *.bcf 2> /dev/null
+    rm *.blg 2> /dev/null
+    rm *.idx 2> /dev/null
+    rm *.ilg 2> /dev/null
+    rm *.ind 2> /dev/null
+    rm *.log 2> /dev/null
+    rm *.ptc 2> /dev/null
+    rm *.glo 2> /dev/null
+    rm *.run.xml 2> /dev/null
+    rm *.toc 2> /dev/null
+    rm *.acn 2> /dev/null
+    rm *.acr 2> /dev/null
+    rm *.alg 2> /dev/null
+    rm *.glg 2> /dev/null
+    rm *.gls 2> /dev/null
+    rm main.ist 2> /dev/null
+    rm *.loa 2> /dev/null
+    rm *.lof 2> /dev/null
+    rm *.lot 2> /dev/null
+    rm *.mw 2> /dev/null
+    rm *.slg 2> /dev/null
+    rm *.slo 2> /dev/null
+    rm *.sls 2> /dev/null
+    rm *.glg-abr 2> /dev/null
+    rm *.glo-abr 2> /dev/null
+    rm *.gls-abr 2> /dev/null
+}
+
+clean
+
+rm main.pdf 2> /dev/null
+
+#sed -i 's/\\\usepackage{.*}//g' sources/*/*.tex
+
+
+texfot pdflatex main.tex
+#texfot pdflatex main1.tex
+#| grep "Output written on"
+#echo "-------------------------------------------------------"
+#echo "-------------------------------------------------------"
+##makeglossaries main
+#echo "-------------------------------------------------------"
+#echo "-------------------------------------------------------"
+##makeindex main.idx -s StyleInd.ist
+#echo "-------------------------------------------------------"
+#echo "-------------------------------------------------------"
+#texfot biber main1
+#echo "-------------------------------------------------------"
+#echo "-------------------------------------------------------"
+#texfot pdflatex  main1.tex | grep "Output written on"
+#echo "-------------------------------------------------------"
+#echo "-------------------------------------------------------"
+##makeglossaries main1
+#echo "-------------------------------------------------------"
+#echo "-------------------------------------------------------"
+##makeindex main1.idx -s StyleInd.ist
+#echo "-------------------------------------------------------"
[…]

### 2019-09-11 paper/talk-15min/main.py

@@ -1,0 +1,2 @@
+import os
+os.system("sh compile.sh")

### 2019-10-14 paper/main.tex
Updates from Overleaf

@@
 % [preprint] option:
-\usepackage[preprint,nonatbib]{neurips_2019}
+\usepackage[nonatbib,final]{neurips_2019}
 
@@
 This observation allows us to introduce natural extensions of Deep Reinforcement Learning algorithms to address large-scale BMDPs.
-We validate our approach on two simulated applications: spoken dialogue and autonomous driving.
+We validate our approach on two simulated applications: spoken dialogue and autonomous driving\footnote{Code and instructions available at \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#how-to-reproduce}{this web address} as required by the \href{https://docs.google.com/document/d/1SBCfisdLU2A_vb7K7wI8oRBin8ZA4cA1BbasqfQr5m0/edit}{Code Submission Policy}.}.
 \end{abstract}
@@
 Second, we use this operator in \BFTQ, a batch Reinforcement Learning algorithm, for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure.
-Third, we scale this algorithm to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
+Third, we scale this algorithm to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex programming, a risk-sensitive exploration procedure, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing.
 Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.
@@
 \end{equation}
-We restrict $\ocS$ to feasible budgets only: $\ocS_f \eqdef \{(s,\beta)\in\ocS: \exists \pi\in\Pi, V_c^\pi(s) \geq \beta\}$ that we assume is non-empty for the BMDP to admit a solution.
+We restrict $\ocS$ to feasible budgets only: $\ocS_f \eqdef \{(s,\beta)\in\ocS: \exists \pi\in\Pi, V_c^\pi(s) \leq \beta\}$ that we assume is non-empty for the BMDP to admit a solution.
 We still write $\ocS$ in place of $\ocS_f$ for brevity of notations.
@@
 Unfortunately, as $\cT$ is not a contraction, we can guarantee neither the convergence of \Cref{algo:bvi} nor the unicity of its fixed points.
-Despite those theoretical limitations, we empirically observed the convergence to a fixed point in our experiments (\Cref{sec:experiements}).
+Despite those theoretical limitations, we empirically observed the convergence to a fixed point in our experiments (\Cref{sec:experiments}).
 We conjecture a possible explanation:
 
-\begin{remark}[Contractivity of $\cT$ on smooth $Q$-functions]
+\begin{theorem}[Contractivity of $\cT$ on smooth $Q$-functions]
 \label{rmk:contractivity-smooth}
-We conjecture that $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
-We lengthily discuss some intuition on why that should be the case in \Cref{proof:contraction-with-smooth}.
+The operator $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
+%We lengthily discuss some intuition on why that should be the case in \Cref{proof:contraction-with-smooth}.
+\end{theorem}
+
+\begin{remark}
+Because of the budget dynamics $\beta' = \beta_a$ in \eqref{eq:dynamics}, we can see in \eqref{eq:bellman_expectation} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
+We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
 \end{remark}
-
-% Because of the budget dynamics $\beta' = \beta_a$ in \eqref{eq:dynamics}, we can see in \eqref{eq:bellman_expectation} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
-We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
 
@@
 
-When the BMDP is unknown, we need to adapt \Cref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$ collected by interaction with the environment.
+When the BMDP is unknown, we need to adapt \Cref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i')\}_{i\in [1,N]}$ collected by interaction with the environment.
 Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$.
@@
 \begin{equation*}
-    \hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
+    \hat{\cT} Q(\os, \oa, r, \os') \eqdef r + \gamma \sum_{\ov{a'}\in \ocA} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q) Q(\ov{s'}, \ov{a'}).
 \end{equation*}
@@
 In order to run \Cref{algo:bftq}, we must first gather a batch of samples $\mathcal{D}$.
+The following strategy is motivated by the intuition that a wide variety of risk levels needs to be experienced during training, which can be achieved by enforcing the risk constraints during data collection.
 Ideally we would need samples from the asymptotic state-budget distribution $\lim_{t\rightarrow\infty}\probability{\os_t}$ induced by an optimal policy $\pi^*$ given an initial distribution $\probability{\os_0}$, but as we are actually building this policy, it is not possible.
@@
 The batch $\cD$ is split into several mini-batches generated sequentially, and $\pi_\text{greedy}$ is updated by running \Cref{algo:bftq} on $\cD$ upon mini-batch completion.
-$\pi_\text{rand}$ is designed to obtain trajectories that only explore feasible budgets: we impose that the joint distribution $\probability{a, \beta_a|s, \beta}$ verifies $\expectedvalue[\beta_a]\leq\beta$.
+$\pi_\text{rand}$ should only pick augmented actions that are admissible candidates for $\pi_\text{greedy}$.
+To that extent $\pi_\text{rand}$ is designed to obtain trajectories that only explore feasible budgets: we impose that the joint distribution $\probability{a, \beta_a|s, \beta}$ verifies $\expectedvalue[\beta_a]\leq\beta$.
 This condition defines a probability simplex $\Delta_{\ocA}$ from which we sample uniformly.
@@
 Importantly, we pick a uniform distribution $\probability{\beta_0} = \cU(\cB)$ so that the entire range of risk-level is explored, and not only reward-seeking behaviours as would be the case with a traditional risk-neutral $\epsilon$-greedy strategy.
-The pseudo-code of our exploration procedure is shown in \Cref{algo:risk-sensitive-exploration} in \Cref{sec:risk-sensitive-supp}.
+The pseudo-code of our exploration procedure is shown in \Cref{algo:risk-sensitive-exploration}.% in \Cref{sec:risk-sensitive-supp}.
+
+%\section{Risk-Sensitive Exploration}
+%\label{sec:risk-sensitive-supp}
+%We recall the Risk-Sensitive Exploration in \Cref{algo:risk-sensitive-exploration}:
+
+\input{source/risk-sensitive-explo-pseudo-code.tex}
 
@@
 As stated in \Cref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested constrained optimisation program \eqref{eq:pi_greedy}.
-% In order to apply $\cT$ in \eqref{eq:bellman-optimality}, it is necessary to compute the greedy policy $\pi_\text{greedy}(s; Q)$.
-Contrary to classical Dynamic Programming where it takes the simple form $\argmax_{a\in\cA} Q(s, a)$, in a budgeted setting is not trivial since it requires solving the nested constrained optimisation program \eqref{eq:pi_greedy}.
 However, it can be solved efficiently by exploiting the \emph{structure} of the set of solutions with respect to $\beta$, that is, concave and increasing.
@@
 }}
-\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(\os,\oa))$\tcp*[f]{budget $\beta$ always respected}}
+\Return $\delta(\oa - \argmax_{\oa} Q_r(\os,\oa))$\tcp*[f]{budget $\beta$ always respected}
 \caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
@@
 
-The computation of $\pi_\text{hull}$ in \Cref{algo:pi_hull} is illustrated in \Cref{fig:hull}.
-
 %\begin{minipage}[t]{0.5\textwidth}
 %\vspace{0pt}
-%\begin{wrapfigure}{r}{0.5\textwidth}
-\begin{figure}[ht]
+\begin{wrapfigure}{r}{0.6\textwidth}
+\vspace{-15pt}
+%\begin{figure}[ht]
     \centering
-    \includegraphics[width=0.7\linewidth]{source/img/pi.pdf}
+    \includegraphics[width=1.0\linewidth]{source/img/pi.pdf}
     \caption{Representation of $\pi_\text{hull}$.
@@
     \label{fig:hull}
-\end{figure}
-%\end{wrapfigure}
+%\end{figure}
+\vspace{-30pt}
+\end{wrapfigure}
 %\end{minipage}
+
+The computation of $\pi_\text{hull}$ in \Cref{algo:pi_hull} is illustrated in \Cref{fig:hull}: first we get rid of dominated points.
+Then we compute the top frontier of the convex hull of the $Q$-function.
+Next, we find the two closest augmented actions $\ov{a}_1$ and $\ov{a}_2$ with cost-value $Q_c$ surrounding $\beta$:  $Q_c(\os,\oa_1)\leq\beta<Q_c(\os,\oa_2)$.
+Finally, we mix the two actions such that the expected spent budget is equal to $\beta$.
+Because of the concavity of the convex hull top frontier, any other combination of augmented actions would lead to a lower expected reward $Q_r$.
 
@@
 \section{Experiments}
-\label{sec:experiements}
+\label{sec:experiments}
 There are two hypotheses we want to validate.
@@
 $\expectedvalue[G_c|\beta=0] \simeq 0.1 $.
-This underestimation can be a consequence of two approximations: the use of the sampling operator $\hat{\cT}$ instead of the true environmental operator $\cT$, and the use of the neural network function approximation $Q_\theta$ instead of $Q$.
+This underestimation can be a consequence of two approximations: the use of the sampling operator $\hat{\cT}$ instead of the true population operator $\cT$, and the use of the neural network function approximation $Q_\theta$ instead of $Q$.
 % This is probably due to \text{highway-env} being actually infeasible for $\beta=0$, since we noticed that about $8\%$ of bad initialisations occur with the front vehicle driving slower than the minimum velocity of the controlled vehicle, leading to either a front collision or a forced change to the opposite lane, both yielding costs.
@@
 
-    \section*{Acknowledgments}
+\section*{Acknowledgements}
     %
     This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
-We thank Guillaume Gautier, Fabrice Clerot, Xuedong Shang for the helpful discussions and valuable insights.
-%\clearpage
+We thank Guillaume Gautier, Fabrice Clerot, and Xuedong Shang for the helpful discussions and valuable insights.
+\clearpage
 \bibliographystyle{named}

### 2019-10-14 paper/supplementary.tex

@@
 This paper gathers all the supplementary material and goes as follows: \Cref{sec:proofs} details all the proofs of the main results.
-\Cref{sec:risk-sensitive-supp} and \Cref{sec:bftq-full} recall respectively the scalable \BFTQ algorithm and the risk-sensitive exploration procedure.
+\Cref{sec:bftq-full} recalls the full scalable \BFTQ algorithm.
 \Cref{sec:lagragian} describes a naive alternative to \BFTQ based on Lagrangian Relaxation.
@@
 Finally we fill the Machine Learning Reproducibility Checklist and we justify each statement in \Cref{sec:ml-checklist}.
+Please note that all the materials referenced through web links are also available in the zip file of supplementary materials.
 
@@
 \begin{proof}
-This proof is the same as that in classical multi-objective MDPs.
+Thanks to the introduction of the augmented spaces $\ocS, \ocA$ and dynamics $\ov{P}$, this proof is the same as that in classical multi-objective MDPs.
 
@@
     Q_r^*(\os, \oa) &=  \max_{\rho, \pi'} Q_r^{\rho, \pi'}(\os', \oa') \label{eq:pthm_def}\\
-    &= \max_{\rho, \pi'} R_r(s, a) + \gamma \sum_{\os'\in\cS} P(\os' | \os, \oa) V_r^{\rho, \pi'}(\os') \label{eq:pthm_exp}\\
-    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
-    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
-    &= R_r(s, a) + \gamma \sum_{\os'\in\cS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
+    &= \max_{\rho, \pi'} R_r(s, a) + \gamma \sum_{\os'\in\ocS} P(\os' | \os, \oa) V_r^{\rho, \pi'}(\os') \label{eq:pthm_exp}\\
+    &= R_r(s, a) + \gamma \sum_{\os'\in\ocS}  P(\os' | \os, \oa) \max_{\rho, \pi'} \sum_{\oa'\in\ocA} \rho(\oa' | \os')Q_r^{\pi'}(\os', \oa') \label{eq:pthm_marg}\\
+    &= R_r(s, a) + \gamma \sum_{\os'\in\ocS}  P(\os' | \os, \oa) \max_\rho\sum_{\oa'\in\ocA}\rho(\oa' | \os')\max_{\pi'\in\Pi_a(\os')}Q_r^{\pi'}(\os', \oa') \label{eq:pthm_max}\\
+    &= R_r(s, a) + \gamma \sum_{\os'\in\ocS}  P(\os' | \os, \oa) \max_\rho\expectedvalueover{\oa'\sim\rho}Q_r^*(\os', \oa') \label{eq:pthm_marg_def2}
 \end{align}
@@
 \item[\eqref{eq:pthm_max}.] \begin{itemize}
-    \item Trivially $\max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\cA} \cdot \leq \sum_{\oa'\in\cA} \max_{ \pi'\in\Pi_a(\os)} \cdot$
+    \item Trivially $\max_{\pi'\in\Pi_a(\os')} \sum_{\oa'\in\ocA} \cdot \leq \sum_{\oa'\in\ocA} \max_{ \pi'\in\Pi_a(\os)} \cdot$
     \item Let $\ov{\pi}\in\argmax_{\pi'\in\Pi_a(\os')} Q_r^{\pi'}(\os', \oa')$, then:
@@
 \label{proof:contraction-with-smooth}
+
+\begin{remark*}
+This proof makes use of insights detailed in the proof of \Cref{prop:bftq_pi_hull} (\Cref{sec:proof_pi_hull}), which we recommend the reader to consult first.
+\end{remark*}
+
 \begin{proof}
@@
 
-This sketch of proof makes use of insights detailed in the proof of \Cref{prop:bftq_pi_hull}, which we recommend the reader to consult first.
+This proof makes use of insights detailed in the proof of \Cref{prop:bftq_pi_hull}, which we recommend the reader to consult first.
 
@@
 
-\section{Risk-Sensitive Exploration}
-\label{sec:risk-sensitive-supp}
-We recall the Risk-Sensitive Exploration in \Cref{algo:risk-sensitive-exploration}:
-
-\input{source/risk-sensitive-explo-pseudo-code.tex}
+%\section{Risk-Sensitive Exploration}
+%\label{sec:risk-sensitive-supp}
+%We recall the Risk-Sensitive Exploration in %\Cref{algo:risk-sensitive-exploration}:
+%\input{source/risk-sensitive-explo-pseudo-code.tex}
 
@@
 We compare two approaches for constructing a batch of samples.
-The animations from the html page \url{exploration.html} display the trajectories collected in each intermediate sub-batch.
+The animations  \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#risk-sensitive-exploration}{from this html page} display the trajectories collected in each intermediate sub-batch.
 The first row corresponds to a classical  risk neutral epsilon-greedy exploration policy while the second row showcases a risk-sensitive exploration strategy introduced in the paper.
@@
 On the \texttt{highway-env} , the budgeted agents display a wide variety of behaviours.
-Animations are displayed on the html page \url{highway-env.html}.
+Animations are displayed \href{https://github.com/budgeted-rl/budgeted-rl.github.io#driving-styles}{on this html page}.
 When $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic.
@@
 
-Animations are displayed on the html page \url{corridors.html} for the \texttt{corridors} environment.
+Animations are displayed  \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#optimal-budgeted-policies-learnt-with-a-risk-sensitive-exploration}{on this html page} for the \texttt{corridors} environment.
 When the budget is low, the agent takes the safest path on the left.
@@
 
-The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \Cref{sec:experiements}.
+The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \Cref{sec:experiments}.
 
@@
 \label{subsubsec:instruction-reproducibility}
-To reproduce the result displayed in \Cref{sec:experiements}, first install the following conventional libraries for python3: pycairo, numpy, scipy and pytorch.
+To reproduce the result displayed in \Cref{sec:experiments}, first install the following conventional libraries for python3: pycairo, numpy, scipy and pytorch.
 Then, execute the commands in \Cref{fig:instructions} on a Linux Operating System.
 The Graphic Processing Unit used for experiments is an NVIDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}.
+Then, follow the instructions in \Cref{fig:instructions}.
+For more details, check this \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#how-to-reproduce}{web address}.
 
@@
 pip3 install --user git+https://github.com/eleurent/rl-agents
+# clone the phd_code repository
+git clone https://github.com/ncarrara/phd_code
 # Change python path to the path of this repository
-export PYTHONPATH="code/scaling-up-brl"
+export PYTHONPATH="phd_code/ncarrara"
 # Navigate to budgeted-rl folder
-cd code/scaling-up-brl/budgeted-rl/
+cd phd_code/ncarrara/budgeted-rl/
 # Run main script using any config file
@@
     \begin{itemize}\item \textbf{yes}, see \Cref{sec:intro}, \Cref{sec:bdp}, \Cref{sec:brl},
-    \Cref{sec:risk-sensitive-supp} and \Cref{sec:bftq-full}.
+    \Cref{sec:exploration} and \Cref{sec:bftq-full}.
     \end{itemize}
@@
         \begin{itemize}
-            \item\textbf{yes}, see \Cref{sec:experiements} and \Cref{sec:algorithms-parameters}.
+            \item\textbf{yes}, see \Cref{sec:experiments} and \Cref{sec:algorithms-parameters}.
         \end{itemize}

### 2019-10-14 paper/main.tex
Update links

@@
 This observation allows us to introduce natural extensions of Deep Reinforcement Learning algorithms to address large-scale BMDPs.
-We validate our approach on two simulated applications: spoken dialogue and autonomous driving\footnote{Code and instructions available at \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#how-to-reproduce}{this web address} as required by the \href{https://docs.google.com/document/d/1SBCfisdLU2A_vb7K7wI8oRBin8ZA4cA1BbasqfQr5m0/edit}{Code Submission Policy}.}.
+We validate our approach on two simulated applications: spoken dialogue and autonomous driving\footnote{Videos and code are available at \href{https://budgeted-rl.github.io/}{this https URL} as required by the \href{https://docs.google.com/document/d/1SBCfisdLU2A_vb7K7wI8oRBin8ZA4cA1BbasqfQr5m0/edit}{Code Submission Policy}.}.
 \end{abstract}
@@
 
-We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available in \Cref{subsec:exploration-examples}}, and the resulting performance of a budgeted policy trained on each batch.
+We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available at \href{https://budgeted-rl.github.io/\#risk-sensitive-exploration}{this https URL}},
+%in \Cref{subsec:exploration-examples}}
+and the resulting performance of a budgeted policy trained on each batch.
 The trajectories (orange) in the risk-neutral batch are concentrated along the risky corridor (red) and ignore the safe corridor (green), which results in bad performances in the low-risk regime.
@@
 In addition, \BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance.
-Examples of policy execution can be found in \Cref{sec:bftq-executions}.
+Examples of policy execution can be found at \href{https://budgeted-rl.github.io/\#optimal-budgeted-policies}{this https URL}.
+%in \Cref{sec:bftq-executions}.
 

### 2019-10-14 paper/supplementary.tex
Update links

@@
 We compare two approaches for constructing a batch of samples.
-The animations  \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#risk-sensitive-exploration}{from this html page} display the trajectories collected in each intermediate sub-batch.
+The animations at  \href{https://budgeted-rl.github.io/\#risk-sensitive-exploration}{this https URL} display the trajectories collected in each intermediate sub-batch.
 The first row corresponds to a classical  risk neutral epsilon-greedy exploration policy while the second row showcases a risk-sensitive exploration strategy introduced in the paper.
@@
 On the \texttt{highway-env} , the budgeted agents display a wide variety of behaviours.
-Animations are displayed \href{https://github.com/budgeted-rl/budgeted-rl.github.io#driving-styles}{on this html page}.
+Animations are displayed at  \href{https://budgeted-rl.github.io/\#driving-styles}{this https URL}.
 When $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic.
@@
 
-Animations are displayed  \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#optimal-budgeted-policies-learnt-with-a-risk-sensitive-exploration}{on this html page} for the \texttt{corridors} environment.
+Animations are displayed at  \href{https://budgeted-rl.github.io/\#optimal-budgeted-policies-learnt-with-a-risk-sensitive-exploration}{this https URL} for the \texttt{corridors} environment.
 When the budget is low, the agent takes the safest path on the left.
@@
 Then, follow the instructions in \Cref{fig:instructions}.
-For more details, check this \href{https://github.com/budgeted-rl/budgeted-rl.github.io\#how-to-reproduce}{web address}.
+For more details, check  \href{https://budgeted-rl.github.io/\#how-to-reproduce}{this https URL}.
 

### 2019-10-26 paper/main.tex
Updates from Overleaf

@@
 This observation allows us to introduce natural extensions of Deep Reinforcement Learning algorithms to address large-scale BMDPs.
-We validate our approach on two simulated applications: spoken dialogue and autonomous driving\footnote{Videos and code are available at \href{https://budgeted-rl.github.io/}{this https URL} as required by the \href{https://docs.google.com/document/d/1SBCfisdLU2A_vb7K7wI8oRBin8ZA4cA1BbasqfQr5m0/edit}{Code Submission Policy}.}.
+We validate our approach on two simulated applications: spoken dialogue and autonomous driving\footnote{Videos and code are available at \url{https://budgeted-rl.github.io/}%{this https URL}
+    %as required by the \href{https://docs.google.com/document/d/1SBCfisdLU2A_vb7K7wI8oRBin8ZA4cA1BbasqfQr5m0/edit}{Code Submission Policy}
+    .}.
 \end{abstract}
@@
 \label{sec:bdp}
- We work in the space of budgeted policies, where $\pi$ both depends on $\beta$ and also outputs a next budget $\beta_a$.
+ We work in the space of budgeted policies, where a policy $\pi$ both depends on the current budget $\beta$ and also outputs a next budget $\beta_a$.
 Hence, the budget $\beta$ is neither fixed nor constant as in the CMDP setting but instead evolves as part of the dynamics.
@@
 \label{prop:greedy_optimal}
-The greedy policy $\pi_\text{greedy}(\cdot~; Q^*)$ is \emph{uniformly} optimal: for all $\os\in\ocS$, $\pi_\text{greedy}(\cdot~; Q^*)\in\Pi^*(\os)$.
+The greedy policy $\pi_\text{greedy}(\cdot~; Q^*)$ is \emph{uniformly} optimal: $\forall\os\in\ocS$, $\pi_\text{greedy}(\cdot~; Q^*)\in\Pi^*(\os)$.
 In particular, $V^{\pi_\text{greedy}(\cdot; Q^*)} = V^*$ and $Q^{\pi_\text{greedy}(\cdot; Q^*)}= Q^*$.
@@
 \label{rmk:contractivity-smooth}
-The operator $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
-%We lengthily discuss some intuition on why that should be the case in \Cref{proof:contraction-with-smooth}.
+The operator $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is Lipschitz with respect to $Q_c$":
+\begin{equation}
+    \cL_\gamma = \left\{\begin{array}{cc}
+         Q\in(\Real^2)^{\ocS\ocA}\text{ s.t. }\exists L<\frac{1}{\gamma}-1: \forall \os\in\ocS,\oa_1,\oa_2\in\ocA,   \\
+         |Q_r(\os,\oa_1) - Q_r(\os,\oa_2)| \leq L|Q_c(\os,\oa_1) - Q_c(\os,\oa_2)|
+    \end{array}\right\}
+\end{equation}
 \end{theorem}
 
-\begin{remark}
-Because of the budget dynamics $\beta' = \beta_a$ in \eqref{eq:dynamics}, we can see in \eqref{eq:bellman_expectation} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
-We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
-\end{remark}
+Thus, we expect that \Cref{algo:bvi} is likely to converge when $Q^*$ is smooth, but could diverge if the slope of $Q^*$ is too high.
+$L^2$-regularisation can be used to encourage smoothness and mitigate risk of divergence.
 
@@
 
-We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy\footnote{Animations are available at \href{https://budgeted-rl.github.io/\#risk-sensitive-exploration}{this https URL}},
-%in \Cref{subsec:exploration-examples}}
+We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: \Cref{fig:exploration} shows the set of trajectories collected by each exploration strategy.
 and the resulting performance of a budgeted policy trained on each batch.
-The trajectories (orange) in the risk-neutral batch are concentrated along the risky corridor (red) and ignore the safe corridor (green), which results in bad performances in the low-risk regime.
+The trajectories (orange) in the risk-neutral batch are concentrated along the risky corridor (right) and ignore the safe corridor (left), which results in bad performances in the low-risk regime.
 Conversely, trajectories in the risk-sensitive batch (blue) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.
@@
     \centering
-    \includegraphics[width=0.23\textwidth]{source/img/test.pdf}
+    \includegraphics[width=0.49\textwidth]{source/img/corridors_densities.pdf}
     %\includegraphics[width=0.23\textwidth]{source/img/risk-sensitive.png}
-    \includegraphics[page=1, width=0.45\textwidth]{source/img/corridors}
-    \caption{Trajectories (left) and performances (right) of two exploration strategies in the \texttt{corridors} environment. }
+    \includegraphics[page=1, width=0.49\textwidth]{source/img/corridors}
+    \caption{Density of explored states (left) and corresponding policy performances (right) of two exploration strategies in the {corridors} environment. }
     \label{fig:exploration}
@@
 % This is probably due to \text{highway-env} being actually infeasible for $\beta=0$, since we noticed that about $8\%$ of bad initialisations occur with the front vehicle driving slower than the minimum velocity of the controlled vehicle, leading to either a front collision or a forced change to the opposite lane, both yielding costs.
-Still, \BFTQ provides a better control on the expected cost of the policy, than \FTQl.
-In addition, \BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance.
-Examples of policy execution can be found at \href{https://budgeted-rl.github.io/\#optimal-budgeted-policies}{this https URL}.
-%in \Cref{sec:bftq-executions}.
+Still, \BFTQ provides a better control on the expected cost of the policy, than \FTQl. In addition, \BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance.
+
+Additional material such as videos of policy executions is provided in \Cref{sec:exp-supp}.
 

### 2019-10-26 paper/poster/poster.tex
Updates from Overleaf

--- 2019-08-19 poster/poster.tex
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{source/img/concavity_example.pdf}    
+\includegraphics[width=0.9\textwidth]{../source/img/concavity_example.pdf}    
 \end{center}
@@
 \bookboxx{
-\begin{remark*}[Contractivity on smooth $Q$-functions]
+\begin{theorem*}[Contractivity on smooth $Q$-functions]
 \label{rmk:contractivity-smooth}
-We conjecture that \hlg{$\cT$ is a contraction when restricted} to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is \hlg{$L$-Lipschitz} with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
-\end{remark*}
+\hlg{$\cT$ is a contraction when restricted} to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is \hlg{$L$-Lipschitz} with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
+\end{theorem*}
 }
@@
 
-\includegraphics[width=0.3\textwidth]{source/img/test.pdf}
-\includegraphics[page=1, width=0.49\textwidth]{source/img/corridors}
+\includegraphics[width=0.49\textwidth]{../source/img/corridors_densities.pdf}
+\includegraphics[page=1, width=0.49\textwidth]{../source/img/corridors}
 \end{minipage}

### 2019-10-26 paper/supplementary.tex
Updates from Overleaf

@@
 
+\begin{remark}
+Because of the budget dynamics $\beta' = \beta_a$ in \eqref{eq:dynamics}, we can see in \eqref{eq:bellman_expectation} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$.
+We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
+\end{remark}
+
 %\begin{minipage}[t]{0.5\textwidth}
@@
 The following section displays environments and algorithms parameters and instructions to reproduce the exact same results displayed in \Cref{sec:experiments}.
-
-\subsubsection{Environments Parameters}
-\label{sec:env-parameters}
-
-All environments parameters are displayed in \Cref{tab:param-corridors}, \Cref{tab:param-slot-filling} and \Cref{tab:param-highway-env}.
-
-\paragraph{State-Space}
-
-The states $s$ (from $\os=(s,\beta)$) of the agent are described in the following:
-
-\begin{itemize}
-    \item \texttt{Corridors}: $s = (x,y)$ where $x$ and $y$ are the 2D coordinates of the agent.
-    \item \texttt{Slot-Filling}: $s = (\text{srs},\text{min},a_u,a_s,t)$ where $\text{srs}$ is a vector of the speech recognition score for each slot, $\text{min}$ is a one hot vector describing the minimum of the $\text{srs}$ vector, $a_u$ is a one hot vector of the last user dialogue act and $a_s$ is the one hot vector of the last system dialogue act.
-Finally $t\in[0,1]$ is the fraction of the current turn with the maximum number of turns authorised.
-    \item \texttt{Highway-Env}: the positions $(x, y)$ and velocities $(\dot{x}, \dot{y})$ of every vehicle on the road.
-\end{itemize}
-
-\begin{table}[ht!]
-    \centering
-    \begin{tabularx}{1.0\textwidth}{lll}
-        \toprule
-        Parameter & Description & Value\tabularnewline
-        \midrule
-        - & Size of the environment & 7 x 6\tabularnewline
-        - & \makecell[l]{Standard deviation of the Gaussian \\noise applied to actions} & (0.25,0.25)\tabularnewline
-        H & Episode duration & 9\tabularnewline
-        \bottomrule
-    \end{tabularx}
-    \caption{Parameters of \texttt{Corridors}}
-    \label{tab:param-corridors}
-\end{table}
-
-\begin{table}[ht!]
-    \centering
-    \begin{tabularx}{1.0\textwidth}{lll}
-        \toprule
-        Parameter & Description & Value\tabularnewline
-        \midrule
-        ser & Sentence Error Rate & 0.6\tabularnewline
-        $\mu_m$& Gaussian mean for misunderstanding & -0.25\tabularnewline
-        $\mu_u$& Gaussian mean for understanding & 0.25\tabularnewline
-        $\sigma$& Gaussian standard deviation & 0.6\tabularnewline
-        $p$& Probability of hang-up & 0.25\tabularnewline
-        H & Episode duration & 10\tabularnewline
-        - & Number of slots & 3\tabularnewline
-        \bottomrule
-    \end{tabularx}
-    \caption{Parameters of \texttt{Slot-Filling}}
-    \label{tab:param-slot-filling}
-\end{table}
-
-\begin{table}[ht!]
-    \centering
-    \begin{tabularx}{1.0\textwidth}{lll}
-        \toprule
-        Parameter & Description & Value\tabularnewline
-        \midrule
-        $N_v$& Number of vehicles & 2 - 6\tabularnewline
-        $\sigma_p$& Standard deviation of vehicles initial positions & 100 m\tabularnewline
-        $\sigma_v$& Standard deviation of vehicles initial velocities & 3 m/s\tabularnewline
-        H & Episode duration & 15 s\tabularnewline
-        \bottomrule
-    \end{tabularx}
-
-    \caption{Parameters of \texttt{highway-env}}
-    \label{tab:param-highway-env}
-\end{table}
 
@@
 
+\subsubsection{Environments Parameters}
+\label{sec:env-parameters}
+
+All environments parameters are displayed in \Cref{tab:param-corridors}, \Cref{tab:param-slot-filling} and \Cref{tab:param-highway-env}.
+
+\paragraph{State-Space}
+
+The states $s$ (from $\os=(s,\beta)$) of the agent are described in the following:
+
+\begin{itemize}
+    \item \texttt{Corridors}: $s = (x,y)$ where $x$ and $y$ are the 2D coordinates of the agent.
+    \item \texttt{Slot-Filling}: $s = (\text{srs},\text{min},a_u,a_s,t)$ where $\text{srs}$ is a vector of the speech recognition score for each slot, $\text{min}$ is a one hot vector describing the minimum of the $\text{srs}$ vector, $a_u$ is a one hot vector of the last user dialogue act and $a_s$ is the one hot vector of the last system dialogue act.
+Finally $t\in[0,1]$ is the fraction of the current turn with the maximum number of turns authorised.
+    \item \texttt{Highway-Env}: the positions $(x, y)$ and velocities $(\dot{x}, \dot{y})$ of every vehicle on the road.
+\end{itemize}
+
+\begin{table}[ht!]
+    \centering
+    \begin{tabularx}{1.0\textwidth}{lll}
+        \toprule
+        Parameter & Description & Value\tabularnewline
+        \midrule
+        - & Size of the environment & 7 x 6\tabularnewline
+        - & \makecell[l]{Standard deviation of the Gaussian \\noise applied to actions} & (0.25,0.25)\tabularnewline
+        H & Episode duration & 9\tabularnewline
+        \bottomrule
+    \end{tabularx}
+    \caption{Parameters of \texttt{Corridors}}
+    \label{tab:param-corridors}
+\end{table}
+
+\begin{table}[ht!]
+    \centering
+    \begin{tabularx}{1.0\textwidth}{lll}
+        \toprule
+        Parameter & Description & Value\tabularnewline
+        \midrule
+        ser & Sentence Error Rate & 0.6\tabularnewline
+        $\mu_m$& Gaussian mean for misunderstanding & -0.25\tabularnewline
+        $\mu_u$& Gaussian mean for understanding & 0.25\tabularnewline
+        $\sigma$& Gaussian standard deviation & 0.6\tabularnewline
+        $p$& Probability of hang-up & 0.25\tabularnewline
+        H & Episode duration & 10\tabularnewline
+        - & Number of slots & 3\tabularnewline
+        \bottomrule
+    \end{tabularx}
+    \caption{Parameters of \texttt{Slot-Filling}}
+    \label{tab:param-slot-filling}
+\end{table}
+
+\begin{table}[ht!]
+    \centering
+    \begin{tabularx}{1.0\textwidth}{lll}
+        \toprule
+        Parameter & Description & Value\tabularnewline
+        \midrule
+        $N_v$& Number of vehicles & 2 - 6\tabularnewline
+        $\sigma_p$& Standard deviation of vehicles initial positions & 100 m\tabularnewline
+        $\sigma_v$& Standard deviation of vehicles initial velocities & 3 m/s\tabularnewline
+        H & Episode duration & 15 s\tabularnewline
+        \bottomrule
+    \end{tabularx}
+
+    \caption{Parameters of \texttt{highway-env}}
+    \label{tab:param-highway-env}
+\end{table}
+
 \subsubsection{Instructions for reproducibility}
@@
 # clone the phd_code repository
-git clone https://github.com/ncarrara/phd_code
+git clone https://github.com/ncarrara/budgeted-rl
 # Change python path to the path of this repository
-export PYTHONPATH="phd_code/ncarrara"
+export PYTHONPATH="budgeted-rl/ncarrara"
 # Navigate to budgeted-rl folder
-cd phd_code/ncarrara/budgeted-rl/
+cd budgeted-rl/ncarrara/budgeted-rl
 # Run main script using any config file
@@
 python3 main/egreedy/main-egreedy.py config/corridors.json 0 4
-python3 main/egreedy/main-egreedy.py config/highway-easy.json 0 10
+python3 main/egreedy/main-egreedy.py config/highway-two-way.json 0 10
 \end{lstlisting}
@@
     \item A link to a downloadable source code, with specification of all dependencies, including external libraries:
-    \begin{itemize}\item \textbf{yes}, see \Cref{subsubsec:instruction-reproducibility} and the folder \texttt{code} in the supplementary material zip file.
+    \begin{itemize}\item \textbf{yes}, see \Cref{subsubsec:instruction-reproducibility}. All resources are available at \href{https://budgeted-rl.github.io}{this https URL}.
     \end{itemize}
@@
         \begin{itemize}
-            \item  \textbf{yes}, two environments   are shipped with the supplementary material (in the \texttt{code} folder) and the third one is fetch from a public repository, see \Cref{subsubsec:instruction-reproducibility} for details.
+            \item  \textbf{yes}, all environments  are fetched from public repositories, see \Cref{subsubsec:instruction-reproducibility} for details.
         \end{itemize}
@@
     \item A description of the computing infrastructure used:     \begin{itemize}
-            \item The Graphic Processing Unit used for experiments is an NVIDIA \texttt{GeForce GTX 1080 Ti} and the Computational Processing Unit is an Intel \texttt{Xeon E7}.
+            \item The Graphic Processing Unit used for experiments is an NVIDIA \texttt{GeForce GTX 1080 Ti} and the Central Processing Unit is an Intel \texttt{Xeon E7}.
         \end{itemize}

### 2019-11-24 paper/talk/edouard/main.tex
Rewrite talk with the new inria template

\documentclass[slideopt,A4,showboxes,svgnames]{beamer}

%% list of packages here
\usepackage[absolute,showboxes,overlay]{textpos}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{multicol}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}

\usepackage{theme/beamerthemeinria}

\input{mathdef}

\input{theme/style}

\title[Budgeted Reinforcement Learning]{Budgeted Reinforcement Learning\\in Continuous State Space}
\subtitle{Sous-titre}
\date[date]{date}
\author[Carrara et al.]{Nicolas Carrara\inst{1},
	Edouard Leurent\inst{1,2},\\
	Tanguy Urvoy\inst{3},
	Romain Laroche\inst{4},\\
	Odalric Maillard\inst{1},
	Olivier Pietquin\inst{1,5}}
\institute{
	\inst{1} Inria SequeL, 
	\inst{2} Renault Group,\\
	\inst{3} Orange Labs,
	\inst{4} Microsoft Montr\'eal,\\
	\inst{5} Google Research, Brain Team}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\frame{\tocpage}

\section{Setting}
\frame{\sectionpage}

\begin{frame}{Motivation}

\begin{alertblock}{Conflicting Objectives}
Single scalar reward to represent multiple contradictory aspects (task completion vs safety)

No explicit control over the tradeoffs
\end{alertblock}

Examples?

\end{frame}

\begin{frame}{Constrained Reinforcement Learning}
\begin{block}{\only<2>{Constrained} Markov Decision Process}
	A\only<1>{n} \alert<2>{\only<2>{C}MDP} is a tuple $(\cS, \cA, P, \Rr, \only<2>{\alert<2>{R_c,}} \gamma \only<2>{\alert{, \beta}})$ with:
	\begin{multicols}{2}
			\begin{itemize}
			\item Rewards $\Rr\in\Real^{\cS \times \cA}$
			\item[]
			\item<2> \alert<2>{Costs $R_c\in\Real^{\cS \times \cA}$}
			\item<2> \alert{Budget $\beta$}
		\end{itemize}
	\end{multicols}
\end{block}
\begin{block}{Objective}
Find $\pi^*$ such that $\forall s\in\cS$:
	\begin{equation*}
	\begin{array}{lcr}
	\displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA)^{\cS}} \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \Rr(s_t, a_t) \condbar s_0=s \right]\\
	\visible<2>{
	\text{ s.t. }  \displaystyle \alert<2>{\expectedvalue\left[\sum_{t=0}^\infty \gamma^t R_c(s_t, a_t) \condbar s_0=s\right] \leq \beta}}
	\end{array}
	\end{equation*}
\end{block}
\end{frame}

\begin{frame}{Budgeted Reinforcement Learning}
\begin{block}{Budgeted Markov Decision Process}
	A \alert{BMDP} is a tuple $(\cS, \cA, P, \Rr, \Rc, \gamma\alert{, \cB})$ with:
	\begin{multicols}{2}
		\begin{itemize}
			\item Rewards $\Rr\in\Real^{\cS \times \cA}$
			\item[]
			\item Costs $\Rc\in\Real^{\cS \times \cA}$
			\item \alert{Budget space $\cB$}
		\end{itemize}
	\end{multicols}
\end{block}
\begin{block}{Objective}
	Find $\pi^*$ such that $\forall\alert{(s,\beta)\in\cS\times\cB}$:
	\begin{equation*}
	\begin{array}{lcr}
	\displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA \alert{\times \cB})^{\cS\alert{\times \cB}}} \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R_r(s_t, a_t) \condbar s_0=s, \alert{\beta_0 = \beta}\right]\\
		\text{ s.t. }  \displaystyle \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \Rc(s_t, a_t) \condbar s_0=s, \alert{\beta_0 = \beta}\right] \leq \beta
	\end{array}
	\end{equation*}
\end{block}
\end{frame}

\begin{frame}{Augmented Settings}
Budgeted policies $\pi$
\begin{itemize}
	\item Take a budget $\beta$ as an additional input
	\item Output a next budget $\beta'$ 
	\item $\pi:\underbrace{(s,\beta)}_{\os} \rightarrow \underbrace{(a,\beta')}_{\oa}$
\end{itemize}

\end{frame}

\begin{frame}{Augmented Setting}

\begin{alertblock}{Definition (Augmented spaces)}
	\begin{itemize}
		\item States $\ocS = \cS\times\cB$.
		\item Actions $\ocA = \cA\times\cB$.
		\item Dynamics $\ov{P}$
		\vspace*{-0.5cm}
		\[\text{state } (s,\beta), \text{ action }(a, \beta_a) \rightarrow \text{ next state }\begin{cases}s'\sim P(s'|s, a)\\\beta' = \beta_a\end{cases}\]
	\end{itemize}
\end{alertblock}

\begin{alertblock}{Definition (Augmented signals)}
	\begin{enumerate}
		\item Rewards $R = ({\green R_r}, {\red R_c})$
		\item Returns $G^\pi = ({\green G_r^\pi}, {\red G_c^\pi}) \eqdef \sum_{t=0}^\infty \gamma^t R(\os_t,\oa_t)$
		\item Value $V^\pi(\os) = ({\green V_r^\pi}, {\red V_c^\pi}) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right]$
		\item Q-Value $Q^\pi(\os, \oa)= ({\green Q_r^\pi}, {\red Q_c^\pi}) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right]$
	\end{enumerate}
\end{alertblock}
\end{frame}

\section{Budgeted Dynamic Programming}
\frame{\sectionpage}

\begin{frame}{Policy Evaluation}
\begin{proposition}[Budgeted Bellman Expectation]
	The Bellman Expectation equations are \alert{preserved}
	\begin{align*}
	V^\pi(\os) = \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa) \\ Q^\pi(\os, \oa) = R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') 
	\end{align*}
\end{proposition}
\pause
\begin{proposition}[Contraction]
	The Bellman Expectation Operator $\cT^\pi$ is a $\gamma$-contraction.
	\begin{align*}
	\cT^\pi Q(\os, \oa) &\eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS}\sum_{\oa'\in\ocA}\ov{P}(\os'|\os, \oa)\pi(\oa'|\os') Q(\os', \oa')
	\end{align*}
\end{proposition}

\begin{itemize}
	\item[\green \checkmark] We can {\green evaluate} a budgeted policy $\pi$
\end{itemize}
\end{frame}

\begin{frame}{Budgeted Optimality}
\begin{alertblock}{Definition (Budgeted Optimality)}
	In that order, we want to:
	\begin{enumerate}
		\item[(i)] {\red Respect the budget $\beta$}: 
		\begin{equation*}
		\Pi_a(\os) \eqdef \{\pi\in\Pi: V_c^\pi(s, \beta) \hler{\leq \beta}\}
		\end{equation*}
		\item[(ii)] {\green Maximise the rewards}:
		\begin{equation*}
		V_r^*(\os) \eqdef \hleg{\max}_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \quad\quad \Pi_r(\os) \eqdef \hleg{\argmax}_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
		\end{equation*}
		\item[(iii)] \alert{Minimise the costs}: 
		\begin{equation*}
		V_c^*(\os) \eqdef \hleb{\min}_{\pi\in\Pi_r(\os)}  V_c^\pi(\os), \quad\quad \Pi^*(\os) \eqdef \hleb{\argmin}_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
		\end{equation*}
	\end{enumerate}
\end{alertblock}
We define the budgeted action-value function $Q^*$ similarly
\end{frame}

\begin{frame}{Budgeted Optimality}
\begin{theorem}[Budgeted Bellman Optimality Equation]
	$Q^*$ verifies the following equation:
	\begin{align*}
	Q^{*}(\os, \oa) &= \cT Q^{*}(\os, \oa)\\
	&\eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in \ocA} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'})
	\end{align*}
\end{theorem}
where the greedy policy {$\pi_\text{greedy}$} is defined by:
\begin{align*}
\pi_\text{greedy}(\oa|\os; Q) \in &\hleb{\argmin}_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa), \\
\text{where }\quad \Pi_r^Q \eqdef &\hleg{\argmax}_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \\
& \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \hler{\leq \beta}
\end{align*}
\end{frame}

\begin{frame}{The optimal policy}
\begin{proposition}[Optimality of the policy]
	$\pi_\text{greedy}(\cdot~; Q^*)$ is \alert{simultaneously optimal} in all states $\os\in\ocS$: $$\pi_\text{greedy}(\cdot~; Q^*)\in \alert{\Pi^*(\os)}$$

	In particular, $V^{\pi_\text{greedy}(\cdot; Q^*)} = V^*$ and $Q^{\pi_\text{greedy}(\cdot; Q^*)}= Q^*$.
\end{proposition}

\begin{proposition}[Solving the non-linear program]
	$\pi_\text{greedy}$ can be computed \alert{efficiently}, as a mixture $\pi_\text{hull}$ of two points that lie on the convex hull of $Q$.
	\[\pi_{\text{greedy}}=\pi_\text{hull}\]
\end{proposition}
\end{frame}

\foreach \n in {0,1,2,3,4}{
	\begin{frame}{Solving the non-linear program: intuition}
	\begin{figure}
		\centering
		\includegraphics[scale=1.0,page=1]{img/b\n}
	\end{figure}
\end{frame}
}

\begin{frame}{Convergence analysis}
Recall what we've shown so far:
\[\cT \xrightarrow[fixed-point]{} Q^* \xrightarrow[tractable]{} \pi_{\text{hull}}(Q^*) \xrightarrow[equal]{} \pi_{\text{greedy}}(Q^*) \xrightarrow[optimal]{} \]

\pause
We're {\green almost there}!

All that is left is to perform \alert{Fixed-Point Iteration} to compute $Q^*$.

\pause
\begin{theorem}[Non-Contractivity]
For any BMDP ($\cS,\cA,P,R_r,R_c,\gamma$) with $|\cA| \geq 2$, $\cT$ is \textbf{\red not} a contraction.
$$\forall\epsilon>0, \exists Q^1,Q^2\in(\Real^2)^{\ocS\ocA}:\|\cT Q^1-\cT Q^2\|_\infty \geq \frac{1}{\epsilon}\|Q^1-Q^2\|_\infty$$
\end{theorem}

\begin{itemize}
	\item[\red \xmark] We {\red cannot guarantee} the convergence of $\cT^n(Q_0)$ to $Q^*$
\end{itemize}
\end{frame}

\foreach \n in {0,1,2}{
	\begin{frame}{Not a contraction: intuition}
	\begin{figure}
		\centering
		\includegraphics[scale=0.9,page=1]{img/d\n}
	\end{figure}
	\end{frame}
}

\begin{frame}{Convergence analysis}
Thankfully,
\begin{theorem}[Contractivity on smooth Q-functions]
	$\cT$ \alert{is a contraction} when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
\end{theorem}
\begin{equation*}
\cL_\gamma = \left\{\begin{array}{cc}
Q\in(\Real^2)^{\ocS\ocA}\text{ s.t. }\exists L<\frac{1}{\gamma}-1: \forall \os\in\ocS,\oa_1,\oa_2\in\ocA,   \\
|Q_r(\os,\oa_1) - Q_r(\os,\oa_2)| \leq L|Q_c(\os,\oa_1) - Q_c(\os,\oa_2)|
\end{array}\right\}
\end{equation*}
\begin{itemize}
	\item[\green \checkmark] We {\green guarantee} convergence under some (strong) assumptions
	\item[\green \checkmark] We {\green observe} empirical convergence
\end{itemize}
\end{frame}

\begin{frame}{Budgeted Dynamic Programming}
\begin{center}
	\scalebox{0.9}{
		\begin{algorithm}[H]
			\DontPrintSemicolon
			\KwData{$P, R_r, R_c$}
			\KwResult{$Q^*$}
			$Q_{0} \leftarrow 0$\;
			\Repeat{convergence}{
				$Q_{k+1} \leftarrow \cT Q_k$\;
			}
			\caption{Budgeted Value-Iteration}
		\end{algorithm}
	}
\end{center}
\end{frame}

\section{Budgeted Reinforcement Learning}
\frame{\sectionpage}

\begin{frame}{Extension to the RL setting}

We address several limitations of Budgeted Value-Iteration

\begin{enumerate}[<+->]
	\item If the $P$, $R_r$ and $R_c$ are unknown:
	\begin{itemize}[<+->]
		\item Work with a {batch} of samples $\cD=\{(\os_i,\oa_i,r_i,\os_i'\}_{i\in [0,N]}$
		\item Replace $\cT$ with a sampling operator $\hat{\cT}$:
		\begin{equation*}
		\hat{\cT} Q(\os_i, \oa_i, r_i, \os'_i) \eqdef r_i + \gamma \sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i}).
		\end{equation*}
	\end{itemize}
	\item If $\cS$ is continuous:
	\begin{itemize}
		\item Employ function approximation $Q_\theta$, and minimise a regression loss
		$$\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, r, \os')||_2^2$$
	\end{itemize}

\end{enumerate}

\end{frame}

\begin{frame}{More scaling}

\begin{itemize}[<+->]
\item CPU parallel computing of the targets $\sum_{\ov{a'_i}\in \cA_i} \pi_\text{greedy}(\ov{a'_i}|\ov{s'_i}; Q) Q(\ov{s'_i}, \ov{a'_i})$
\item Same for interactions with the environment.
\item Neural Network for function approximation: 
\begin{center}
	\resizebox{.5\textwidth}{!}{%
		\input{../../source/architecture.tex}
	}
\end{center}
\end{itemize}
\end{frame}

\section{Experiments}
\frame{\sectionpage}

\begin{frame}{Performances}
\begin{itemize}
	\item Baseline: $\lambda$-FTQ, Lagrangian relaxation
	\begin{itemize}
		\item $R_r(s,a) \leftarrow R_r(s,a) - \lambda R_c(s,a) \text{ where } \lambda \geq 0$
	\end{itemize}.
\end{itemize}
\end{frame}

\begin{frame}{Dialogue systems}
A slot-filling problem: the agent (the dialogue system) fills a form by asking the user each slot. It can either:
\begin{itemize}
	\item ask to answer using {\green voice} {\green (safe/slow)};
	\item ask to answer with a {\red numeric pad} {\red (unsafe/fast)}.
\end{itemize}

\begin{center}
	\includegraphics[trim={0 0.1cm 0 0.8cm}, clip, width=0.6\textwidth]{../../source/img/slot-filling.pdf}
\end{center}
\end{frame}

\begin{frame}{Autonomous driving}

The agent (the car) is on a two-way road with a car in front of it,
\begin{itemize}
	\item it can {\green stay behind} {\green (safe/slow)};
	\item it can {\red overtake} {\red (unsafe/fast)}.
	\end{itemize}

\begin{center}
\includegraphics[trim={0 0cm 0 0.8cm}, clip, width=0.5\textwidth]{../../source/img/highway.pdf}\\
\href{https://budgeted-rl.github.io/\#driving-styles}{\includegraphics[width=0.5\linewidth]{img/highway_env}}
\end{center}

\end{frame}

\begin{frame}{Risk-sensitive exploration}
\begin{alertblock}{How to collect the batch $\cD$?}
We propose an $\epsilon$-greedy exploration procedure:
\pause
\begin{itemize}[<+->]
	\item Sample an initial budget $\beta_0\sim\cU(\cB)$
	\item At each step, where $\os=(s,\beta)$ only explore feasible budgets:
	\begin{align*}
	&\oa = (a, \beta_a)\sim\mathcal{U}(\Delta_{\cA\cB})\\
	&\text{ where }  \Delta \text{ is such that }\probability{a, \beta_a|s, \beta} \text{verifies} \expectedvalue[\beta_a]\leq\beta
	\end{align*}
\end{itemize}
\end{alertblock}
\end{frame}

\begin{frame}{Corridors}

Two corridors:
\begin{enumerate}
\item one with {\red high costs / high rewards}
\item the other with {\green no costs / low rewards}
\end{enumerate}

\begin{center}
	\includegraphics[width=0.3\textwidth]{../../source/img/corridors_paths}
\end{center}

\begin{itemize}
	\item[$\rightarrow$] Validate the \alert{risk-sensitive exploration} procedure
\end{itemize}

\end{frame}

\begin{frame}{Corridors}

\begin{columns}
\begin{column}{0.5\linewidth}
\includegraphics[trim={0.5cm 0.1cm 0.7cm 0.2cm}, clip, page=1, width=\linewidth]{../../source/img/corridors}
\end{column}
\begin{column}{0.46\linewidth}
\begin{center}
\href{https://budgeted-rl.github.io/\#risk-sensitive-exploration}{\includegraphics[trim={2cm 0.1cm 1cm 0.5cm}, clip, width=\linewidth]{../../source/img/corridors_densities.pdf}}
\end{center}
\end{column}
\end{columns}
\end{frame}

\end{document}

### 2019-11-24 paper/talk/nicolas/talk-autonomous-driving-oriented.tex
Rewrite talk with the new inria template

--- 2019-08-19 talk/talk.tex
@@
 
-\input{talk/mathdef}
+\input{mathdef}
 
@@
 \title[]{Budgeted Reinforcement Learning in Continuous State Space}
-\date{10 august 2018.}
+\date{November the 14th, 2019.}
 \begin{document}
@@
 
-    \foreach \n in {0,1,2,3,4,5}{
-    \begin{frame}{\textbf{Introduction} | Use-case}
-        \begin{figure}
-            \centering
-            \includegraphics[scale=0.5]{talk/img/\n}
-        \end{figure}
+    \begin{frame}{\textbf{Introduction} | Context}
+        \begin{itemize}
+            \item The \textbf{autonomous driving} problem is hard to solve. Why ?
+            \begin{itemize}
+                \item we need to model the physic laws
+                \item but also and the human behaviors (negociation/randomness).
+            \end{itemize}
+            \item \textbf{Reinforcement Learning} is a solution, but
+            \begin{itemize}
+                \item It only works with one reward signal.
+                \item Reward is hard to design, need a lot of tuning
+                \item It usually implies too conversatives or too risky behaviors.
+            \end{itemize}
+
+            \item \textbf{Robust Control} faces the same problem.
+            \begin{itemize}
+                \item Propagating the incertitudes implies too safe behaviors.
+                \item Models used have to be simple and it can't apply in sim2real.
+            \end{itemize}
+            \item The solution: \textbf{Safe Reinforcement Learning}
+            \begin{itemize}
+                \item Control the risk and the reward signal.
+                \item Keeps the objectives simple.
+                \item Add a new DOF: easier to design a more complexe reward.
+            \end{itemize}
+        \end{itemize}
     \end{frame}
-    }
 
-    \begin{frame}{\textbf{Introduction} | Use-case}
+    \begin{frame}{\textbf{Introduction} | Problem}
         \begin{block}{Problem}
@@
             \begin{itemize}
-                \pause\item gather gold as much as possible;
-                \pause\item limit the villager casualties under some budget;
+                \pause\item maximize a signal;
+                \pause\item limit the frequency of catastrophic events given some budget;
                 \pause\item being able to change the budget in real time.
@@
             %%\vspace{-1.5em}
-            \includegraphics[scale=0.6,page=1]{talk/img/a\n}
+            \includegraphics[scale=0.6,page=1]{img/a\n}
         \end{figure}
@@
             \pause\item Actions $\ocA = \cA\times\cB$.
-                \pause\item Dynamics $\ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) \eqdef P(s'|s, a)\delta(\beta' - \beta_a)$.
+            \pause\item Dynamics $\ov{P}$ %$\ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) \eqdef P(s'|s, a)\delta(\beta' - \beta_a)$.
         \end{itemize}
@@
             %%\vspace{-1.5em}
-            \includegraphics[scale=0.9,page=1]{talk/img/d\n}
+            \includegraphics[scale=0.9,page=1]{img/d\n}
         \end{figure}
@@
     \begin{frame}{Contractivity on smooth $Q$-functions}
-        \begin{conjecture}[Contractivity $\cL_\gamma$]
+        \begin{Theorem}[Contractivity $\cL_\gamma$]
             $\cT$ is a contraction when restricted to the subset $\cL_\gamma$ of $Q$-functions such that "$Q_r$ is $L$-Lipschitz with respect to $Q_c$", with $L<\frac{1}{\gamma}-1$.
-        \end{conjecture}
+        \end{Theorem}
 
@@
         \begin{center}
-            \includegraphics[width=\textwidth]{talk/img/slot-filling.pdf}
+            \includegraphics[width=\textwidth]{img/slot-filling.pdf}
         \end{center}
@@
         \begin{center}
-            \includegraphics[width=\textwidth]{talk/img/highway.pdf}
+            \includegraphics[width=\textwidth]{img/highway.pdf}
         \end{center}
@@
     \begin{frame}{Experiments: autonomous driving}
-        \href{scaling-up-brl.github.io-master/highway-env.html}{\beamergotobutton{BFTQ on the highway environment}}
+        \href{../../assets/highway-env.html}{\beamergotobutton{BFTQ on the highway environment}}
 
@@
                 \begin{center}
-                    \includegraphics[width=\textwidth]{talk/img/env-corridor.pdf}
+                    \includegraphics[width=\textwidth]{img/env-corridor.pdf}
                 \end{center}
@@
             \begin{column}{0.7\textwidth}
-                \includegraphics[page=1, width=\textwidth]{talk/img/corridors}
+                \includegraphics[page=1, width=\textwidth]{img/corridors}
             \end{column}
@@
                 \begin{center}
-                    \includegraphics[width=\textwidth]{talk/img/test.pdf}
+                    \includegraphics[width=\textwidth]{img/test.pdf}
                 \end{center}
@@
 
-    \begin{frame}{Experiments: corridors}
-        \href{scaling-up-brl.github.io-master/corridors.html}{\beamergotobutton{Risk-sensitive vs Risk-Neutral on the corridors environment}}
-
-    \end{frame}
+    %    \begin{frame}{Experiments: corridors}
+    %        \href{scaling-up-brl.github.io-master/corridors.html}{\beamergotobutton{Risk-sensitive vs Risk-Neutral on the corridors environment}}
+    %
+    %    \end{frame}
 

### 2019-11-24 paper/talk/edouard/main.tex
Update BMDP definition
Update examples and baseline

@@
 Examples?
+\end{frame}
 
+\begin{frame}{Examples}
+\begin{alertblock}{Dialogue systems}
+A slot-filling problem: the agent fills a form by asking the user each slot. It can either:
+\begin{itemize}
+	\item ask to answer using {\green voice} {\green (safe/slow)};
+	\item ask to answer with a {\red numeric pad} {\red (unsafe/fast)}.
+\end{itemize}
+\end{alertblock}
+
+\begin{alertblock}{Autonomous Driving}
+The agent is driving on a two-way road with a car in front of it,
+\begin{itemize}
+\item it can {\green stay behind} {\green (safe/slow)};
+\item it can {\red overtake} {\red (unsafe/fast)}.
+\end{itemize}
+\begin{center}
+\href{https://budgeted-rl.github.io/assets/highway-neutral.gif}{\includegraphics[width=0.5\linewidth]{img/highway_env}}
+\end{center}
+\end{alertblock}
 \end{frame}
@@
 \begin{frame}{Constrained Reinforcement Learning}
-\begin{block}{\only<2>{Constrained} Markov Decision Process}
+\begin{alertblock}{\only<2>{Constrained} Markov Decision Process}
 	A\only<1>{n} \alert<2>{\only<2>{C}MDP} is a tuple $(\cS, \cA, P, \Rr, \only<2>{\alert<2>{R_c,}} \gamma \only<2>{\alert{, \beta}})$ with:
@@
 	\end{multicols}
-\end{block}
+\end{alertblock}
 \begin{block}{Objective}
-Find $\pi^*$ such that $\forall s\in\cS$:
+Maximise rewards \only<2>{\alert{while keeping costs under a fixed budget}}
 	\begin{equation*}
-	\begin{array}{lcr}
-	\displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA)^{\cS}} \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \Rr(s_t, a_t) \condbar s_0=s \right]\\
+	\begin{array}{rl}
+	 \max_{\pi\in\cM(\cA)^{\cS}} &\expectedvalue\left[\sum_{t=0}^\infty \gamma^t \Rr(s_t, a_t) \condbar s_0=s \right]\\
 	\visible<2>{
-	\text{ s.t. }  \displaystyle \alert<2>{\expectedvalue\left[\sum_{t=0}^\infty \gamma^t R_c(s_t, a_t) \condbar s_0=s\right] \leq \beta}}
+	\text{ s.t. }  \displaystyle &\alert<2>{\expectedvalue\left[\sum_{t=0}^\infty \gamma^t R_c(s_t, a_t) \condbar s_0=s\right] \leq \beta}}
 	\end{array}
@@
 \begin{frame}{Budgeted Reinforcement Learning}
-\begin{block}{Budgeted Markov Decision Process}
+\begin{alertblock}{Budgeted Markov Decision Process}
 	A \alert{BMDP} is a tuple $(\cS, \cA, P, \Rr, \Rc, \gamma\alert{, \cB})$ with:
@@
 	\end{multicols}
-\end{block}
+\end{alertblock}
 \begin{block}{Objective}
-	Find $\pi^*$ such that $\forall\alert{(s,\beta)\in\cS\times\cB}$:
-	\begin{equation*}
-	\begin{array}{lcr}
-	\displaystyle \pi^* \in \argmax_{\pi\in\cM(\cA \alert{\times \cB})^{\cS\alert{\times \cB}}} \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R_r(s_t, a_t) \condbar s_0=s, \alert{\beta_0 = \beta}\right]\\
-		\text{ s.t. }  \displaystyle \expectedvalue\left[\sum_{t=0}^\infty \gamma^t \Rc(s_t, a_t) \condbar s_0=s, \alert{\beta_0 = \beta}\right] \leq \beta
+Maximise rewards while keeping costs under an \alert{adjustable} budget. $\alert{\forall \beta\in\cB},$
+	\begin{align*}
+	\begin{array}{rl}
+	 \max_{\pi\in\cM(\cA \alert{\times \cB})^{\cS\alert{\times \cB}}} &\expectedvalue\left[\sum_{t=0}^\infty \gamma^t \Rr(s_t, a_t) \condbar s_0=s, \alert{\beta_0 = \beta}\right]\\
+		\text{ s.t. } &\expectedvalue\left[\sum_{t=0}^\infty \gamma^t \Rc(s_t, a_t) \condbar s_0=s, \alert{\beta_0 = \beta}\right] \leq \beta
 	\end{array}
-	\end{equation*}
+	\end{align*}
 \end{block}
@@
 
-\begin{frame}{Performances}
+\begin{frame}{A baseline approximate solution}
+\begin{alertblock}{Lagrangian Relaxation}
+	Consider the dual problem so as to replace the hard constraint by a soft constraint penalised by a \alert{Lagrangian  multiplier $\lambda$}:
+	\begin{equation*}
+	\label{eq:lagrangian-relaxation}
+	\max_\pi \expectedvalue \sum_t {\green \gamma^t R_r(s,a)} - \alert{\lambda}{\red \gamma^t R_c(s,a)}
+	\end{equation*}
 	\begin{itemize}
-	\item Baseline: $\lambda$-FTQ, Lagrangian relaxation
-	\begin{itemize}
-		\item $R_r(s,a) \leftarrow R_r(s,a) - \lambda R_c(s,a) \text{ where } \lambda \geq 0$
-	\end{itemize}.
+		\item Train many policies $\pi_k$ with penalties $\lambda_k$ and recover the cost budgets $\beta_k$% (Theorem 4.4, \cite{Beutler85})
+		\item Very data/memory-heavy
 	\end{itemize}
+\end{alertblock}
 \end{frame}
@@
 \end{center}
-
 \end{frame}

### 2019-12-01 paper/talk/edouard/main.tex
Update talk

@@
 
-\section{Setting}
+\section{Motivation and Setting}
 \frame{\sectionpage}
 
-\begin{frame}{Motivation}
+\begin{frame}{Learning to act}
+\begin{alertblock}{Optimal Decision-Making}
+	Which action $a_t$ should we choose in state $s_t$ to maximise a cumulative reward $R$?
+	\begin{equation*}
+	\max_\pi \expectedvalueover{a_t\sim\pi(a_t|s_t)} \left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\right]
+	\end{equation*}
+\end{alertblock}
+
+\pause
+\begin{itemize}
+	\item[\green \checkmark] A very general formulation
+	\item[\green \checkmark]<2|only@2> Widely used in the industry
+	\item[\red \xmark]<3-4> {\red \textbf{Not}} widely used in the industry
+	\only<4>{\begin{itemize}
+			\item Sample efficiency
+			\item Trial and error
+			\item Unpredictable behaviour
+	\end{itemize}}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Limitation of Reinforcement Learning}
+
+Reinforcement learning relies on a \alert{single reward function $R$}
+\begin{itemize}[<+->]
+	\item[\green \checkmark] A convenient formulation, but;
+	\item[\red \xmark] $R$ is not always easy to design. 
+\end{itemize}
 
 \begin{alertblock}{Conflicting Objectives}
-Single scalar reward to represent multiple contradictory aspects (task completion vs safety)
+Complex tasks require multiple \alert{contradictory} aspects. Typically:
 
-No explicit control over the tradeoffs
+\begin{center}
+	\Large {\green Task completion} \quad vs \quad {\red Safety}
+\end{center}
 \end{alertblock}
-
-Examples?
+\begin{flushright}
+	For example...
+\end{flushright}
 \end{frame}
 
-\begin{frame}{Examples}
+\begin{frame}{Example problems with conflicts}
 \begin{alertblock}{Dialogue systems}
@@
 \end{alertblock}
-
+\pause
 \begin{alertblock}{Autonomous Driving}
@@
 \end{alertblock}
+\end{frame}
+
+\begin{frame}{Limitation of Reinforcement Learning}
+\color{gray}
+Reinforcement learning relies on a single reward function $R$
+\begin{itemize}
+	\item[\color{gray} \checkmark] \color{gray} A convenient formulation, but;
+	\item[\color{gray} \xmark] \color{gray} $R$ is not always easy to design. 
+\end{itemize}
+
+\begin{alertblock}{\color{gray}Conflicting Objectives}
+	\color{gray}
+	Complex tasks require multiple {contradictory} aspects. Typically:
+
+	\begin{center}
+		\Large { Task completion} \quad vs \quad { Safety}
+	\end{center}
+\end{alertblock}
+\begin{flushright}
+	For example...
+\end{flushright}
+
+\color{black}
+For a fixed reward function $R$,
+\begin{itemize}
+	\item[\incarrow] $\pi^*$ is only guaranteed to lie on a \alert{Pareto-optimal} curve \alert{$\Pi^*$}
+	\item[\incarrow] no control over the $\frac{\green \text{Task Completion}}{\red \text{Safety}}$ trade-off
+\end{itemize}
+\end{frame}
+
+\begin{frame}{The Pareto-optimal curve}
+\centering
+\includegraphics[width=0.9\textwidth]{img/pareto1}
+\end{frame}
+\begin{frame}{From maximal safety to minimal risk}
+\centering
+\includegraphics[width=0.9\textwidth]{img/pareto2}
+\end{frame}
+\begin{frame}{The optimal policy can move freely along $\Pi^*$}
+\centering
+\includegraphics[width=0.9\textwidth]{img/pareto3}
+\end{frame}
+\begin{frame}{How to choose a desired trade-off}
+\centering
+\includegraphics[width=0.9\textwidth]{img/pareto4}
 \end{frame}
@@
 
+\begin{frame}{We want to learn $\Pi^*$ rather than $\pi^*_\beta$}
+\centering
+\includegraphics[width=0.9\textwidth]{img/pareto5}
+\end{frame}
+
 \begin{frame}{Budgeted Reinforcement Learning}
@@
 
-\begin{frame}{Augmented Settings}
+\begin{frame}{Problem formulation}
 Budgeted policies $\pi$
@@
 \end{itemize}
+\vspace*{\baselineskip}
+\begin{flushright}
+		\incarrow~Augment the spaces with the budget $\beta$
+\end{flushright}
 
@@
 
-\begin{frame}{More scaling}
+\begin{frame}{Scalable implementation}
 

### 2020-10-19 paper/poster/poster.tex
Update webpage style

@@
 
-\documentclass[a0paper,portrait,fontscale=0.355, margin=2cm]{baposter}
+\documentclass[paperwidth=36in,paperheight=48in,portrait,fontscale=0.355, margin=2cm]{baposter}
 
@@
 \usepackage{cleveref}
+\usepackage{pifont}
 
@@
 \input{../mathdef}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{lemma}{Lemma}
+\newtheorem{proposition}{Proposition}
+\newtheorem{remark}{Remark}
+\newtheorem{property}{Property}
+\newtheorem{assumption}{Assumption}
+\newtheorem{conjecture}{Conjecture}
+\newtheorem*{definition*}{Definition}
+\newtheorem*{theorem*}{Theorem}
+\newtheorem*{proposition*}{Proposition}
+\newtheorem*{remark*}{Remark}
 
@@
 {
-Nicolas Carrara, Edouard Leurent, Romain Laroche, \\Tanguy Urvoy, Odalric-Ambrym Maillard, Olivier Pietquin
+Nicolas Carrara$^*$, Edouard Leurent$^*$, Romain Laroche, \\Tanguy Urvoy, Odalric-Ambrym Maillard, Olivier Pietquin
 \vspace{-3\baselineskip}
@@
 
-\begin{itemize}[nolistsep]
-    \item[\incarrow] \hlr{Single scalar reward} to represent \hlr{multiple} \hlr{contradictory} aspects (e.g. complete task and avoid collisions)
-    \item[\incarrow] \hlr{No control over the spread} of the performance distribution\\
+    \hlr{Single scalar} reward for \hlb{multiple contradictory} aspects 
+    \begin{center}
+    	\includegraphics[trim={0 1.7cm 0 3cm}, clip, width=0.9\linewidth]{img/pareto}
+    \end{center}
 
-\end{itemize}
-
-\textbf{Constrained MDP} $(\cS, \cA, P, R_r, R_c, \gamma)$
-\begin{itemize}[nolistsep]
+\textbf{Constrained MDP} $(\cS, \cA, P, R_r, $\hlb{$R_c$}$, \gamma, $\hlb{$\beta$}$)$
+\begin{itemize}[]
     \item $[$\cite{BEUTLER1985236,Altman95constrainedmarkov}$]$
-    \item Introduce a \hlb{cost signal} $R_c$
-    \item Constrained objective
+    \item Introduce a cost signal $R_c$ and constrained objective
 \begin{equation*}
@@
 
-\textbf{Budgeted MDP}
+\textbf{Budgeted MDP} $(\cS, \cA, P, R_r, {R_c}, \gamma, \hleb{\cB})$
 
-\begin{itemize}[nolistsep]
+\begin{itemize}[]
 	\item $[$\cite{Boutilier_Lu:uai16}$]$
-    \item We seek \hlg{one general policy} $\pi(s,\beta)$ that solves every CMDP for any $\beta$
+    \item We seek \hlg{one general policy} $\pi(s,\beta)$ that solves every CMDP for any $\beta\in\cB$
     \item[\incarrow] Can only be solved for \hlr{finite} $\cS$ and \hlr{known} $P, R_r, R_c$.
@@
 
-The Bellman Expectation equations% [\cite{Bellman}]
-are preserved, and the Bellman Expectation Operator $\cT^\pi$ is a \hlg{$\gamma$-contraction}.
+The Bellman Expectation equations are \hlg{preserved}, and the Bellman Expectation Operator $\cT^\pi$ is a \hlg{$\gamma$-contraction}.
 }
@@
     \begin{equation*}
-        V_r^*(\os) \eqdef {\max}_{\pi\in\Pi_a(\os)}  V_r^\pi(\os) \qquad\quad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
+        V_r^*(\os) \eqdef {\max}_{\pi\in\Pi_a(\os)}  V_r^\pi(\os), \quad \Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)
     \end{equation*}
@@
     \begin{equation*}
-        V_c^*(\os) \eqdef {\min}_{\pi\in\Pi_r(\os)}  V_c^\pi(\os), \qquad\quad \Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
+        V_c^*(\os) \eqdef {\min}_{\pi\in\Pi_r(\os)}  V_c^\pi(\os), \quad \Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)
     \end{equation*}
@@
 
-Despite those theoretical limitations, we observed \hlg{empirical convergence} in our experiments
+\begin{itemize}[nolistsep]
+	\item[\hlr{\xmark}] We \hlr{cannot guarantee} the convergence of $\cT^n(Q_0)$ to $Q^*$.
+\end{itemize}
 
@@
 }
+\begin{itemize}[nolistsep]
+	\item[\hlg{\checkmark}] We \hlg{guarantee} convergence under some (strong) assumptions.
+	\item[\hlg{\checkmark}] We \hlg{observe} empirical convergence.
+\end{itemize}
 }
@@
 \begin{center}
-\includegraphics[width=0.8\linewidth]{../source/img/pi.pdf}    
+\includegraphics[width=0.76\linewidth]{../source/img/pi.pdf}    
 \end{center}
@@
 \begin{center}
-\resizebox{.7\textwidth}{!}{%
+\resizebox{.65\textwidth}{!}{%
 \input{../source/architecture.tex}
@@
 %----------------------------------------------------------------------------------------
-\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=bopt}{
+\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=bopt, above=bottom}{
 This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
@@
 %----------------------------------------------------------------------------------------
-\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
-    {
-    \AtNextBibliography{\footnotesize}
-    \setlength{\bibitemsep}{3pt}
-    \printbibliography[heading=none]
-    }
-}
+%\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
+%    {
+%    \AtNextBibliography{\footnotesize}
+%    \setlength{\bibitemsep}{3pt}
+%    \printbibliography[heading=none]
+%    }
+%}
 

### 2020-10-19 paper/talk/edouard/main.tex
Minor update
Recall CRL graph when introducing BRL
Update webpage style

@@
 Reinforcement learning relies on a \alert{single reward function $R$}
-\begin{itemize}[<+->]
-	\item[\green \checkmark] A convenient formulation, but;
-	\item[\red \xmark] $R$ is not always easy to design. 
+\begin{itemize}
+	\pause\item[\green \checkmark] A convenient formulation, but;
+	\pause\item[\red \xmark] $R$ is not always easy to design. 
 \end{itemize}
-
+\pause
 \begin{alertblock}{Conflicting Objectives}
@@
 \end{alertblock}
+\pause
 \begin{flushright}
@@
 \begin{itemize}
-	\item[\incarrow] $\pi^*$ is only guaranteed to lie on a \alert{Pareto-optimal} curve \alert{$\Pi^*$}
 	\item[\incarrow] no control over the $\frac{\green \text{Task Completion}}{\red \text{Safety}}$ trade-off
-\end{itemize}
+	\item[\incarrow] $\pi^*$ is only guaranteed to lie on a \alert{Pareto-optimal} curve \alert{$\Pi^*$}\end{itemize}
 \end{frame}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{img/pareto5}
+\includegraphics<1>[width=0.9\textwidth]{img/pareto4}%
+\includegraphics<2>[width=0.9\textwidth]{img/pareto5}
 \end{frame}
@@
 
+\begin{frame}
+\centering \LARGE Thank You!
+\end{frame}
+
 \end{document}
